# Background
Here I process and explore the PRIDE human ubn build
I also map our omics-derived ensp-based sites into the PRIDE build, and 
I combine this with a mapping for the PRIDE build from uniprot to ENSP

Note that all annotation of the PRIDE build with features and other information happens in 
the script 04_featurePrep.Rmd

# Initialise
```{r}
library(tidyverse)
library(purrr)
library(ggplot2)
library(ggridges)
library(DT)
library(readxl)
library(openxlsx)

home_directory <- ""
source(paste0(home_directory, "scripts/essential_visualisations_JvG.R"))
source(paste0(home_directory, "scripts/enrichment_analysis_JvG.R"))
source("scripts/00_objects.R")
source("scripts/01_general.R")
```

Load in relevant proteomes
```{r}
##Function: read in a fasta file of protein sequences and convert them into a data frame
library(seqinr)
read_fastaseq <- function(directory){
  seqs_raw <- read.fasta(directory)
  seqs <- data.frame("sequence" = map(seqs_raw, ~paste(., collapse = "")) %>% unlist)
  seqs$sequence <- toupper(seqs$sequence)
  seqs$uniprotkb <- strsplit(rownames(seqs), "\\|") %>% map(~.[2]) %>% unlist
  return(seqs)
}
##Load proteomes
uniprotHsIso <- read_fastaseq(paste(home_directory, 
                                         "data/biol_databases/uniprot/uniprotkb_proteome_UP000005640_2025_02_19.fasta", sep = ""))
uniprotHsIsoRevi <- read_fastaseq(paste(home_directory, 
                                         "data/biol_databases/uniprot/uniprotkb_proteome_UP000005640_AND_revi_2025_02_19.fasta", sep = ""))
uniprotHsRevi <- subset(uniprotHsIsoRevi, !grepl("-", uniprotkb))
```


# Load and process relevant files
```{r}
processBuild <- function(data){
  dataProc <- data %>%
    mutate(uniprot = strsplit(Protein, "|", fixed = T) %>% map(~.[2]) %>% unlist,
           unipGene = strsplit(Protein, "|", fixed = T) %>% map(~.[3]) %>% unlist) %>%
    mutate(site = paste0(PTM_residue, Protein_pos)) %>%
    mutate(uniprot_site = paste(uniprot, site, sep = "_"),
           uniprot_pos = paste(uniprot, Protein_pos, sep = "_")) %>%
    #Remove A decoys
    subset(PTM_residue != "A") %>%
    as.data.frame
  return(dataProc)
}
#buildPos
buildPos <- read_csv("data/PRIDE/DATE/G2S1B_0.05_protein_pos_single_prot_mapping.csv") %>% processBuild
#buildPosAllprots
buildPosAllprots <- read_csv("data/PRIDE/DATE/G2S1B_0.05_protein_pos_all_prot_mapping.csv") %>% processBuild
```

Explore
```{r}
nrow(buildPos)
length(unique(paste(buildPos$Protein, buildPos$Protein_pos, sep = "")))
table(buildPos$PTM_residue)

#Are uniprots reviewed?
uniqUnips <- list("buildPos" = buildPos, "buildPosAllprots" = buildPosAllprots) %>% 
  map(~unique(.$uniprot))
map(uniqUnips, length)
map(uniqUnips, ~length(intersect(., uniprotHsIso$uniprotkb)))
map(uniqUnips, ~length(intersect(., uniprotHsIsoRevi$uniprotkb)))
```


## Export for supps
```{r}
# Flag redundant mappings
buildExp <- buildPosAllprots %>% 
  mutate(redundantSite = uniprot_pos %in% buildPos$uniprot_pos == F) %>% 
  .[order(.$redundantSite, decreasing = F), ]

# Rename
renameCols <- c(
    "uniprot" = "Protein",
    "unipGene" = "Gene Symbol",
    "site" = "Ubi-site",
    "PTM_FLR_category" = "Localisation confidence",
    "Sum_of_PSM_counts(5%FLR)" = "Sum_of_PSM_counts(5%FLR)",
    "redundantSite" = "Site excluded (redundant mapping)"
)
PRIDEcols <- colnames(buildExp) %>% .[grep("^PXD", .)]
buildExp <- buildExp[, c(names(renameCols), PRIDEcols)] %>% 
    dplyr::rename(!!setNames(names(renameCols), renameCols))
outDir = "output/data/PRIDE/SuppTable_referenceUbnome.xlsx"
sheet = "ubnome"
if (file.exists(outDir)){
  wb <- loadWorkbook(outDir)
} else {
  wb <- createWorkbook(outDir)
}
removeWorksheet(wb, sheet)
addWorksheet(wb, sheet)
writeData(wb, sheet = sheet, x = buildExp)
saveWorkbook(wb, outDir, overwrite = TRUE)
```



# Identifier mapping

## Map between the omics-derived ensp sites and the PRIDE build
Here I map the sites in Inigo's data

### Set up data
Set up Inigo's sites
```{r}
#All sites, filtered for PRIDE
inigoFiltered <- read_csv("data/features/Beltrao_sites_annotated_filtered.csv") %>%
  mutate(uniprot = gsub("_\\d+$", "", uniprot_pos))

#All sites before filtering
inigoAll <- read_csv("data/features/ubnOmics/All human sites merged WITH AGE and Q.csv") %>%
  dplyr::rename("uniprot" = "Entry")

datLst <- list("inigoFiltered" = inigoFiltered, "inigoAll" = inigoAll)

#Do the IDpos agree?
IDpos <- datLst %>% map(~unique(.$IDpos))
ggVennDiagram_nice(IDpos)

#Do the uniprot identifiers agree?
siteUnipLst <- map(datLst, ~distinct(.[, c("IDpos", "uniprot")]))
colnames(siteUnipLst$inigoAll)[2] <- "uniprot_all"
siteUnipDf <- left_join(siteUnipLst$inigoFiltered, siteUnipLst$inigoAll, by = "IDpos")
siteUnipDf$unipSame <- siteUnipDf$uniprot == siteUnipDf$uniprot_all
siteUnipDf %>% dplyr::select(-c("IDpos")) %>% distinct %>% .[, "unipSame"] %>% table


#Inspect uniprot identifiers from inigoAll. Are they reviewed/present?
inigoAllUnips <- unique(inigoAll$uniprot)
length(inigoAllUnips)
length(grep("-", inigoAllUnips))
length(intersect(inigoAllUnips, uniprotHsIso$uniprotkb))
length(intersect(inigoAllUnips, uniprotHsIsoRevi$uniprotkb))
```


### Map uniprot identifiers into the (reviewed) proteome
```{r}
#Set up uniprot identifier data
inigoAll <- inigoAll %>%
  mutate("inProteome" = uniprot %in% uniprotHsIso$uniprotkb,
         "inProteomeRevi" = uniprot %in% uniprotHsIsoRevi$uniprotkb)
inigoAllUnipdf <- inigoAll[, c("uniprot", "inProteome", "inProteomeRevi")] %>% distinct
#Set up mapping
library(UniProt.ws)
up <- UniProt.ws()


#Step 1: Map all identifiers not in the proteome
notInProteomeMap <- mapUniProt(
  from = "UniProtKB_AC-ID",
  to = "UniParc",
  query = subset(inigoAllUnipdf, inProteome == F & !is.na(uniprot))$uniprot
)
save(notInProteomeMap, file = "output/data/PRIDE/notInProteomeMap__1.RData")
notInProteomeMap <- notInProteomeMap %>%  mutate(UniProtKB = gsub("\\.\\d+", "", UniProtKB))
##Subset for those in proteome and/or reviewed
notInProteomeMap$unipsSplit <- strsplit(notInProteomeMap$UniProtKB, "; ")
notInProteomeMap$unipsSplitInproteome <- map(notInProteomeMap$unipsSplit, ~intersect(., uniprotHsIso$uniprotkb))
notInProteomeMap$unipsSplitRevi <- map(notInProteomeMap$unipsSplit, ~intersect(., uniprotHsIsoRevi$uniprotkb))
##How many hits?
map(notInProteomeMap$unipsSplitInproteome, length) %>% unlist %>% table
map(notInProteomeMap$unipsSplitRevi, length) %>% unlist %>% table
##Select matches
notInProteomeMap$finalUnip <- NA
for (i in 1:nrow(notInProteomeMap)){
  #Use reviewed of possible
  if (length(notInProteomeMap$unipsSplitRevi[[i]]) > 0){
    unips <- notInProteomeMap$unipsSplitRevi[[i]]
  #Otherwise use present but unreviewed
  } else if (length(notInProteomeMap$unipsSplitInproteome[[i]]) > 0){
    unips <- notInProteomeMap$unipsSplitInproteome[[i]]
  } else {
  #Otherwise keep the original
    notInProteomeMap$finalUnip[i] <- notInProteomeMap$From[i]
  }
  #If there are still multiple uniprots, take the first one in the PRIDE data
  if (length(unips) > 1){
    print(i)
    print(unips)
    print(intersect(unips, buildPos$uniprot))
    unips <- intersect(unips, buildPos$uniprot)
  }
  notInProteomeMap$finalUnip[i] <- unips[1]
}
#Map back
inigoAllUnipdf$uniprotNew <- inigoAllUnipdf$uniprot
for (i in 1:nrow(notInProteomeMap)){
  inigoAllUnipdf$uniprotNew[which(inigoAllUnipdf$uniprot == notInProteomeMap$From[i])] <- notInProteomeMap$finalUnip[i]
}

#Step 2: Map unreviewed using gene names
#Get gene names for all reviewed uniprots
uniprotHsIsoRevi_geneNames <- mapUniProt(
  from = "UniProtKB_AC-ID",
  to = "UniProtKB",
  query = subset(uniprotHsIsoRevi, !grepl("-", uniprotkb))$uniprotkb
)
save(uniprotHsIsoRevi_geneNames, file = "output/data/PRIDE/uniprotHsIsoRevi_geneNames__1.RData")
uniprotHsIsoRevi_geneNames$Gene.NamesLst <- strsplit(uniprotHsIsoRevi_geneNames$Gene.Names, " ")
uniprotHsIsoRevi_geneNames$firstGene <- map(uniprotHsIsoRevi_geneNames$Gene.NamesLst, ~.[1]) %>% unlist
#Get gene names for ubn unreviewed uniprots
notReviMap <- mapUniProt(
  from = "UniProtKB_AC-ID",
  to = "UniProtKB",
  query = setdiff(inigoAllUnipdf$uniprotNew, uniprotHsIsoRevi$uniprotkb) %>% .[which(!is.na(.))]
)
notReviMap$Gene.NamesLst <- strsplit(notReviMap$Gene.Names, " ")
notReviMap$firstGene <- map(notReviMap$Gene.NamesLst, ~.[1]) %>% unlist

#Overlap of all gene names
length(which(notReviMap$Gene.Names %in% uniprotHsIsoRevi_geneNames$Gene.Names))

#Subset reviewed proteome for gene names present in mapping
allMappingGenes <- unique(unlist(notReviMap$Gene.NamesLst))
uniprotHsIsoRevi_geneNames$inMapping <- map(uniprotHsIsoRevi_geneNames$Gene.NamesLst, ~length(intersect(., allMappingGenes)) > 0) %>% unlist
uniprotHsIsoRevi_geneNamesSub <- subset(uniprotHsIsoRevi_geneNames, inMapping == T)

#Multiple gene names are listed, so find matches based on when any single one lines up
notReviMap$geneMatchInds <- map(notReviMap$Gene.NamesLst, function(genes){
  return(which(uniprotHsIsoRevi_geneNamesSub$firstGene %in% genes))
})
notReviMap$geneMatchNum <- map(notReviMap$geneMatchInds, length) %>% unlist

#Get matches based on this
notReviMap$finalUnip <- NA
for (i in 1:nrow(notReviMap)){
  inds <- notReviMap$geneMatchInds[[i]]
  if (length(inds) == 1){
    notReviMap$finalUnip[i] <- uniprotHsIsoRevi_geneNamesSub$From[inds]
  } else {
    notReviMap$finalUnip[i] <- notReviMap$From[i]
  }
}

#Map back
for (i in 1:nrow(notReviMap)){
  inigoAllUnipdf$uniprotNew[which(inigoAllUnipdf$uniprot == notReviMap$From[i])] <- notReviMap$finalUnip[i]
}

#Remove isoforms
inigoAllUnipdf$uniprotNew_wIso <- inigoAllUnipdf$uniprotNew
inigoAllUnipdf$uniprotNew <- gsub("-\\d+$", "", inigoAllUnipdf$uniprotNew)

#Map into site data
inigoAllMapped <- left_join(inigoAll, inigoAllUnipdf[, c("uniprot", "uniprotNew")], by = "uniprot")
```


### Map sequence windows into uniprots
```{r}
inigoAllMapped <- inigoAllMapped[order(inigoAllMapped$uniprotNew), ]
unipsOffset <- c("", inigoAllMapped$uniprotNew[-nrow(inigoAllMapped)])
inigoAllMapped$newUnip <- inigoAllMapped$uniprotNew != unipsOffset
inigoAllMapped$uniprot_pos_matches <- NA
unipSeq <- ""
for (i in 1:nrow(inigoAllMapped)){
  #Get objects
  unip <- inigoAllMapped$uniprotNew[i]
  seqWind <- inigoAllMapped$sequence_window[i]
  #Only proceed if uniprot is in the proteome
  if (!is.na(unip) & unip %in% uniprotHsIso$uniprotkb){
    #Get uniprot sequence
    if (inigoAllMapped$newUnip[i]){
      unipSeq <- uniprotHsIso$sequence[which(uniprotHsIso$uniprotkb == unip)]
      unipSeq <- paste(c(rep("_", 7), unipSeq, rep("_", 7)), collapse = "")
    }
    #Match sequence window in
    matchesDf <- stringr::str_locate_all(unipSeq, seqWind)[[1]]
    if (nrow(matchesDf) == 0){
      matches <- NA
    } else {
      newPosns <- as.numeric(matchesDf[, 1])
      matches <- map(newPosns, ~paste0(unip, "_", .)) %>% unlist
      matches <- paste(matches, collapse = ";")
    }
    inigoAllMapped$uniprot_pos_matches[i] <- matches
  }
}

#Inspect matches
inigoAllMapped$uniprot_pos_matches_num <- strsplit(inigoAllMapped$uniprot_pos_matches, ";") %>% 
  map(function(x){
    if (is.na(x[1])){ return(0) } else { return(length(x))}
  }) %>% unlist
table(inigoAllMapped$uniprot_pos_matches_num)
##Inspect those with 0 matches
###Uniprot not in the proteome?
inigoAllMapped$uniprotNotInProt <- inigoAllMapped$uniprotNew %in% uniprotHsIso$uniprotkb == F
table(inigoAllMapped$uniprot_pos_matches_num, inigoAllMapped$uniprotNotInProt)
###Changed uniprot?
inigoAllMapped$uniprotChanged <- inigoAllMapped$uniprotNew != inigoAllMapped$uniprot
inigoAllMapped$uniprotChanged[is.na(inigoAllMapped$uniprotChanged)] <- FALSE
table(inigoAllMapped$uniprot_pos_matches_num, inigoAllMapped$uniprotChanged)

##For those with multiple matches, take those in the PRIDE atlas
buildPos$PTM_FLR_category <- factor(buildPos$PTM_FLR_category, levels = c("Gold", "Silver", "Bronze"))
buildPos <- buildPos[order(buildPos$PTM_FLR_category), ] #Order so that we pick Gold when given a choice
inigoAllMapped$uniprot_pos_matches_noMult <- inigoAllMapped$uniprot_pos_matches
for (i in 1:nrow(inigoAllMapped)){
  if (inigoAllMapped$uniprot_pos_matches_num[i] > 1){
    uniprot_pos_vec <- strsplit(inigoAllMapped$uniprot_pos_matches[i], ";")[[1]]
    uniprot_pos_vec_inPRIDE <- intersect(buildPos$uniprot_pos, uniprot_pos_vec)
    #If no matches are in PRIDE, just take the first
    if (length(uniprot_pos_vec_inPRIDE) == 0){
      inigoAllMapped$uniprot_pos_matches_noMult[i] <- uniprot_pos_vec[1]
    } else {
      inigoAllMapped$uniprot_pos_matches_noMult[i] <- uniprot_pos_vec_inPRIDE[1] #Take the first if there are still multiple
    }
  }
}
inigoAllMapped$uniprot_pos_matches_noMult_num <- strsplit(inigoAllMapped$uniprot_pos_matches_noMult, ";") %>% 
  map(function(x){
    if (is.na(x[1])){ return(0) } else { return(length(x))}
  }) %>% unlist
table(inigoAllMapped$uniprot_pos_matches_noMult_num)
```


### Overlap with PRIDE after mapping
```{r}
#Refilter inigoFiltered using the  pride set
inigoFilteredFilt <- subset(inigoFiltered, uniprot_pos %in% buildPos$uniprot_pos)
inigoAllMappedFilt <- subset(inigoAllMapped, uniprot_pos_matches_noMult %in% buildPos$uniprot_pos)
nrow(inigoFilteredFilt)
nrow(inigoAllMappedFilt)

#Map in PRIDE info
inigoAllMappedFilt <- dplyr::rename(inigoAllMappedFilt, "uniprot_pos" = "uniprot_pos_matches_noMult")
inigoAllMappedFilt <- left_join(inigoAllMappedFilt, 
                                buildPos[, c("PTM_FLR_category", "Sum_of_PSM_counts(5%FLR)", "uniprot_pos")],
                                by = "uniprot_pos") %>%
                      dplyr::rename("PRIDE_set" = "PTM_FLR_category", "PSM_sum" = "Sum_of_PSM_counts(5%FLR)")

#Overlap between inigoFiltered and inigoAll
lst <- list("filtered" = inigoFilteredFilt, "all" = inigoAllMappedFilt)
ggVennDiagram_nice(map(lst, ~.$IDpos))
inigoFilteredFilt[which(inigoFilteredFilt$IDpos %in% inigoAllMappedFilt$IDpos == F), ]
```


### Prepare for downstream analysis
```{r}
#Rename
InigoFeatMapped <- inigoAllMappedFilt %>%
  dplyr::rename("uniprotOld" = "uniprot") %>%
  dplyr::rename(
    "uniprot" = "uniprotNew"
  )
#Select columns
cols <- c("IDpos", "ENSP", "uniprot_pos", "sequence_window", "ensp_pos",
           "PRIDE_set", "PSM_sum", "common_ancestor_tax", "age_mya",
           "PSP.ubi", "PSP.act", "PSP.met", "abundance",
           "Function..CC.", "Calcium.binding", "Nucleotide.binding", "Nucleotide.binding.INSIDE",
           "FinalBox", "HotSpot.pfam", "HotSpot.name", "times.quantified",
           "times.quantified.proteasome", "proteasome.smaller.Q3", "DNAdam.bigger.Q3")
InigoFeatMapped <- InigoFeatMapped[, cols]
save(InigoFeatMapped, file = "output/data/InigoFeatMapped__3.RData")
```


### Make a map of identifiers for all human sites
```{r}
ubnOmicsSites_unipMap_hs <- inigoAllMapped %>% 
  dplyr::rename("uniprot_old" = "uniprot") %>%
  dplyr::rename("uniprot" = "uniprotNew",
                "uniprot_pos" = "uniprot_pos_matches_noMult") %>%
  dplyr::select(c("IDpos", "ENSP", "uniprot", "uniprot_pos", "sequence_window"))
save(ubnOmicsSites_unipMap_hs, file = "output/data/ubnOmicsSites_unipMap_hs__3.RData")
```


## Map between the PRIDE uniprot sites and ENSP identifiers
Here I combine the map derived above with another map from Inigo, to 
map (almost) all sites in the PRIDE build to ENSP. This is necessary 
for the conservation analysis
```{r}
# Load and set up data
buildPos_wMap <- buildPos  # use the non-redundant database
# Load site-uniprot map
enspUnipMap <- read_delim("data/conservation/final_mapping_uniprot_ensembl.tab") %>% 
    dplyr::rename(
        "uniprot" = "Uniprot_protein",
        "pos_fUniprot" = "Uniprot_position",
        "sequence_window" = "seq_window",
        "ensp" = "Ensembl_protein",
        "pos_fEnsp" = "Ensembl_position") %>% 
    mutate(uniprot_pos = paste(uniprot, pos_fUniprot, sep = "_"),
        ensp_pos = paste(ensp, pos_fEnsp, sep = "_")) %>% 
    mutate(ensp_pos = ifelse(is.na(pos_fEnsp), NA, ensp_pos))

# Add Inigo's new map
collapser <- function(x){ 
        if (is.na(x)[1]){
            return(NA)
        } else {
            return(paste(x, collapse = ";"))
        }
    }
enspUnipMap_proc <- enspUnipMap[, c("uniprot_pos", "ensp_pos")] %>% 
    group_by(uniprot_pos) %>% 
    summarise(ensp_pos = collapser(ensp_pos), .groups = "drop") %>%  #Collapse unambiguous mapping
    dplyr::rename("ensp_pos_InigoNew" = "ensp_pos")
buildPos_wMap <- left_join(buildPos_wMap, enspUnipMap_proc, by = "uniprot_pos")

# Add my ensp->uniprot map for the original sites
InigoFeatMapped_proc <- InigoFeatMapped[, c("ensp_pos", "uniprot_pos")] %>% 
    dplyr::rename("ensp_pos_MeOld" = "ensp_pos")
buildPos_wMap <- left_join(buildPos_wMap, InigoFeatMapped_proc, by = "uniprot_pos")


## Overview of mapping
# How many sites get mapped
table("ensp_pos_InigoNew" = !is.na(buildPos_wMap$ensp_pos_InigoNew), 
    "ensp_pos_MeOld" = !is.na(buildPos_wMap$ensp_pos_MeOld))

# Compare the two mappings
buildPos_wMap$map_agrees <- buildPos_wMap$ensp_pos_InigoNew == buildPos_wMap$ensp_pos_MeOld
## Inspect those that disagree
subset(buildPos_wMap, map_agrees == F) %>% View

## Assign mapping
# Use my mappings, and when they are missing use Inigo's
buildPos_wMap$ensp_pos_final <- ifelse(
    is.na(buildPos_wMap$ensp_pos_MeOld),
    buildPos_wMap$ensp_pos_InigoNew,
    buildPos_wMap$ensp_pos_MeOld
)

## Export
PRIDE_uniprotToEnsp <- buildPos_wMap[, c("uniprot_pos", "ensp_pos_final")] %>% 
  dplyr::rename("ensp_pos" = "ensp_pos_final")
save(PRIDE_uniprotToEnsp, file = "output/data/PRIDE_uniprotToEnsp__3.RData")
```





# Overview and broad characterisation of ubi sites in the build
```{r}
#Define build data to use. For now I use the build without subsetting for Inigo's data
buildDat <- buildPos %>%
  mutate(PTM_FLR_category = factor(PTM_FLR_category, levels = c("Gold", "Silver", "Bronze")))
```

## Number of sites and proteins
```{r}
#Number of sites
numSummSites <- buildDat %>% group_by(PTM_FLR_category) %>%
      summarise(num = length(unique(uniprot_site)),
                group = "Sites")
#Number of proteins. When a protein has sites of different quality, I pick the best
bestLevel <- function(x){sort(x, decreasing = F)[1]}
numSummProt <- buildDat[, c("PTM_FLR_category", "uniprot")] %>% 
      group_by(uniprot) %>% summarise(PTM_FLR_category = bestLevel(PTM_FLR_category), .groups = "drop") %>%
      group_by(PTM_FLR_category) %>% summarise(num = length(unique(uniprot)), group = "Proteins") 
#Combine
numSumm <- rbind(numSummSites, numSummProt) %>%
      mutate(group = factor(group, levels = c("Sites", "Proteins")))

#Plot
ggplot(numSumm, aes(x = group, y = num, fill = PTM_FLR_category)) +
  geom_col(position = "stack", width = 0.8) +
  geom_text(data = numSumm %>% group_by(group) %>% 
            summarise(total = sum(num), totalLab = scales::comma(sum(num))),
            aes(x = group, y = total, label = totalLab),
            vjust = -0.5, size = 1.875*7/5,
            inherit.aes = F) +
  comfy_theme(include_xaxis = F, rotate_x_text = T) +
  theme(axis.title.x = element_blank()) +
  labs(y = "# ubiquitination events") +
  lims(y = c(0, 120000)) +
  scale_fill_manual(values = prideSetColours, name = "Confidence")
ggsave_pdfpng(file = "output/images/overview/PRIDE/bplot_numSitesProts",
      width = 2.25, height = 2)
```

## Number of sites per protein
```{r}
numSitesPerprot <- buildDat %>% group_by(uniprot) %>% summarise(numSites = n()) %>%
                    mutate(group = "all")
#Summarise with bins
numSitesPerprotSumm <- binAndSummarise(data = numSitesPerprot, bin_breaks = c(1, 2, 3, 4, 5, 6, 11, 51, Inf),
                            groupCol = "group", valCol = "numSites",
                            breakType = "integer") %>%
              mutate(BinLab = Bin)
levels(numSitesPerprotSumm$BinLab) <- map(levels(numSitesPerprotSumm$BinLab), function(x){
  xSplit <- strsplit(x, "-")[[1]]
  if (length(xSplit) > 1){
    if (xSplit[1] == xSplit[2]){
      x <- xSplit[1]
    }
  }
  return(x)
}) %>% unlist
##Plot
n <- length(levels(numSitesPerprotSumm$Bin))
palette <- brewer.pal(9, "Blues")[(9 - n):9]
ggplot(numSitesPerprotSumm, aes(x = BinLab, 
                                fill = BinLab,
                                y = n)) +
      #geom_col(fill = prideSetColours["Gold"]) +
      geom_col(colour = "#3a3a3a", linewidth = 0.235/2,
      width = 0.85) +
      comfy_theme(include_xaxis = F, include_legend = F) +
      scale_fill_manual(values = palette) +
      labs(x = "# ubi-sites per protein", y = "# proteins")
ggsave_pdfpng(file = "output/images/overview/PRIDE/bplot_numSitesPerProt",
      width = 2.65, height = 2)
```



## Protein abundance
Set up protein abundance data
```{r}
#Map in paxDB data
load(paste0(home_directory, "data/biol_databases/PaxDB/20250320/9606/9606-WHOLE_ORGANISM-integrated_mappedFID.RData"))

#Get uniprot id
uniprotHsIso$unip_id <- strsplit(rownames(uniprotHsIso), "|", fixed = T) %>% map(~.[3]) %>% unlist 
buildDat <- left_join(buildDat, uniprotHsIso[, c("uniprotkb", "unip_id")] %>% dplyr::rename("uniprot" = "uniprotkb"))

#Map abundance based on this
buildDat <- left_join(buildDat, paxDBunipid[, c("abundance", "unip_id")])

#How many get abundance?
table(!is.na(buildDat$abundance))
buildDat[, c("uniprot", "abundance")] %>% distinct %>% .[, "abundance"] %>% is.na %>% table
```


Explore protein abundance for ubn
```{r}
#Ubiquitinated proteins vs all proteins
paxDBunipid$hasUbi <- paxDBunipid$unip_id %in% buildDat$unip_id 
protAbundAll <- paxDBunipid %>%  
      mutate(group = ifelse(hasUbi, "ubi", "all"))
View(protAbundAll)
##Plot

###Histogram
ggplot(protAbundAll, aes(x = abundance, fill = group)) +
      geom_histogram(colour = "#3a3a3a", linewidth = 0.235/2) +
      comfy_theme() +
      scale_x_log10() +
      scale_fill_manual(values =  setNames(c(PSPcolours["notReg"], "#2172b6"), c("all", "ubi")),
                        labels = c("All", "Ubiquitinated"),
                        name = "Proteins") + 
      labs(x = "Protein abundance (ppm)", y = "# proteins")
ggsave_pdfpng(file = "output/images/overview/PRIDE/hist_abund_ubiVsAll",
          width = 2.75, height = 1.5)
###Ridges
ggplot(protAbundAll, aes(x = abundance, y = group, fill = group)) +
      geom_density_ridges(linewidth = 0.235) +
      comfy_theme() +
      scale_x_log10() +
      scale_fill_manual(values =  setNames(c(prideSetColours["Gold"], PSPcolours["notReg"]), c("ubi", "all"))) +
      labs(x = "Protein abundance (ppm)")
ggsave_pdfpng(file = "output/images/overview/PRIDE/ridge_abund_ubiVsAll",
          width = 2.25, height = 1.5)

#Different confidence scores
ggplot(buildDat, aes(x = abundance, y = PTM_FLR_category, fill = PTM_FLR_category)) +
      geom_density_ridges(linewidth = 0.235) +
      comfy_theme() +
      scale_x_log10(limits = c(1e-03, NA)) +
      scale_fill_manual(values = prideSetColours, name = "Confidence") +
      labs(x = "Protein abundance", y = "Confidence") 
ggsave_pdfpng(file = "output/images/overview/PRIDE/ridge_abund_confidence",
          width = 2.5, height = 1.5)
## Hist
ggplot(buildDat, aes(x = abundance, fill = factor(PTM_FLR_category, levels = rev(levels(PTM_FLR_category))))) +
  geom_histogram(colour = "#3a3a3a", linewidth = 0.235/2) +
  comfy_theme() +
  scale_x_log10(limits = c(1e-03, NA)) +
  scale_fill_manual(values =  prideSetColours,
        labels = c("≥ Bronze", "≥ Silver", "Gold"),
        name = "Confidence") + 
  labs(x = "Protein abundance (ppm)", y = "# ubisites")
ggsave_pdfpng(file = "output/images/overview/PRIDE/hist_abund_confidence",
          width = 2.75, height = 1.5)
```


### Number of sites vs other features
```{r} 
# Set up features from uniprot
uniprotHsIso <- uniprotHsIso %>% 
        mutate(seqLength = str_length(sequence),
             kCount = str_count(sequence, "K"))
numSitesPerprot <- left_join(numSitesPerprot, uniprotHsIso[, c("uniprotkb", "seqLength", "kCount")] %>% 
                                                dplyr::rename("uniprot" = "uniprotkb"), 
                                                by = "uniprot")
numSitesPerprot <- left_join(numSitesPerprot, distinct(buildDat[, c("uniprot", "abundance")]), 
                                                by = "uniprot")

# Plot
prefix <- "output/images/overview/pplot_numSites_vs_"
plot_params <- list(
  list(var = "abundance", yaxis = "Protein abundance (ppm)"),
  list(var = "seqLength", yaxis = "Protein length"),
  list(var = "kCount",    yaxis = "# lysines per protein")
)
plot_list <- list()
for (i in 1:length(plot_params)){
  param <- plot_params[[i]]
  var <- param$var
  yaxis <- param$yaxis
  p <- ggplot(numSitesPerprot, aes(y = !!sym(var), x = numSites)) +
         geom_point(shape = 16, alpha = 0.25, size = 0.5, colour = "#28275d") +
         comfy_theme() + scale_x_log10() + scale_y_log10() +
         labs(x = "# ubi-sites per protein", y = yaxis)
  plot_list[[i]] <- p
}
ggarrange(plotlist = plot_list, nrow = 1)
ggsave_pdfpng(file = paste0(prefix, "allVars"), height = 1.6, width = 1.75*3)
```



## GO enrichment
```{r}
# Set up
onts <- list(BP = "BP", MF = "MF", CC = "CC")

# All proteins with ubisites
background_genes <- unique(uniprotHsRevi$uniprotkb) #Use the reviewed proteome without isoforms
DE_genes <- unique(buildDat$uniprot)
GOresList_allProts <- lapply(onts, function(ont) {
  GO_enricher(background_genes = background_genes,
              DE_genes = DE_genes,
              gene_label = "UNIPROT",
              organism_database = org.Hs.eg.db,
              ontology = ont,
              pathway_DE_intersection_threshold = 3,
              alternative = "greater")
})

# Proteins with > X ubisites vs anythign with a ubi-site
background_genes <- unique(buildDat$uniprot) #Use all proteins with ubi-sites
DE_genes <- subset(numSitesPerprot, numSites >= 11)$uniprot %>% unique
GOresList_manySites <- lapply(onts, function(ont) {
  GO_enricher(background_genes = background_genes,
              DE_genes = DE_genes,
              gene_label = "UNIPROT",
              organism_database = org.Hs.eg.db,
              ontology = ont,
              pathway_DE_intersection_threshold = 3,
              alternative = "greater")
})
View(head(GOresList_manySites$BP, 20))


# Plot
for (ont in names(GOresList_manySites)) {
  enrichment_barplot_pval(
    data = GOresList_manySites[[ont]] %>% arrange(adj_pval) %>% head(15),
    pval_col = "adj_pval",
    colour_col = "ES_log2",
    colours = brewer.pal(9, "RdBu"),
    order_col = "adj_pval",
    order_decreasing = TRUE,
    sig_line = FALSE,
    xlab = "-log10 adj. p-value",
    legend_title = "Fold enrichment (log2)"
  )  + 
  ggtitle("Proteins with > 10 ubi-sites")
  ggsave_pdfpng(file = paste0("output/images/overview/enrichment/bplot_GOenr_geq11_", ont),
                width = 4.25, height = 2.4)
}
```


## Comparison to PhosphositePlus
### Set up numbers
```{r}
# PhosphositePlus
## Load all PSP ubi sites and prep
PSPDir <- paste(home_directory, "data/biol_databases/phosphosite_plus/DATE", sep = "")
f <- "Ubiquitination_site_dataset"
PSP_ubn_raw <- read_delim(file.path(PSPDir, f)) %>%
  as.data.frame
PSP_ubn_hs <- subset(PSP_ubn_raw, ORGANISM == "human") %>% 
              mutate(site = gsub("-ub$", "", MOD_RSD)) %>% 
              mutate(uniprot_site = paste(ACC_ID, site, sep = "_"))
## Evidence of identification
PSP_ubn_hs$MS_LIT[which(is.na(PSP_ubn_hs$MS_LIT))] <- 0
PSP_ubn_hs$MS_CST[which(is.na(PSP_ubn_hs$MS_CST))] <- 0
PSP_ubn_hs <- mutate(PSP_ubn_hs, MS_total = MS_LIT + MS_CST) # Sum LIT and CST evidence

# PRIDE build
PXDIds <- colnames(buildPos)[grep("^PXD", colnames(buildPos))] %>% gsub("_.+", "", .) %>% unique
FLR_cols <- paste(PXDIds, "_FLR", sep = "")
# Count number of datasets with FLR < 5%
buildPos$FLR5count <- rowSums(buildPos[, FLR_cols] < 0.05, na.rm = T)

# Combine
PSP_PRIDE_MS2nums <- map_dfr(list("PSP" = PSP_ubn_hs[, c("uniprot_site", "MS_total")],
                                  "PRIDE" = buildPos[, c("uniprot_site", "FLR5count")]),
                                  function(x){
                                    colnames(x) <- c("uniprot_site", "MS2_count")
                                    return(x)
                                  }, .id = "source") %>% 
                      subset(MS2_count > 0)
## Bin summary
PSP_PRIDE_MS2nums_binSumm <- PSP_PRIDE_MS2nums %>% 
                binAndSummarise(., bin_breaks = c(0, 1, 2, 3, 4, 5, 6, Inf),
                            groupCol = "source", valCol = "MS2_count",
                            breakType = "integer")
levels(PSP_PRIDE_MS2nums_binSumm$Bin) <- gsub("^([0-9]+)-\\1$", "\\1", levels(PSP_PRIDE_MS2nums_binSumm$Bin))
PSP_PRIDE_MS2nums_binSumm$sourceLabel <- factor(PSP_PRIDE_MS2nums_binSumm$source,
  levels = c("PRIDE", "PSP"),
  labels = c("PRIDE (11 datasets)", "PSP (> 150 datasets)")
)
### Plot
n <- length(levels(PSP_PRIDE_MS2nums_binSumm$Bin))
#palette <- rev(brewer.pal(9, "Blues")[c(2, 4, 6, 8)])
#palette <- rev(brewer.pal(n, "Blues"))
palette <- rev(brewer.pal(9, "Blues")[c(1, 3, 4, 6, 8, 9)])
ggplot(PSP_PRIDE_MS2nums_binSumm, aes(x = sourceLabel, y = n, fill = factor(Bin, levels = rev(levels(Bin))))) +
  geom_col(position = "stack", colour = "#3a3a3a", linewidth = 0.235/2) +
  scale_fill_manual(values = palette) +
  comfy_theme(include_xaxis = F, rotate_x_text = F) + 
  theme(axis.title.x = element_blank()) +
  labs(x = "Source", y = "# human ubi-sites", fill = "# datasets") + 
  ggtitle("Site identification across MS datasets")
ggsave_pdfpng(file = "output/images/overview/PRIDE/vsPSP/bplot_numMS2", width = 2.25, height = 1.75)
```

### Overlap with PSP 
```{r}
# Make a list of dataframes containing sites
sitelst_usVsPsp <- PSP_PRIDE_MS2nums %>% 
  mutate(source = factor(source, levels = c("PRIDE", "PSP"), labels = c("This study", "PhosphositePlus"))) %>% 
  split(.$source)

# Plot venn diagram
map(sitelst_usVsPsp, ~unique(.$uniprot_site)) %>% 
  ggVennDiagram_nice(.)  +
  ggtitle("Overlap of ubi-sites")
ggsave_pdfpng(file = "output/images/overview/PRIDE/vsPSP/venn_numSites_usVsPSP",
              width = 3.5, height = 3.5)

# Only sites supported by >= 2 databases
map(sitelst_usVsPsp, function(x){
  unique(subset(x, MS2_count > 1)$uniprot_site)
}) %>% 
  ggVennDiagram_nice(.)  +
  ggtitle("Overlap of ubi-sites")
ggsave_pdfpng(file = "output/images/overview/PRIDE/vsPSP/venn_numSites_usVsPSP_MSgeq2",
              width = 3.5, height = 3.5)
```


## Overlap with Wen 2025 Nature methods
I compare to sites identified by re-searching mass spec data in Wen 2025 (doi.org/10.1038/s41592-025-02797-x)
```{r}
# Prep data
PTMatlasComb <- read_csv(paste0(home_directory, "data/supplementary_tables/Wen_NatMethods_2025/PTMatlasComb.csv"))
PTMatlasUbi <- subset(PTMatlasComb, PTM == "ubiquitination") %>% 
  mutate(uniprot_site = paste0(protein, "_", aa, pos))

# Get list of sites
sitelst_usVsPspVsptmatlas <- sitelst_usVsPsp
sitelst_usVsPspVsptmatlas$PTMAtlas <- PTMatlasUbi

# Plot venn diagram
siteLst <- map(sitelst_usVsPspVsptmatlas, ~unique(.$uniprot_site))
map(siteLst) %>% 
  ggVennDiagram_nice(.)  +
  ggtitle("Overlap of ubi-sites")
ggsave_pdfpng(file = "output/images/overview/PRIDE/vsPSP/venn_numSites_usVsPSPVsPTMAtlas",
              width = 4.5, height = 4.5)

# Individual numbers
map(siteLst, length)
length(intersect(siteLst$`This study`, siteLst$PhosphositePlus))
length(intersect(siteLst$`This study`, siteLst$PTMAtlas))

# Is it our bronze sites that are unique?
ourSitesCat <- buildPos %>% split(.$PTM_FLR_category) %>% map(~unique(.$uniprot_site))
map(ourSitesCat, length)
map(ourSitesCat, ~length(setdiff(., siteLst$PTMAtlas)))
map(ourSitesCat, ~length(setdiff(., c(siteLst$PTMAtlas, siteLst$PhosphositePlus))))
```

What PRIDE datasets do they look at vs us?
```{r}
# Load
PTMatlasDatasets <- read_xlsx(paste0(home_directory, "data/supplementary_tables/Wen_NatMethods_2025/Table_S1.xlsx"), 
                              sheet = "S1", skip = 2)
ourDatasets <- read_xlsx("data/supp_tables/SuppTable_PRIDE_datasets.xlsx", sheet = "table")
# Extract datasets
datasetsLst <- list()
datasetsLst$`This study` <- ourDatasets$Datasets %>% gsub("-.+$", "", .) %>% unique
#datasetsLst$`This study` <- ourDatasets$`Pubmed ID` %>% unique
datasetsLst$PTMAtlas <- subset(PTMatlasDatasets, PTM == "Ubiquitination")$Dataset
#datasetsLst$PTMAtlas <- subset(PTMatlasDatasets, PTM == "Ubiquitination")$PMID
# Overlap 
ggVennDiagram_nice(datasetsLst)
datasetsTab <- data.frame("Dataset" = unique(unlist(datasetsLst)))
for (name in names(datasetsLst)){
  datasetsTab[, name] <- datasetsTab$Dataset %in% datasetsLst[[name]]
}
View(datasetsTab)
```