# Background
Here I annotate the PRIDE human ubi-sites with features that I later use for functional score training


## Checklist - things that need to be run manually
- Get VEPs and interface residues - scripts/11_VEP.py
- Disorder, surface accessibility, and secondary structure - run NetSurfP3.0
- Protein domains, motifs etc - run uniprot API call (this script)




# Initialise
```{r}
library(tidyverse)
library(purrr)
library(ggplot2)

home_directory <- ""
source(paste0(home_directory, "scripts/essential_visualisations_JvG.R"))
source("scripts/01_general.R")
source("scripts/03_visualisations.R")

#Load data
load("data/PSP/PSP_reg_ubCorr__3.RData")

# Set up a log to track missing data in various features
featLog = list()
```

Load in relevant proteomes
```{r}
##Function: read in a fasta file of protein sequences and convert them into a data frame
library(seqinr)
read_fastaseq <- function(directory){
  seqs_raw <- read.fasta(directory)
  seqs <- data.frame("sequence" = map(seqs_raw, ~paste(., collapse = "")) %>% unlist)
  seqs$sequence <- toupper(seqs$sequence)
  seqs$uniprotkb <- strsplit(rownames(seqs), "\\|") %>% map(~.[2]) %>% unlist
  return(seqs)
}
##Load proteomes with cache
prtmCacheDir <- "output/data/proteomes/"
prtmRealDir <-  paste0(home_directory, "data/biol_databases/uniprot/")
# Iso
cacheF <- paste0(prtmCacheDir, "uniprotkb_proteome_UP000005640_2025_02_19", ".RData")
if (file.exists(cacheF)){
  load(cacheF)
} else {
  uniprotHsIso <- read_fastaseq(paste(prtmRealDir, 
    "uniprotkb_proteome_UP000005640_2025_02_19", ".fasta", sep = ""))
  save(uniprotHsIso, file = cacheF)
}
# Iso revi
cacheF <- paste0(prtmCacheDir, "uniprotkb_proteome_UP000005640_AND_revi_2025_02_19", ".RData")
if (file.exists(cacheF)){
  load(cacheF)
} else {
  uniprotHsIsoRevi <- read_fastaseq(paste(prtmRealDir, 
    "uniprotkb_proteome_UP000005640_AND_revi_2025_02_19", ".fasta", sep = ""))
  save(uniprotHsIsoRevi, file = cacheF)
}
```


# Load and prep the PRIDE build
```{r}
processBuild <- function(data){
  dataProc <- data %>%
    mutate(uniprot = strsplit(Protein, "|", fixed = T) %>% map(~.[2]) %>% unlist,
           unipGene = strsplit(Protein, "|", fixed = T) %>% map(~.[3]) %>% unlist) %>%
    mutate(site = paste0(PTM_residue, Protein_pos)) %>%
    mutate(uniprot_site = paste(uniprot, site, sep = "_"),
           uniprot_pos = paste(uniprot, Protein_pos, sep = "_")) %>%
    #Remove A decoys
    subset(PTM_residue != "A") %>%
    as.data.frame
  return(dataProc)
}
#buildPos - No redundant sites. This is what I use
buildPos <- read_csv("data/PRIDE/DATE/G2S1B_0.05_protein_pos_single_prot_mapping.csv") %>% processBuild
#buildPosAllprots
buildPosAllprots <- read_csv("data/PRIDE/DATE/G2S1B_0.05_protein_pos_all_prot_mapping.csv") %>% processBuild

# Define data to use
ubiFeatRaw <- buildPos %>% as.data.frame %>% 
  dplyr::rename(pos = Protein_pos)

#Set up a subsetted proteome
uniprotHsIsoSub <- subset(uniprotHsIso, uniprotkb %in% ubiFeatRaw$uniprot)
```

## Get sequence windows from uniprot
```{r}
# Are all uniprots present?
table(ubiFeatRaw$uniprot %in% uniprotHsIsoSub$uniprotkb)
setdiff(ubiFeatRaw$uniprot, uniprotHsIsoSub$uniprotkb)
## Filter out those that are not present (only 2 proteins and 22 sites)
ubiFeatRaw <- subset(ubiFeatRaw, uniprot %in% uniprotHsIsoSub$uniprotkb)

# Extract sequence windows
## Set up caching: load existing cache or initialise a new one to store computed results for unipAA and unipSeqWind
cache_file <- "output/data/unipSeqWindCache.rds"
if (file.exists(cache_file)) {
  resCache <- readRDS(cache_file)
} else {
  resCache <- list()
}
## Process each unique protein and cache the results per position
uniqProteins <- unique(ubiFeatRaw$uniprot)
for (up in uniqProteins) {
  idx <- which(ubiFeatRaw$uniprot == up)
  # If not already cached, compute the results for all positions of this protein
  if (!(up %in% names(resCache))) {
    seq_raw <- uniprotHsIsoSub$sequence[which(uniprotHsIsoSub$uniprotkb == up)]
    padded_seq <- paste(c(rep("_", 7),
                          seq_raw,
                          rep("_", 7)), collapse = "")
    positions <- ubiFeatRaw$pos[idx]
    unipAA_vals <- sapply(positions, function(pos) {
      if (pos <= str_length(padded_seq) - 14) {
         substr(padded_seq, pos + 7, pos + 7)
      } else { 
         NA 
      }
    })
    unipSeqWind_vals <- sapply(positions, function(pos) {
      if (pos <= str_length(padded_seq) - 14) {
         substr(padded_seq, pos, pos + 14)
      } else { 
         NA 
      }
    })
    resCache[[up]] <- data.frame(pos = positions,
                                 unipAA = unipAA_vals,
                                 unipSeqWind = unipSeqWind_vals,
                                 stringsAsFactors = FALSE)
  }
}

# Use cache results all at once
resCacheDf <- map_dfr(resCache, function(x){x}, .id = "uniprot") %>% 
  mutate(uniprot_pos = paste(uniprot, pos, sep = "_"))
ubiFeatProc <- left_join(ubiFeatRaw, resCacheDf[, c("uniprot_pos", "unipAA", "unipSeqWind")], by = "uniprot_pos")

# Save/Update the cache for future runs
saveRDS(resCache, file = cache_file)

# Check sequences
## Anything missing?
table(is.na(ubiFeatProc$unipAA))
## Anything not lysine?
table(ubiFeatProc$unipAA)
subset(ubiFeatProc, unipAA != "K") %>% View

# Filter out missing and not lysine
ubiFeatProc <- subset(ubiFeatProc, !is.na(unipAA) & unipAA == "K")
nrow(ubiFeatProc)
nrow(buildPos)
```



# Map in PhosphositePlus annotations
```{r}
PSP_reg_ubSub <- subset(PSP_reg_ubCorr, uniprot_site %in% ubiFeatProc$uniprot_site)

#Big numbers
print("# of PSP sites")
print(nrow(PSP_reg_ubCorr))
print("# of PSP sites in Inigo's sites")
print(nrow(PSP_reg_ubSub))

#Different functions
PSP_reg_ub_FUNC <- split_ON_FUNCTION(PSP_reg_ubSub) 
##Number of unique sites per ON_FUNCTION
onFuncSum <- PSP_reg_ub_FUNC %>% split(.$ON_FUNCTION_strp) %>% map(~unique(.$uniprot_site)) %>% map_dbl(~length(.)) %>% as.data.frame
colnames(onFuncSum)[1] <- "numSites"
onFuncSum$FUNCTION <- rownames(onFuncSum)
##Plot
ggplot(subset(onFuncSum, numSites > 10), 
       aes(x = reorder(FUNCTION, numSites, decreasing = TRUE), 
           y = numSites)) +
  geom_bar(stat = "identity") +
  comfy_theme(include_xaxis = F, rotate_x_text = T) +
  xlab("Function") +
  ylab("# ubn sites") +
  ggtitle("Regulatory ubn sites")
ggsave_pdfpng(file = "output/images/overview/bplot_InigoOverlap_ubnFuncs",
              width = 3, height = 2.25)

#Degradation vs other functions
##Flag in inigo's data
protDegSites <- subset(PSP_reg_ub_FUNC, ON_FUNCTION_strp == "protein degradation")$uniprot_site
otherFuncSites <- subset(PSP_reg_ub_FUNC, ON_FUNCTION_strp != "protein degradation")$uniprot_site
ubiFeatProc$PSP_reg <- ubiFeatProc$uniprot_site %in% PSP_reg_ubSub$uniprot_site
ubiFeatProc$PSP_reg_ProtDeg <- ubiFeatProc$uniprot_site %in% protDegSites
ubiFeatProc$PSP_reg_otherFunc <- ubiFeatProc$uniprot_site %in% otherFuncSites
ubiFeatProc$PSP_reg_notProtDeg <- ubiFeatProc$uniprot_site %in% setdiff(otherFuncSites, protDegSites)
ubiFeatProc$PSP_reg_ProtDegOnly <- ubiFeatProc$uniprot_site %in% setdiff(protDegSites, otherFuncSites)

cols <- colnames(ubiFeatProc) %>% .[grep("^PSP_", .)]
for (col in cols){
  print(col)
  print(sum(ubiFeatProc[, col]))
}
```   

# Prepare and inspect PRIDE confidence
```{r}
ubiFeatProc$PRIDE_set <- factor(ubiFeatProc$PTM_FLR_category, levels = c("Bronze", "Silver", "Gold"))
table(ubiFeatProc$PRIDE_set)

#Inspect for PSP regulatory sites
PRIDESetSumm <- catBoolSummariser(data = ubiFeatProc, catCol = "PRIDE_set",
                  boolCols = c("PSP_reg_ProtDeg", "PSP_reg_otherFunc", "PSP_reg"))
PRIDESetSumm %>% pivot_longer(cols = c("PSP_reg_ProtDeg_pct", "PSP_reg_otherFunc_pct"), names_to = "type", values_to = "perc") %>%
  ggplot(., aes(x = PRIDE_set, y = perc, fill = type)) +
  geom_col(position = position_dodge2()) +
  comfy_theme() +
  labs(x = "PRIDE confidence", y = "% of site group")
ggsave_pdfpng(file = "output/images/features/PRIDE/bplot_boxes_PSPreg",
              width = 3, height = 2)
```


# Map in conservation boxes
```{r}
load("data/conservation/conservationBoxesDfUniq__1.RData")
conservationBoxesDfUniq_filt <- conservationBoxesDfUniq %>% 
  .[, c("uniprot_site", "ensp_pos", "consBox_LysCont", "consBox_v1")] %>% 
  dplyr::rename("Cons_ensp_pos" = "ensp_pos" )
# Add to features
ubiFeatProc <- left_join(ubiFeatProc, conservationBoxesDfUniq_filt, by = "uniprot_site")
# Overview
table(!is.na(ubiFeatProc$consBox_LysCont))
table(ubiFeatProc$consBox_LysCont)
```


# Protein abundance from PaxDB
```{r}
#Map in paxDB data
load(paste0(home_directory, "data/biol_databases/PaxDB/20250320/9606/9606-WHOLE_ORGANISM-integrated_mappedFID.RData"))

#Get uniprot id
uniprotHsIso$unip_id <- strsplit(rownames(uniprotHsIso), "|", fixed = T) %>% map(~.[3]) %>% unlist 
ubiFeatProc <- left_join(ubiFeatProc, uniprotHsIso[, c("uniprotkb", "unip_id")] %>% dplyr::rename("uniprot" = "uniprotkb"))

#Map abundance based on this
ubiFeatProc <- left_join(ubiFeatProc, paxDBunipid[, c("abundance", "unip_id")])

# Overview
table(!is.na(ubiFeatProc$abundance))
```




# Get VEPs and interface sites
```{r}
#Set up mutations
muts <- list(
  "R" = paste(ubiFeatProc$uniprot_site, "R", sep = "") %>% gsub("_", "/", .),
  "A" = paste(ubiFeatProc$uniprot_site, "A", sep = "") %>% gsub("_", "/", .)
)
mutsDf <- data.frame("variant_id" = unlist(muts))
write.csv(mutsDf, file = "output/data/VEP/ubnVariants.csv")

#HERE YOU NEED TO RUN ubnVariants.csv through the VEP script: scripts/11_VEP.py

# Load in results
mutsDfVEP_merged <- read_csv("output/data/VEP/ubnVariantsVEP_merged.csv") %>% as.data.frame %>% .[, -1]

# Flag what is missing
VEPmissingSites <- setdiff(ubiFeatProc$uniprot_site, mutsDfVEP_merged$uniprot_site)
featLog[["VEP"]] <- list(
  "status" = ifelse(length(VEPmissingSites) == 0, "All sites covered", paste0(length(VEPmissingSites), " sites missing")),
  "missingSites" = VEPmissingSites
)

# Add to data
ubiFeatProc <- left_join(ubiFeatProc, mutsDfVEP_merged, by = "uniprot_site")

# Overview of coverage
colSums(is.na(ubiFeatProc[, c("mutR_am_class_all", "interface_label", "mutR_ESM1b_LLR")]))
```


# Protein length and related
```{r}
#Protein length
uniprotHsIsoSub$length <- str_length(uniprotHsIsoSub$sequence)
rownames(uniprotHsIsoSub) <- uniprotHsIsoSub$uniprotkb
ubiFeatProc$proteinLength <- factor(ubiFeatProc$uniprot)
levels(ubiFeatProc$proteinLength) <- uniprotHsIsoSub[levels(ubiFeatProc$proteinLength), "length"]
ubiFeatProc$proteinLength <- as.numeric(as.character(ubiFeatProc$proteinLength))
#Distance from C and N termini
ubiFeatProc <- mutate(ubiFeatProc, 
                        distN = pos,
                        distC = proteinLength - pos + 1) %>%
  mutate(distNFrac = distN / proteinLength,
         distCFrac = distC / proteinLength)

#Inspect for PSP sites
combDat <- makeCombDat(ubiFeatProc)
vars <- c("proteinLength", "distN", "distC")
for (var in vars){
  plt <- ViolBoxplot(data = combDat, vars = list(x = "group", y = var, colour = "group")) +
    scale_y_log10()
  plot(plt)
  ggsave_pdfpng(file = paste0("output/images/features/structure/viol_", var, "_PSPreg"),
              width = 3.5, height = 2)
}

```



# NetSurfP3.0: Disorder, surface accessibility, and secondary structure
Here I export sequences to be run through NetSurfP3.0. I then load in the results 
Many steps of this are slow (e.g. reading in many small files), so I use caches 
to speed it up
### Prepare and filter sequences
```{r}
# Get sequences to run
library(Biostrings)
seqDat <- subset(uniprotHsIsoSub, length <= 5000)
seqDat$sequence <- gsub("U", "C", seqDat$sequence) #Change selenocysteine to cysteine

# Get sites that should be run
preNSPsites <- ubiFeatProc[, c("uniprot", "uniprot_site", "uniprot_pos", "pos")] %>% 
  subset(uniprot %in% seqDat$uniprotkb)
```

### Load in results (including cache) and export missing sequences
```{r}
# Load in cache
cacheF <- "output/data/structure/NSPdat_forUbn_cache.RData"
if (file.exists(cacheF)){
  load(cacheF)  # Data is called NSPdat
} else {
  NSPdat <- NULL
}
# Check sites that are not in the cache
preNSPsitesMissing <- subset(preNSPsites, uniprot_pos %in% NSPdat$uniprot_pos == F)

# Make map to existing result files
outputBatchDir <- "output/data/structure"
outputBatches <- dir(outputBatchDir) %>% .[grep("^batch\\d+$", .)]
unipDirDf <- map_dfr(outputBatches, function(f){
  batchDir <- file.path(outputBatchDir, f, "01")
  unipFiles <- dir(batchDir) %>% .[grep("^\\d+_", .)]
  unipFilesFull <- paste0(outputBatchDir, "/", f, "/01/", unipFiles, "/", unipFiles, ".csv")
  unipFilesDf <- data.frame("uniprot" = gsub("^\\d+_", "", unipFiles), "fDir" = unipFilesFull)
}, .id = "batch")
## Flag uniprots that are missing - need to be rerun
## Do this based on the results map, not cache
missingUnips <- unique(setdiff(preNSPsitesMissing$uniprot, unipDirDf$uniprot))

# Export sequences for uniprots that need to be rerun, in batches
## Get existing batches
#batchOutDir <- "output/data/structure"
#existingBatches <- dir(batchOutDir) %>% .[grep("UbnUnipSeqs_batch\\d+\\.fasta", .)] %>% 
#  gsub("UbnUnipSeqs_batch", "", .) %>% gsub(".fasta", "", ., fixed = T) %>% as.numeric
existingBatches <- outputBatches %>% gsub("^batch", "", .) %>% as.numeric
## Make new batches, max size 4900
if (length(missingUnips) > 0){
  newBatches <- split(missingUnips, ceiling(seq_along(missingUnips) / 4900))
  newBatchIDs <- 1:length(newBatches) + max(existingBatches)
  for (i in 1:length(newBatchIDs)){
    seqDatSub <- subset(seqDat, uniprotkb %in% newBatches[[i]])
    fasta_seqs <- BStringSet(seqDatSub$sequence)
    names(fasta_seqs) <- seqDatSub$uniprotkb
    writeXStringSet(fasta_seqs, paste0("output/data/structure/UbnUnipSeqs_batch", newBatchIDs[i],  ".fasta"))
  }
}

##### HERE YOU NEED TO RUN NSP3 ON ANY NEW BATCHES #####
# conda activate nsp3
# python nsp3.py -m models/nsp3.pth -i (input file) -o (output file)
# Load in results for missing sites
## Subet for missing sites where the uniprots have results
preNSPsitesMissing_wRes <- subset(preNSPsitesMissing, uniprot %in% unipDirDf$uniprot)
## Read in results for new sites by looping over uniprots
if (nrow(preNSPsitesMissing_wRes) > 0){
  NSPdatNewLst <- list()
  for (unip in unique(preNSPsitesMissing_wRes$uniprot)){
    ## Get posns
    posns <- subset(preNSPsitesMissing_wRes, uniprot == unip)$pos
    ## Load in NSP data
    resDir <- unipDirDf$fDir[which(unipDirDf$uniprot == unip)[1]]
    dat <- read_csv(resDir)
    datSub <- subset(dat, n %in% posns) %>% as.data.frame
    NSPdatNewLst[[unip]] <- datSub
  }
  NSPdatNew <- purrr::reduce(NSPdatNewLst, rbind)
  NSPdatNew <- NSPdatNew %>% 
    mutate(uniprot = gsub("^>", "", id)) %>%
    mutate(uniprot_pos = paste(uniprot, n, sep = "_")) %>%
    dplyr::select(-c("id", "seq", "n"))
  inds <- which(colnames(NSPdatNew) != "uniprot_pos")
  colnames(NSPdatNew)[inds] <- paste0("NSP_", colnames(NSPdatNew)[inds])
} else {
  NSPdatNew <- NULL
}
## Process and add to cache
NSPdat <- rbind(NSPdat, NSPdatNew) %>% distinct
save(NSPdat, file = cacheF)
```

### Add to table and check what is still missing
```{r}
# Add to data
ubiFeatProc <- left_join(ubiFeatProc, NSPdat, by = "uniprot_pos")

# Flag anything still missing
NSPmissingSites <- setdiff(preNSPsites$uniprot_pos, NSPdat$uniprot_pos)
featLog[["NSP"]] <- list(
  "status" = ifelse(length(NSPmissingSites) == 0, "All sites covered", paste0(length(NSPmissingSites), " sites missing")),
  "missingSites" = NSPmissingSites
)
```


### Explore results
```{r}
#Overview
table(!is.na(ubiFeatProc$NSP_asa))

#Inspect for PSP sites
combDat <- makeCombDat(ubiFeatProc)
vars <- c("NSP_rsa", "NSP_disorder", "NSP_p[q3_H]", "NSP_p[q3_E]", "NSP_p[q3_C]")
for (var in vars){
  plt <- ViolBoxplot(data = combDat, vars = list(x = "group", y = var, colour = "group")) +
    scale_y_log10() +
    ggtitle(var)
  plot(plt)
  ggsave_pdfpng(file = paste0("output/images/features/structure/viol_", var, "_PSPreg"),
              width = 3.5, height = 2)
}
```



# Protein domains
## Set up input
```{r}
# Set up input uniprots
preDomainUnips <- uniprotHsIsoSub$uniprotkb

# Load cache
cacheF <- "output/data/structure/domainDat_cache.RData"
if (file.exists(cacheF)){
  load(cacheF)  # Data is called domainDatRaw
} else {
  domainDatRaw <- NULL
}
# Define missing uniprots
preDomainUnipsMissing <- setdiff(preDomainUnips, rownames(domainDatRaw))
```

## Run domain extraction on missing results 
```{r}
library(UniprotR)
if (length(preDomainUnipsMissing) > 0){
  # Make batches of <= 100 uniprots, to help with internet connection
  v <- preDomainUnipsMissing
  batch_size <- 100
  nBatches <- ceiling(length(v) / batch_size)
  batches <- split(v, ceiling(seq_along(v) / batch_size))
  lst <- list()
  i <- 1
  # Run uniprotR domain extraction on batches
  while (i <= nBatches){
    dat <- GetFamily_Domains(batches[[i]])
    if (!is.null(dat)){
      lst[[i]] <- dat
      i <- i + 1
    }
  }
  domainDatRawNew <- purrr::reduce(lst, rbind)
  domainDatRawNew <- domainDatRawNew[which(rownames(domainDatRawNew) %in% rownames(domainDatRaw) == F), ]
} else {
  domainDatRawNew <- NULL
}

# Update cache
domainDatRaw <- rbind(domainDatRaw, domainDatRawNew)
save(domainDatRaw, file = cacheF)
```

## Extract results
```{r}
#Process domain annotations + similar
domainProcessArgs <- list(
  c("Domain..FT.", "DOMAIN", "domain"),
  c("Compositional.bias", "COMPBIAS", "compbias"),
  c("Motif", "MOTIF", "motif"),
  c("Region", "REGION", "region"),
  c("Repeat", "REPEAT", "repeat"),
  c("Zinc.finger", "ZN_FING", "znFing")
)
for (x in domainProcessArgs){
  domainDat <- filter(domainDatRaw, !is.na(!!sym(x[1]))) %>% as.data.frame
  #Process annotations
  annot <- domainDat[, x[1]]
  annot <- gsub(paste0("^", x[2], " "), "", annot)
  posns <- strsplit(annot, paste0("; ", x[2], " ")) %>% map(function(vec){
    vec <- gsub(";.+$", "", vec)
    vec <- gsub(">", "", vec) # Sometimes this in the entry e.g. A0A0B4J1U4: 18..>118; /note=Ig-like; /evidence=ECO:0000255|PROSITE-ProRule:PRU00114
    vec <- vec[!grepl("-", vec)] #Remove matches that are only for isoform uniprots
    vec <- vec[!grepl("?", vec, fixed = T)] #Remove matches that have ?
    if (length(vec) == 0){
      return(NULL)
    } else {
      vecSplt <- strsplit(vec, "..", fixed = T)
      sequence <- map(vecSplt, function(y){
        if (length(y) == 1){
          return(as.numeric(y))
        } else {
          return(y[1]:y[2])
        }
      }) %>% unlist
      return(sequence)
    }
  })
  names(posns) <- rownames(domainDat)
  posns <- posns[map(posns, ~!is.null(.)) %>% unlist]
  
  #Identify ubn sites in annotated regions
  ubnSitesLst <- ubiFeatProc %>% split(.$uniprot) %>% map(~.$pos)
  sitesInDomains <- map2(posns, names(posns), function(domain, unip){
    sites <- ubnSitesLst[[unip]]
    int <- intersect(sites, domain)
    if (length(int) == 0){
      return(NULL)
    } else {
      return(paste(unip, int, sep = "_"))
    }
  }) %>% unlist
  ubiFeatProc[, paste0("UNIPROT_in_", x[3])] <- ubiFeatProc$uniprot_pos %in% sitesInDomains
}

ubiFeatProc %>% .[, grep("^UNIPROT_", colnames(.))] %>% colSums


#Inspect for PSP regulatory sites
UNIPROTCols <- colnames(ubiFeatProc) %>% .[grep("^UNIPROT_", .)]
names(UNIPROTCols) <- UNIPROTCols
UNIPROTSumm <- map(UNIPROTCols, function(col){
  dat <- filter(ubiFeatProc, !is.na(!!sym(col)))
  catBoolSummariser(data = dat, catCol = col,
                  boolCols = c("PSP_reg_ProtDeg", "PSP_reg_otherFunc", "PSP_reg"))
})
map2(UNIPROTSumm, names(UNIPROTSumm), function(summ, col){
  summ %>% pivot_longer(cols = c("PSP_reg_ProtDeg_pct", "PSP_reg_otherFunc_pct"), names_to = "type", values_to = "perc") %>%
    ggplot(., aes(x = !!sym(col), y = perc, fill = type)) +
    geom_col(position = position_dodge2()) +
    comfy_theme(rotate_x_text = T) +
    labs(x = col, y = "% of site group")
  ggsave_pdfpng(file = paste0("output/images/features/structure/bplot_", col, "_PSPreg"),
                width = 3, height = 2)
})
```


# PTMs at the same position
```{r}
#Load PSP data for all modifications
PSPDir <- paste(home_directory, "data/biol_databases/phosphosite_plus/DATE", sep = "")
Fs <- dir(PSPDir) %>% .[grep("_dataset$", .)]
PSPCombDat <- map(Fs, function(f){
  PSP_raw <- read_delim(file.path(PSPDir, f)) %>%
    as.data.frame
  PSP_sub <- subset(PSP_raw, ORGANISM == "human" & ACC_ID %in% ubiFeatProc$uniprot) %>%
    mutate(mod = gsub("^.+-", "", MOD_RSD),
           site = gsub("-.+$", "", MOD_RSD)) %>%
    mutate(pos = as.numeric(gsub("[A-Z]", "", site))) %>%
    mutate(uniprot_pos = paste(ACC_ID, pos, sep = "_"),
           uniprot_site = paste(ACC_ID, site, sep = "_")) %>%
    dplyr::select("ACC_ID", "mod", "site", "pos", "uniprot_pos", "uniprot_site")
  return(PSP_sub)
}) %>% purrr::reduce(rbind)

#Number of other lysine modifications at the same position
PSPCombDat_KnotUb <- subset(PSPCombDat, mod != "ub" & grepl("^K", site))
ubiFeatProc$PSP_numOtherLysMod <- map(ubiFeatProc$uniprot_site, ~length(which(PSPCombDat_KnotUb$uniprot_site == .))) %>% unlist


#Inspect for PSP sites
combDat <- makeCombDat(ubiFeatProc)
vars <- c("PSP_numOtherLysMod")
for (var in vars){
  plt <- ViolBoxplot(data = combDat, vars = list(x = "group", y = var, colour = "group")) +
    #scale_y_log10() +
    ggtitle(var)
  plot(plt)
  ggsave_pdfpng(file = paste0("output/images/features/PTMs/viol_", var, "_PSPreg"),
              width = 3.5, height = 2)
}

ubiFeatProc$PSP_boolOtherLysMod <- ubiFeatProc$PSP_numOtherLysMod > 0
catBoolSummariser(data = ubiFeatProc, catCol = "PSP_boolOtherLysMod",
                  boolCols = c("PSP_reg_ProtDeg", "PSP_reg_otherFunc", "PSP_reg"))
```



# Export
```{r}
save(ubiFeatProc, file = "output/data/features/ubiFeatProc__1.RData")
```












