# Background
Here I prepare data for the hotspot analysis. This involves mapping non-human ubn sites to uniprot identifiers, extracting domains with InterPro API call, and preparing domains for alignment


## How to run the analysis 
- Run up to step 6b in this document. This will map identifiers, extract domains, and prepare them for hotspot analysis
- Move output from output/data/hotspots/domains/domainsPreAlign and output/data/hotspots/ubnSites_forHotspots to computational cluster
- Run scripts/hotspot_scripts/02_runMAFFT_for_hotspots.py to align domain sequences
- Run scripts/hotspot_scripts/03_run_ptm_hotspots.py to run hotspot analysis
- Move output from output/data/hotspots/hotspotOutput.csv back to local
- Run steps 7 up to 7b in this document 
- Move output from output/data/hotspots/domains/representativeStrs/strRep_seqDomain.fasta to the cluster 
- Run scripts/hotspot_scripts/04_repStrsMAFFT.py to align domain sequences with the representative structure 
- Run scripts/hotspot_scripts/05_repStrsHotspots.py to run hotspot analysis with the representative structure
- Move output from output/data/hotspots/domains/representativeStrs back to local 
- Run steps 8 and beyond in this document


# Initialise
```{r}
library(tidyverse)
library(purrr)
library(ggplot2)
library(seqinr)
library(data.table)
library(UniProt.ws)
up <- UniProt.ws()
library(Biostrings)

home_directory <- ""
source(paste0(home_directory, "scripts/essential_visualisations_JvG.R"))
source("scripts/00_objects.R")
source("scripts/01_general.R")
source("scripts/hotspot_scripts/01_hotspotPrepFns.R")
```


Load in relevant proteomes
```{r}
##Function: read in a fasta file of protein sequences and convert them into a data frame
library(seqinr)
read_fastaseq <- function(directory){
  seqs_raw <- read.fasta(directory)
  seqs <- data.frame("sequence" = map(seqs_raw, ~paste(., collapse = "")) %>% unlist)
  seqs$sequence <- toupper(seqs$sequence)
  seqs$uniprotkb <- strsplit(rownames(seqs), "\\|") %>% map(~.[2]) %>% unlist
  return(seqs)
}
##Load proteomes
uniprotHsIso <- read_fastaseq(paste(home_directory, 
                                         "data/biol_databases/uniprot/uniprotkb_proteome_UP000005640_2025_02_19.fasta", sep = ""))
uniprotHsIsoRevi <- read_fastaseq(paste(home_directory, 
                                         "data/biol_databases/uniprot/uniprotkb_proteome_UP000005640_AND_revi_2025_02_19.fasta", sep = ""))
```


# 1. Map protein identifiers to uniprot
## Perform mapping
### Set up
```{r}
ubnOmicsSites <- read_csv(file = "data/features/ubnOmics/ensp_site_canonical.csv")
identifsSpecies <-ubnOmicsSites %>% split(.$scientific_name) %>% map(~unique(.$ensp))
sampleIdentifs <- ubnOmicsSites %>% split(.$scientific_name) %>% map(~unique(.$ensp)[1:10])

identifMaps <- list()
```


### Map using uniprot API
```{r}
##Databases to use
unipDbSpecies <- list(
  "Arabidopsis thaliana" = "Araport"
)

##Extra processing of identifies
arabIDs <- data.frame("ID" = identifsSpecies$`Arabidopsis thaliana`,
                      "IDstrp" = gsub("\\..+$", "", identifsSpecies$`Arabidopsis thaliana`))

##Perform mapping
arabMap <- mapUniProt(
    from = "Araport",
    to = "UniProtKB",
    query = unique(arabIDs$IDstrp)
)

##Add the .1 back
arabMapProc <- left_join(
  dplyr::rename(arabIDs, "From"= "ID"),
  dplyr::rename(arabMap, "IDstrp" = "From"),
  by = "IDstrp"
)

##Assign
identifMaps$`Arabidopsis thaliana` <- arabMapProc
```


### Map using annotationDBI
```{r}
library(AnnotationDbi)
library(org.Sc.sgd.db)
identifMaps[["Saccharomyces cerevisiae"]] <- AnnotationDbi::select(org.Sc.sgd.db, 
                                                                   keys = identifsSpecies[["Saccharomyces cerevisiae"]], keytype = "ORF",
                                                                   columns = "UNIPROT")
```

### Use BioMart manually
```{r}
#Export
expSpecies <- c("Caenorhabditis elegans", "Drosophila melanogaster", 
                "Gallus gallus", 
                "Mus musculus", "Rattus norvegicus")
for (spec in expSpecies){
  dat <- data.frame("identif" = identifsSpecies[[spec]])
  write_tsv(dat, file = paste0("output/data/hotspots/identifiers/IDs_", spec, ".tsv"),
            col_names = F)
}
#Load back in
for (spec in expSpecies){
  identifMaps[[spec]] <- read_delim(paste0("output/data/hotspots/identifiers/output_", spec, ".txt"))
}

map(identifMaps, nrow)
names(identifMaps)
```


## Clean results and compare to uniprot proteomes
### Clean results
```{r}
#Functions
biomartPickUnip <- function(dat){
  mutate(dat, uniprot = ifelse(!is.na(`UniProtKB/Swiss-Prot ID`), `UniProtKB/Swiss-Prot ID`, `UniProtKB/TrEMBL ID`))
}
#Rename and pick identifiers
identifMaps$`Arabidopsis thaliana` <- identifMaps$`Arabidopsis thaliana` %>% dplyr::rename("uniprot" = "Entry")
identifMaps$`Saccharomyces cerevisiae` <- identifMaps$`Saccharomyces cerevisiae` %>% dplyr::rename("uniprot" = "UNIPROT", "From" = "ORF")
identifMaps$`Caenorhabditis elegans` <- identifMaps$`Caenorhabditis elegans` %>% dplyr::rename("From" = "WormBase CDS ID") %>% biomartPickUnip
identifMaps$`Drosophila melanogaster` <- identifMaps$`Drosophila melanogaster` %>% dplyr::rename("From" = "Protein stable ID") %>% biomartPickUnip
identifMaps$`Mus musculus` <- identifMaps$`Mus musculus` %>% dplyr::rename("From" = "Protein stable ID") %>% biomartPickUnip
identifMaps$`Rattus norvegicus` <- identifMaps$`Rattus norvegicus` %>% dplyr::rename("From" = "Protein stable ID") %>% biomartPickUnip
identifMaps$`Gallus gallus` <- identifMaps$`Gallus gallus` %>% dplyr::rename("From" = "Protein stable ID") %>% biomartPickUnip

#Add C elegans extra results
celegans_protID <- read_tsv("output/data/hotspots/identifiers/output_PROTEINID_Caenorhabditis elegans.txt") %>%
  mutate(From = `Protein stable ID`) %>% biomartPickUnip
identifMaps$`Caenorhabditis elegans` <- rbind(identifMaps$`Caenorhabditis elegans`, celegans_protID)

#export
save(identifMaps, file = "output/data/hotspots/identifiers/identifMaps__2.RData")
```

```{r}
# Checkpoint
#load("output/data/hotspots/identifiers/identifMaps__2.RData")
```


### Load proteomes
```{r}
#Function to load proteomes
library(seqinr)
read_fastaseq <- function(directory){
  seqs_raw <- read.fasta(directory)
  seqs <- data.frame("sequence" = map(seqs_raw, ~paste(., collapse = "")) %>% unlist)
  seqs$sequence <- toupper(seqs$sequence)
  seqs$uniprotkb <- strsplit(rownames(seqs), "\\|") %>% map(~.[2]) %>% unlist
  return(seqs)
}

#Set up directories
proteomeCodes <- c(
  "Arabidopsis thaliana" = "UP000006548",
  "Saccharomyces cerevisiae" = "UP000002311",
  "Caenorhabditis elegans" = "UP000001940",
  "Drosophila melanogaster" = "UP000000803",
  "Mus musculus" = "UP000000589",
  "Rattus norvegicus" = "UP000002494",
  "Gallus gallus" = "UP000000539"
)
# Checkpoint
#load("output/data/hotspots/identifiers/prtmAllLst.RData")
```


Activate this chunk when you want to reload proteomes from scratch. Otherwise just run the chunk above
```{r}
prtmDirsAll <- map(proteomeCodes, ~paste0(home_directory, "data/biol_databases/uniprot/uniprotkb_proteome_",
                                          ., "_2025_02_19.fasta"))
prtmDirsRevi <- map(proteomeCodes, ~paste0(home_directory, "data/biol_databases/uniprot/uniprotkb_proteome_",
                                          ., "_AND_revi_2025_02_19.fasta"))
prtmDirsRevi$`Saccharomyces cerevisiae` <- prtmDirsAll$`Saccharomyces cerevisiae`

#Load in proteomes
prtmAllLst <- map(prtmDirsAll, read_fastaseq)
prtmReviLst <- map(prtmDirsRevi, read_fastaseq)
for (spec in names(prtmAllLst)){
  prtmAllLst[[spec]]$revi <- prtmAllLst[[spec]]$uniprotkb %in% prtmReviLst[[spec]]$uniprotkb
}
#Extra arabidopsis proteome
arabPrtmExtra <- read_fastaseq(paste0(home_directory, "data/biol_databases/uniprot/uniprotkb_proteome_",
                                          "UP000426265", "_2025_02_19.fasta")) %>%
  mutate(revi = F)
prtmAllLst$`Arabidopsis thaliana` <- rbind(prtmAllLst$`Arabidopsis thaliana`,
                                           subset(arabPrtmExtra, uniprotkb %in% prtmAllLst$`Arabidopsis thaliana`$uniprotkb == F))
save(prtmAllLst, file = "output/data/hotspots/identifiers/prtmAllLst.RData")
```


### Compare matches to proteomes
```{r}
#Flag reviewed matches
for (spec in names(identifMaps)){
  identifMaps[[spec]]$inPrtm <- identifMaps[[spec]]$uniprot %in% prtmAllLst[[spec]]$uniprotkb
  identifMaps[[spec]]$inRevi <- identifMaps[[spec]]$uniprot %in% subset(prtmAllLst[[spec]], revi == T)$uniprotkb
}

#Calculate numbers
matchNumM <- map(names(identifMaps), function(spec){
  mapDat <- identifMaps[[spec]]
  input <- identifsSpecies[[spec]]
  prtm <- prtmAllLst[[spec]]
  mapDat <- mapDat[!is.na(mapDat$uniprot), ]
  #How many matches?
  inputNum <- length(input)
  matchNum <- length(which(input %in% mapDat$From))
  #How many matches in proteome?
  matchPrtmNum <- length(which(input %in% subset(mapDat, uniprot %in% prtm$uniprotkb)$From))
  matchReviNum <- length(which(input %in% subset(mapDat, uniprot %in% subset(prtm, revi == T)$uniprotkb)$From))
  #Output
  numVec <- c(inputNum, matchNum, matchPrtmNum, matchReviNum)
  return(numVec)
}) %>% purrr::reduce(rbind)
rownames(matchNumM)  <- names(identifMaps)
colnames(matchNumM) <- c("inputNum", "matchNum", "matchPrtmNum", "matchReviNum")
```

### Fuse matches
```{r}
identifMapDf <- map_dfr(identifMaps, function(dat){dat[, c("From", "uniprot", "inPrtm", "inRevi")]}, .id = "scientific_name")
nrow(identifMapDf)
nrow(distinct(identifMapDf))
length(unique(identifMapDf$uniprot))
```



# 2. Non-human mapping: Pick uniprots
```{r}
identifMapDf <- identifMapDf[which(!is.na(identifMapDf$uniprot)), ] #Remove NAs

#Make map data to hold map choices
mapResultsDf <- map_dfr(identifsSpecies[-which(names(identifsSpecies) == "Homo sapiens")], function(x){data.frame("From" = unique(x))}, .id = "scientific_name")
##Inspect number of matches at different levels
numsTotal <- identifMapDf %>% distinct(., From, uniprot, .keep_all = T) %>% group_by(From) %>% 
  summarise(numMatches = n(), 
            numPrtmMatches = sum(inPrtm), numReviMatches = sum(inRevi))
mapResultsDf <- left_join(mapResultsDf, numsTotal, by = "From")
mapResultsDf[is.na(mapResultsDf)] <- 0

mapResultsDf %>% split(.$scientific_name) %>% map(~table(.$numReviMatches))

#Make matches
pickUnip <- function(fromTemp){
  mapDat <- subset(identifMapDf, From == fromTemp)
  #Pick Revi if we can
  reviInds <- which(mapDat$inRevi == T)
  prtmInds <- which(mapDat$inPrtm == T)
  if (length(reviInds) > 0){
    choiceInds <- reviInds
  #Else pick in prtm if we can
  } else if (length(prtmInds) > 0) {
    choiceInds <- prtmInds
  } else {
    choiceInds <- 1:nrow(mapDat)
  }
  choice <- mapDat$uniprot[choiceInds[1]]
  return(choice)
}
mapResultsDf$uniprot <- NA
for (i in 1:nrow(mapResultsDf)){
  fromTemp <- mapResultsDf$From[i]
  if (fromTemp %in% identifMapDf$From){
    mapResultsDf$uniprot[i] <- pickUnip(fromTemp)
  }
}
sum(is.na(mapResultsDf$uniprot))

#Numbers for mapping
mapResultsDf <- left_join(mapResultsDf, distinct(identifMapDf[, c("uniprot", "inPrtm", "inRevi")]), by = "uniprot")
mapResultsDf$startingNum <- TRUE
mapResultsDf$anyMatch <- !is.na(mapResultsDf$uniprot)
boolCols <- c("startingNum", "anyMatch", "inPrtm", "inRevi")
for (col in boolCols){
  mapResultsDf[which(is.na(mapResultsDf[, col])), col] <- FALSE
}
mat <- mapResultsDf %>% split(.$scientific_name) %>% map(~colSums(.[, boolCols])) %>% purrr::reduce(rbind)
rownames(mat) <- sort(unique(mapResultsDf$scientific_name))

#Manually inspect those with multiple reviewed matches
subset(mapResultsDf, numReviMatches > 2) %>% .[, c("From", "uniprot", "numReviMatches")]
##Mouse histones
mouseHistoneENSP <- read_delim("output/data/hotspots/identifiers/IDs_mouse_histone.txt") %>% 
  .[, c("Protein stable ID", "Gene name", "Gene Synonym")] %>% distinct %>% as.data.frame
mouseHistoneUnips <- mapUniProt(
    from = "UniProtKB",
    to = "UniProtKB",
    query = subset(identifMapDf, From == "ENSMUSP00000077477" & inRevi == T)$uniprot
)
histoneMap <- list( #I manualy map these 9 proteins by using common gene names/synonyms
  c("ENSMUSP00000077477", "H2ac4", "C0HKE1"),
  c("ENSMUSP00000080088", "Hist1h2ap", "C0HKE9"),
  c("ENSMUSP00000088281", "H2ac7", "C0HKE3"),
  c("ENSMUSP00000088285", "Hist1h2ai", "C0HKE6"),
  c("ENSMUSP00000089335", "Hist1h2ag", "C0HKE5"),
  c("ENSMUSP00000089339", "Hist1h2ao", "C0HKE8"),
  c("ENSMUSP00000089345", "Hist1h2an", "C0HKE7"),
  c("ENSMUSP00000100034", "H2ac8 Hist1h2ae", "C0HKE4"),
  c("ENSMUSP00000127684", "H2ac6 Hist1h2ac", "C0HKE2")
) %>% purrr::reduce(rbind) %>% as.data.frame
colnames(histoneMap) <- c("ENSP", "gene", "uniprot")
for (i in 1:nrow(histoneMap)){
  mapResultsDf[which(mapResultsDf$From == histoneMap$ENSP[i]), "uniprot"] <- histoneMap$uniprot[i]
}
##Calmodulins
calmMap <- list( #I manually map these 3 genes by using common gene names
  c("ENSMUSP00000019514", "Calm3", "P0DP28"),
  c("ENSMUSP00000048857", "Calm2", "P0DP27"),
  c("ENSMUSP00000105709", "Calm1", "P0DP26")
) %>% purrr::reduce(rbind) %>% as.data.frame
colnames(calmMap) <- c("ENSP", "gene", "uniprot")
for (i in 1:nrow(calmMap)){
  mapResultsDf[which(mapResultsDf$From == calmMap$ENSP[i]), "uniprot"] <- calmMap$uniprot[i]
}
```



# 3. Non-human uniprots: Map sites
## Get sequence windows (slow)
```{r}
#Load in
ubnOmicsProtSeqs <- read_csv("data/features/ubnOmics/proteins_all.csv") %>%
  dplyr::rename("ensp" = "id.2")
#Are all proteins present?
ubnOmicsSites %>% subset(ensp %in% ubnOmicsProtSeqs$ensp == F)

#Extract sequence windows
ubnOmicsSites <- ubnOmicsSites[order(ubnOmicsSites$scientific_name, ubnOmicsSites$ensp), ]
unipsOffset <- c("", ubnOmicsSites$ensp[-nrow(ubnOmicsSites)])
ubnOmicsSites$newUnip <- ubnOmicsSites$ensp != unipsOffset
ubnOmicsSites$ensp_seqWind <- NA
unipSeq <- ""
##Loop over species to subset data
species <- unique(ubnOmicsProtSeqs$scientific_name)
for (spec in species){
  inds <- which(ubnOmicsSites$scientific_name == spec)
  seqDat <- subset(ubnOmicsProtSeqs, scientific_name == spec)
  for (i in inds){
    #Get objects
    ensp <- ubnOmicsSites$ensp[i]
    pos <- ubnOmicsSites$position[i]
    #Get protein sequence
    if (ubnOmicsSites$newUnip[i]){
      unipSeq <- seqDat$sequence[which(seqDat$ensp == ensp)]
      unipSeq <- paste(c(rep("_", 7), unipSeq, rep("_", 7)), collapse = "")
    }
    #Get position and sequence window
    ubnOmicsSites$ensp_seqWind[i] <- substr(unipSeq, start = pos, stop = pos + 14)
  }
}
ubnOmicsSitesSeqwinds <- ubnOmicsSites
save(ubnOmicsSitesSeqwinds, file = "output/data/ubnOmicsSitesSeqwinds.RData")

#Check that central positions are lysines
ubnOmicsSitesSeqwinds$AA <- substr(ubnOmicsSitesSeqwinds$ensp_seqWind, 8, 8)
subset(ubnOmicsSitesSeqwinds, AA != "K")
```

```{r}
# Checkpoint
#load("data/features/ubnOmics/ubnOmicsSitesSeqwinds.RData")
```



## Mapping
### Set up
```{r}
#Set up map and get uniprot selections
ubnOmicsSites_nonHsMap <- ubnOmicsSitesSeqwinds %>% subset(scientific_name != "Homo sapiens") %>% .[, c("ensp", "position", "scientific_name", "ensp_seqWind")] %>% distinct
ubnOmicsSites_nonHsMap <- left_join(ubnOmicsSites_nonHsMap, 
                                    mapResultsDf[, c("From", "uniprot")] %>% dplyr::rename("ensp" = "From"),
                                    by = "ensp")
#Flag missing uniprots to be dealt with later
ubnOmicsSites_nonHsMap$uniprot[which(is.na(ubnOmicsSites_nonHsMap$uniprot))] <- "missing"
```

### Map into uniprot sequences, where possible
```{r}
#Set up looping in chunks of species and uniprots
ubnOmicsSites_nonHsMap <- ubnOmicsSites_nonHsMap[order(ubnOmicsSites_nonHsMap$scientific_name, ubnOmicsSites_nonHsMap$uniprot), ]
unipsOffset <- c("", ubnOmicsSites_nonHsMap$ensp[-nrow(ubnOmicsSites_nonHsMap)])
ubnOmicsSites_nonHsMap$newUnip <- ubnOmicsSites_nonHsMap$ensp != unipsOffset
unipSeq <- ""
ubnOmicsSites_nonHsMap$matches <- NA

#Loop over species
species <- unique(ubnOmicsSites_nonHsMap$scientific_name)
for (spec in species){
  #Get the relevant proteome and subset it
  specUnips <- subset(ubnOmicsSites_nonHsMap, scientific_name == spec)$uniprot %>% unique
  seqDat <- prtmAllLst[[spec]] %>% subset(uniprotkb %in% specUnips) 
  #Loop over sites in this proteome
  inds <- which(ubnOmicsSites_nonHsMap$scientific_name == spec)
  for (i in inds){
    #Get objects
    unip <- ubnOmicsSites_nonHsMap$uniprot[i]
    seqWind <- ubnOmicsSites_nonHsMap$ensp_seqWind[i]
    #Only proceed if uniprot is in the proteome
    if (!is.na(unip) & unip %in% seqDat$uniprot){
      #Get uniprot sequence
      if (ubnOmicsSites_nonHsMap$newUnip[i]){
        unipSeq <- seqDat$sequence[which(seqDat$uniprot == unip)]
        unipSeq <- paste(c(rep("_", 7), unipSeq, rep("_", 7)), collapse = "")
      }
      #Match sequence window in
      matchesDf <- stringr::str_locate_all(unipSeq, seqWind)[[1]]
      if (nrow(matchesDf) == 0){
        matches <- NA
      } else {
        newPosns <- as.numeric(matchesDf[, 1])
        matches <- map(newPosns, ~paste0(unip, "_", .)) %>% unlist
        matches <- paste(matches, collapse = ";")
      }
      ubnOmicsSites_nonHsMap$matches[i] <- matches
    }
  }
}

#Inspect matches
ubnOmicsSites_nonHsMap$matches_num <- strsplit(ubnOmicsSites_nonHsMap$matches, ";") %>% 
  map(function(x){
    if (is.na(x[1])){ return(0) } else { return(length(x))}
  }) %>% unlist

table(ubnOmicsSites_nonHsMap$scientific_name, ubnOmicsSites_nonHsMap$matches_num)

temp <- subset(ubnOmicsSites_nonHsMap, uniprot != "missing")
table(temp$scientific_name, temp$matches_num)
```

### Map into the entire proteome when there are no matches
```{r}
#Perform mapping
ubnOmicsSites_nonHsMap$matchesWholeProt <- NA
for (spec in species){
  #Get the relevant proteome and subset it
  seqDat <- prtmAllLst[[spec]]
  buff <- paste(rep("_", 7), collapse = "")
  seqDat$sequence <- paste(buff, seqDat$sequence, buff, sep = "") #Pad all sequences
  #Loop over sites in this proteome that have no matches
  inds <- which(ubnOmicsSites_nonHsMap$scientific_name == spec &
                  ubnOmicsSites_nonHsMap$matches_num == 0)
  for (i in inds){
    #Get objects
    seqWind <- ubnOmicsSites_nonHsMap$ensp_seqWind[i]
    #Mapping
    mapping <- stringr::str_locate_all(seqDat$sequence, seqWind)
    mappingBool <- map(mapping, ~(nrow(.) > 0)) %>% unlist
    mappingInds <- which(mappingBool == TRUE)
    if (length(mappingInds) == 0){ ##Return NULL if no map
      output <- NA
    } else {
      ##Extract matches
      matchesLst <- map(mappingInds, function(i){ #This accounts for if there are multiple matches in the same protein
        matchesDf <- mapping[[i]]
        unip <- seqDat$uniprotkb[i]
        newPosns <- as.numeric(matchesDf[, 1])
        matches <- map(newPosns, ~paste0(unip, "_", .)) %>% unlist
        matches <- paste(matches, collapse = ";")
        return(matches)
      }) %>% 
        unlist %>% paste(., collapse = ";") #Collapse multiple matches into one vector
      output <- matchesLst
    }
    ubnOmicsSites_nonHsMap$matchesWholeProt[i] <- output
  }
} 

#Get reviewed only 
wholeProtMatchesLst <- strsplit(ubnOmicsSites_nonHsMap$matchesWholeProt, ";")
wholeProtMatchesLst_Unips <- map(wholeProtMatchesLst, ~gsub("_\\d+$", "", .))
wholeProtMatchesLst_revUnipsInds <- wholeProtMatchesLst_Unips
for (spec in unique(ubnOmicsSites_nonHsMap$scientific_name)){
  prtm <- prtmAllLst[[spec]] %>% subset(revi == TRUE)
  for (i in which(ubnOmicsSites_nonHsMap$scientific_name == spec)){
    wholeProtMatchesLst_revUnipsInds[[i]] <- wholeProtMatchesLst_revUnipsInds[[i]] %in% prtm$uniprotkb
  }
}
wholeProtMatchesLst_revi <- map2(wholeProtMatchesLst, wholeProtMatchesLst_revUnipsInds, 
                                 function(x, y){x[y]})
ubnOmicsSites_nonHsMap$matchesWholeProt_revi <- map(wholeProtMatchesLst_revi, function(x){
  if (length(x) == 0){
    return(NA)
  } else {
    return(paste(x, collapse = ";"))
  }
}) %>% unlist

#Inspect matches
ubnOmicsSites_nonHsMap$matchesWholeProts_num <- strsplit(ubnOmicsSites_nonHsMap$matchesWholeProt, ";") %>% 
  map(function(x){
    if (is.na(x[1])){ return(0) } else { return(length(x))}
  }) %>% unlist
ubnOmicsSites_nonHsMap$matchesWholeProt_reviNum <- strsplit(ubnOmicsSites_nonHsMap$matchesWholeProt_revi, ";") %>% 
  map(function(x){
    if (is.na(x[1])){ return(0) } else { return(length(x))}
  }) %>% unlist
subset(ubnOmicsSites_nonHsMap, matches_num == 0)$matchesWholeProts_num %>% table
subset(ubnOmicsSites_nonHsMap, matches_num == 0)$matchesWholeProt_reviNum %>% table

temp <- subset(ubnOmicsSites_nonHsMap, matches_num == 0)
table(temp$scientific_name, temp$matchesWholeProt_reviNum)
```

### Select final sites
```{r}
#Set up
ubnOmicsSites_nonHsMap$finalMatches <- ubnOmicsSites_nonHsMap$matches

#Proteome-wide matches: Pick reviewed where possible
for (i in which(ubnOmicsSites_nonHsMap$matches_num == 0)){
  if (ubnOmicsSites_nonHsMap$matchesWholeProt_reviNum[i] > 0){
    ubnOmicsSites_nonHsMap$finalMatches[i] <- ubnOmicsSites_nonHsMap$matchesWholeProt_revi[i]
  } else {
    ubnOmicsSites_nonHsMap$finalMatches[i] <- ubnOmicsSites_nonHsMap$matchesWholeProt[i]
  }
}

#When there are still multiple choices, prioritise those closest to the original position
ubnOmicsSites_nonHsMap$finalMatchesNum <- strsplit(ubnOmicsSites_nonHsMap$finalMatches, ";") %>% 
  map(function(x){
    if (is.na(x[1])){ return(0) } else { return(length(x))}
  }) %>% unlist
ubnOmicsSites_nonHsMap$finalMatch <- ubnOmicsSites_nonHsMap$finalMatches
for (i in 1:nrow(ubnOmicsSites_nonHsMap)){
  if (ubnOmicsSites_nonHsMap$finalMatchesNum[i] > 1){
    pos <- ubnOmicsSites_nonHsMap$position[i]
    matches <- strsplit(ubnOmicsSites_nonHsMap$finalMatches[i], ";")[[1]]
    matchesPos <- gsub("^.+_", "", matches) %>% as.numeric
    closestInds <- which.min(abs(matchesPos - pos))
    match <- matches[closestInds[1]]
    ubnOmicsSites_nonHsMap$finalMatch[i] <- match
  }
}

#Inspect
table(!is.na(ubnOmicsSites_nonHsMap$finalMatch))
```

### Tidy and export
```{r}
ubnOmicsSites_nonHsMap <- ubnOmicsSites_nonHsMap %>%
  dplyr::rename("uniprotOld" = "uniprot",
                "positionOld" = "position",
                "uniprot_pos" = "finalMatch") %>%
  mutate(uniprot = gsub("_\\d+$", "", uniprot_pos),
         position = gsub("^.+_", "", uniprot_pos))
ubnOmicsSites_nonHsMapSub <- subset(ubnOmicsSites_nonHsMap, !is.na(uniprot_pos))
#save(ubnOmicsSites_nonHsMapSub, file = "output/data/hotspots/ubnOmicsSites_nonHsMapSub__1.RData")
```


```{r}
# Checkpoint
# load("output/data/hotspots/ubnOmicsSites_nonHsMapSub__1.RData")
```


# 4. Combine mapped ubn sites

## Load human sites
I use sites from the PRIDE build
```{r}
load("data/features/ubiFeatProc__1.RData")

humanSites <- ubiFeatProc %>%
                           mutate(scientific_name = "Homo sapiens") %>%
                           dplyr::rename("IDpos" = "Cons_ensp_pos", "sequence_window" =  "unipSeqWind") %>% 
                           .[, c("IDpos", "scientific_name", "uniprot_pos", "sequence_window")]
```

## Combine
```{r}
ubnSitesUnipmap <- rbind(humanSites,
                         ubnOmicsSites_nonHsMapSub %>% 
                           mutate(IDpos = paste(ensp, positionOld, sep = "_")) %>%
                           dplyr::rename("sequence_window" = "ensp_seqWind") %>% 
                           .[, c("IDpos", "scientific_name", "uniprot_pos", "sequence_window")]) %>%
  mutate(uniprot = gsub("_\\d+$", "", uniprot_pos),
         position = gsub("^.+_", "", uniprot_pos) %>% as.numeric)
ubnSitesUnipmap <- subset(ubnSitesUnipmap, !is.na(uniprot_pos))
save(ubnSitesUnipmap, file = "output/data/hotspots/ubnSitesUnipmap__2.RData")
# Checkpoint
#load("output/data/hotspots/ubnSitesUnipmap__2.RData")
```


# 5. Extract InterPro domains with API call

Load intermediates, in case you didn't run the above chunks
```{r}
load("output/data/hotspots/identifiers/prtmAllLst.RData")
load("output/data/hotspots/ubnSitesUnipmap__2.RData")
#load("output/data/hotspots/domains/IPRresLst__4.RData")
```



## Run API call, using cached results
```{r}
# Load from cache
cacheF <- "output/data/hotspots/domains/IPRresLst__cache.RData"
if (file.exists(cacheF)){
  load(cacheF) # Data is a list called IPRresLst
} else {
  IPRresLst <- list()
}

# Set up uniprots
apiUnips <- unique(ubnSitesUnipmap$uniprot)
apiUnipsMissing <- setdiff(apiUnips, names(IPRresLst))

# Run API on any uniprots that are missing
if (length(apiUnipsMissing) > 0){
  #Run APIs
  start <- Sys.time()
  IPRresLstNew <- list()
  for (i in 1:length(apiUnipsMissing)){
    if (i %% 10 == 0){print(i)}
    unip <- apiUnipsMissing[i]
    IPRresLstNew[[i]] <- callInterproAPI(unip)
  }
  names(IPRresLstNew) <- apiUnipsMissing
  end <- Sys.time()
  print(end - start)
} else {
  IPRresLstNew <- list()
}

# Update and save cache
IPRresLst <- c(IPRresLst, IPRresLstNew)
save(IPRresLst, file = cacheF)
```

## Process domain results
```{r}
#Combine
IPRresDfLst <- map2(IPRresLst, names(IPRresLst), function(res, unip){
  dat <- processInterproAPI(res)
  if (!is.null(dat)){
    dat$uniprot <- unip
    dat <- as.data.table(dat)
  }
  return(dat)
})
boolVec <- map(IPRresDfLst, ~!is.null(.)) %>% unlist
IPRresDf <- rbindlist(IPRresDfLst[boolVec])
```

## Check that uniprots from the list names match uniprots from results
```{r}
uniprotComp <- map2(IPRresLst, names(IPRresLst), function(res, unipFLst){
  if (is.null(res)){
    unipFRes <- NA
  } else {
    unipFRes <- toupper(res$results$proteins[[1]]$accession)
  }
  return(c(unipFLst, unipFRes))
})
unipComp <- purrr::reduce(uniprotComp, rbind)
unipComp <- as.data.frame(unipComp)
unipComp$comp <- unipComp$V1 == unipComp$V2
table(unipComp$comp)
```


## Inspect domain results 
```{r}
#What failed?
failedBool <- map(IPRresLst, is.null) %>% unlist
failedNames <- names(IPRresLst)[failedBool]
finalUnips <- ubnSitesUnipmap[, c("scientific_name", "uniprot")] %>% distinct
finalUnips$IPRfailed <- finalUnips$uniprot %in% failedNames

##Numbers that failed
matchNumM <- finalUnips %>% split(.$scientific_name) %>% map(function(mapDat){
  spec <- mapDat$scientific_name[1]
  input <- identifsSpecies[[spec]]
  prtm <- prtmAllLst[[spec]]
  mapDat <- mapDat[!is.na(mapDat$uniprot), ]
  #How many matches?
  inputNum <- length(input)
  #How many uniprots failed?
  failNum <- subset(mapDat, IPRfailed == T)$uniprot %>% unique %>% length
  #How many matches didn't fail?
  matchNumNoFail <- length(which(input %in% subset(mapDat, IPRfailed == F)$From))
  #Output
  numVec <- c(inputNum, failNum, matchNumNoFail)
  return(numVec)
}) %>% purrr::reduce(rbind)
rownames(matchNumM)  <- sort(unique(identifMapDf$scientific_name))
colnames(matchNumM) <- c("inputNum", "failNum", "matchNumNoFail")

##Inspect those that failed
finalUnips %>% split(.$scientific_name) %>% map(~subset(., IPRfailed == T)$uniprot %>% unique %>% .[1:20])
```


Protein lengths
```{r}
prtmProtLengths <- map_dfr(prtmAllLst, function(x){
  x$length <- str_length(x$sequence)
  return(x[, c("uniprotkb", "length")])
}, .id = "scientific_name")
prtmProtLengths <- rbind(prtmProtLengths, 
                         uniprotHsIso %>% 
                           mutate(length = str_length(sequence),
                                                 scientific_name = "Homo sapiens") %>%
                           .[, c("scientific_name", "uniprotkb", "length")])

IPRprotLengths <- data.frame("length" = map(IPRresLst, ~.$results$proteins[[1]]$protein_length) %>% unlist,
                             "uniprotkb" = map(IPRresLst, ~.$results$proteins[[1]]$accession) %>% unlist) %>%
  mutate(uniprotkb = toupper(uniprotkb))
IPRprotLengths <- left_join(IPRprotLengths, dplyr::rename(prtmProtLengths, "lengthPrtm" = "length"), by = "uniprotkb") %>%
  mutate(comp = length == lengthPrtm)
subset(IPRprotLengths, comp == F)
```



# 6. Prepare InterPro domain results
## Map in ubn sites
```{r}
#Split out multiple matches within one protein into separate entires
IPRresDfSplt <- separate_rows(IPRresDf, c("start", "end"), sep = ";") %>%
  mutate(start = as.numeric(start), end = as.numeric(end))

#Subset for domains containing ubn sites
##Define total catchment per protein
domainVecs <- map(1:nrow(IPRresDfSplt), ~IPRresDfSplt$start[.]:IPRresDfSplt$end[.])
##Get positions per uniprot
posnVecs <- ubnSitesUnipmap %>% split(.$uniprot) %>% map(~.$position)
##Intersect uniprot positions with domain positions
intersectVec <- map(1:length(domainVecs), function(i){
  unip <- IPRresDfSplt$uniprot[i]
  posns <- posnVecs[[unip]]
  return(intersect(posns, domainVecs[[i]]))
})
head(intersectVec)
IPRresDfSplt$ubnSites <- map(intersectVec, function(x){
  if (length(x) == 0){
    return(NA)
  } else {
    return(paste(x, collapse = ";"))
  }
}) %>% unlist
IPRresDfSplt$hasUbnSites <- !is.na(IPRresDfSplt$ubnSites)
table(IPRresDfSplt$hasUbnSites)
IPRresDfSpltSub <- subset(IPRresDfSplt, hasUbnSites == T)
IPRresDfSpltSub$ubnSiteNum <- strsplit(IPRresDfSpltSub$ubnSites, ";") %>% map(length) %>% unlist
```

## Extract table of all proteins, their domains, and their ubi-sites
```{r}
UniprotDomainUbisites <- IPRresDfSpltSub
save(UniprotDomainUbisites, file = "output/data/hotspots/UniprotDomainUbisites__2.RData")
# Checkpoint
#load("output/data/hotspots/UniprotDomainUbisites__2.RData")
#IPRresDfSpltSub <- UniprotDomainUbisites
```


## Extract domain sequences
```{r}
#Add scientific name
IPRresDfSpltSub <- left_join(IPRresDfSpltSub, distinct(ubnSitesUnipmap[, c("uniprot", "scientific_name")]), by = "uniprot")

#Extract domain sequences
IPRresDfSpltSub <- IPRresDfSpltSub[order(IPRresDfSpltSub$scientific_name, IPRresDfSpltSub$uniprot), ]
unipsOffset <- c("", IPRresDfSpltSub$uniprot[-nrow(IPRresDfSpltSub)])
IPRresDfSpltSub$newUnip <- IPRresDfSpltSub$uniprot != unipsOffset
IPRresDfSpltSub$sequence <- NA
unipSeq <- ""
##Loop over species to subset data
species <- unique(IPRresDfSpltSub$scientific_name)
for (spec in species){
  inds <- which(IPRresDfSpltSub$scientific_name == spec)
  if (spec == "Homo sapiens"){
    seqDat <- uniprotHsIso
  } else {
    seqDat <- prtmAllLst[[spec]]
  }
  for (i in inds){
    #Get objects
    unip <- IPRresDfSpltSub$uniprot[i]
    start <- IPRresDfSpltSub$start[i]
    end <- IPRresDfSpltSub$end[i]
    #Get protein sequence
    if (IPRresDfSpltSub$newUnip[i]){
      unipSeq <- seqDat$sequence[which(seqDat$uniprotkb == unip)]
    }
    #Get position and sequence window
    IPRresDfSpltSub$sequence[i] <- substr(unipSeq, start = start, stop = end)
  }
}
```


## Filter domains
```{r}
IPRresDfSpltSub <- IPRresDfSpltSub %>%
   group_by(accession, uniprot) %>%
  mutate(UID = paste(accession, uniprot, row_number(), sep = "_")) %>%
  ungroup()
#Summarise at domain level
IPRresDfSpltSub_summ <- IPRresDfSpltSub[, c("accession", "pfam", "ubnSiteNum", "UID")] %>%
  group_by(accession) %>%
  summarise(instanceNum = n(), ubnSiteNum = sum(ubnSiteNum))
#Filter
filteredDomains <- subset(IPRresDfSpltSub_summ,
                          instanceNum >= 15 &
                          ubnSiteNum >= 50)
```


## Export for alignment
```{r}
for (j in 1:nrow(filteredDomains)){
  domainName <- filteredDomains$accession[j]
  domainDatSub <- subset(IPRresDfSpltSub, accession == domainName)
  #Prep fasta
  fastaNames <- map(1:nrow(domainDatSub), function(i){
    unip <- domainDatSub$uniprot[i]
    startEnd <- as.character(domainDatSub[i, c("start", "end")])
    paste(c(unip, startEnd), collapse = ";")
  }) %>% unlist
  fastaSeqs <- domainDatSub$sequence
  fastaSeqs <- BStringSet(fastaSeqs)
  names(fastaSeqs) <- fastaNames
  #Write fasta
  writeXStringSet(fastaSeqs, filepath = paste0("output/data/hotspots/domains/domainsPreAlign/", domainName, ".fasta"),
                  width = 20000)
}
```


## Format sites
```{r}
#Subset for tested domains
domainSites <- IPRresDfSpltSub %>% subset(accession %in% filteredDomains$accession) %>%
  separate_rows("ubnSites", sep = ";") %>%
  mutate(uniprot_pos = paste(uniprot, ubnSites, sep = "_")) %>%
  mutate(AA = "K")

#Format
domainSitesExp <- domainSites[, c("scientific_name", "uniprot", "ubnSites", "AA", "accession", "start", "end", "pfam")]
write_csv(domainSitesExp, file = "output/data/hotspots/ubnSites_forHotspots", col_names = F)
```


# 6b. Run the hotspot script 
- Move output from output/data/hotspots/domains/domainsPreAlign and output/data/hotspots/ubnSites_forHotspots to the cluster
- Run scripts/hotspot_scripts/02_runMAFFT_for_hotspots.py to align domain sequences
- Run scripts/hotspot_scripts/03_run_ptm_hotspots.py to run hotspot analysis
- Move output from output/data/hotspots/hotspotOutput.csv back to local

# 7. Process hotspot output
## Make table for hotspots
```{r}
hotspotRes <- read_csv("output/data/hotspots/domains/hotspotOutput/hotspotOutput.csv")
#Make table for individual hotspots
hotspotDf <- hotspotRes[, c("domain", "hotspot_id", "start_aln", "end_aln", "min_adj_pval")] %>% distinct 
##Add pfam info
hotspotDf <- left_join(hotspotDf, 
                       IPRresDfSpltSub[, c("accession", "pfam")] %>% dplyr::rename("domain" = "accession") %>% distinct,
                       by = "domain")
hotspotDfPfams <- hotspotDf %>% separate_rows("pfam", sep = ";")
length(unique(hotspotDfPfams$pfam))
##Extract domain names
hotspotDf[, c("interpro_name", "interpro_nameShort")] <- NA
for (i in 1:nrow(hotspotDf)){
  inteproID <- hotspotDf$domain[i]
  ###make API call
  base_url <- "https://www.ebi.ac.uk/interpro/api/entry/interpro/"
  url <- paste0(base_url, "/", inteproID, "?format=json")
  response <- GET(url)
  apiDat <- fromJSON(httr::content(response, as = "text"), flatten = F)
  ###Get output
  hotspotDf$interpro_name[i] <- apiDat$metadata$name$name
  hotspotDf$interpro_nameShort[i] <- apiDat$metadata$name$short
}
#Make domain identifier with domain domain
hotspotDf <- get_hotspot_idwName(data = hotspotDf, interpro_name = "interpro_name", hotspot_id = "hotspot_id")

#Extract # structures and representative structures
hotspotDf[, c("strRep", "strNum")] <- NA
for (i in 1:nrow(hotspotDf)){
  inteproID <- hotspotDf$domain[i]
  #Representative structure
  base_url <- "https://www.ebi.ac.uk/interpro/api/entry/interpro/"
  url <- paste0(base_url, "/", inteproID, "?format=json")
  response <- GET(url)
  apiDat <- fromJSON(httr::content(response, as = "text"), flatten = F)
  ###Get output
  res <- apiDat$metadata$representative_structure$accession
  if (!is.null(res)){
    hotspotDf$strRep[i] <- res[1]
  }
  
  #Number of structures
  base_url <- "https://www.ebi.ac.uk/interpro/api/structure/entry/interpro/"
  url <- paste0(base_url, "/", inteproID, "?format=json")
  response <- GET(url)
  apiDat <- fromJSON(httr::content(response, as = "text"), flatten = F)
  ###Get output
  res <- apiDat$structures$pdb
  if (!is.null(res)){
    hotspotDf$strNum[i] <- res[1]
  }
}

#Manually adjust representative structures
repStr_manual_lst <- list(
  c("IPR000719", "5uq2") #Protein kinase. I change to one with a resolved activation loop
)
for (x in repStr_manual_lst){
  hotspotDf$strRep[which(hotspotDf$domain == x[1])] <- x[2]
}
```


## Overview
```{r}
# Overivew
print("# hotspots")
print(nrow(hotspotDf))

print("# domains with hotspots")
print(length(unique(hotspotDf$domain)))
```


## Map hotspot regions into the corresponding domains
```{r}
# Map maps between absolute positions and alignment positions
alignDir <- "output/data/hotspots/domains/domainsPostAlignUnwrap"
alignColmaps <- list()
for (domainName in unique(hotspotDf$domain)){
  fastaF <- paste0(alignDir, "/", domainName, ".fasta")
  alignment_position_maps <- get_alignmentMap(fastaF)
  alignColmaps[[domainName]] <- alignment_position_maps
}
# Collect all positions present in each hotspot
hotspot_absolute_regions <- list()
k <- 1
for (i in 1:nrow(hotspotDf)){
  hotspot_id_sel <- hotspotDf$hotspot_id[i]
  domainName <- gsub("_.+$", "", hotspot_id_sel)
  hotspot <- hotspotDf$start_aln[i]:hotspotDf$end_aln[i]
  # Get alignment map
  alignColmap <- alignColmaps[[domainName]]
  # Loop over alignment map and extract absolute positions in the hotspot region
  for (j in 1:length(alignColmap)){
    domainId <- names(alignColmap)[j]
    mapp <- alignColmap[[j]]
    absoluteRegion <- subset(mapp, alignment_position %in% hotspot)$absolute_position
    if (length(absoluteRegion) > 0){
      resLst <- list(
        "domainId" = domainId, "pos" = absoluteRegion, "hotspot" = hotspot_id_sel
      )
      hotspot_absolute_regions[[k]] <- resLst
      k <- k + 1
    }
  }
}
hotspotAbsoluteRegions <- map_dfr(hotspot_absolute_regions, function(lst){
  data.frame(
    "domainId" = lst$domainId, "hotspot" = lst$hotspot, "pos" = lst$pos
  ) %>% mutate(
    uniprot = strsplit(domainId, ";") %>% map(~.[1]) %>% unlist,
    domainStart = strsplit(domainId, ";") %>% map(~.[2]) %>% unlist %>% as.numeric,
    domainEnd = strsplit(domainId, ";") %>% map(~.[3]) %>% unlist %>% as.numeric
    )
}, .id = "id")
```



## Also extract absolute hotspot positions from hotspotResidueOutput 
```{r}
# Load residue-level hotspot results
hotspotResidueRes <- read_csv("output/data/hotspots/domains/hotspotOutput/hotspotResidueOutput.csv")

# Subset for alignment regions
hotspotAbsoluteRegions_fResidue <- hotspotResidueRes
hotspotAbsoluteRegions_fResidue$in_hotspot <- FALSE
for (i in 1:nrow(hotspotDf)){
  ind <- which(
    hotspotAbsoluteRegions_fResidue$domain == hotspotDf$domain[i] & 
    hotspotAbsoluteRegions_fResidue$position_aln >= hotspotDf$start_aln[i] & 
    hotspotAbsoluteRegions_fResidue$position_aln <= hotspotDf$end_aln[i]
  )
  hotspotAbsoluteRegions_fResidue$in_hotspot[ind] <- TRUE
}
hotspotAbsoluteRegions_fResidue <- subset(hotspotAbsoluteRegions_fResidue, in_hotspot == T)
```


## Annotate domainSites with presence in hotspot region
```{r}
domainSites$position <- gsub("^.+_", "", domainSites$uniprot_pos) %>% as.numeric

# from hotspotAbsoluteRegions
hotspotAbsoluteRegions$uniprot_pos <- paste(hotspotAbsoluteRegions$uniprot, hotspotAbsoluteRegions$pos, sep = "_")
## Subset for domain sites
hotspotAbsoluteRegions_siteSub <- subset(hotspotAbsoluteRegions, uniprot_pos %in% domainSites$uniprot_pos)
## Collapse hotspots
hotspotAbsoluteRegions_siteSub <- hotspotAbsoluteRegions_siteSub %>% 
  .[order(.$hotspot), ] %>% group_by(uniprot_pos) %>% 
  summarise(hotspot = paste(hotspot, collapse = ";")) %>% ungroup
## Add to domainSites
domainSites <- left_join(domainSites, hotspotAbsoluteRegions_siteSub, by = "uniprot_pos")
domainSites$in_hotspot <- !is.na(domainSites$hotspot)
domainSites <- domainSites %>% dplyr::rename("hotspot_id" = "hotspot")

# from hotspotAbsoluteRegions_fResidue
hotspotAbsoluteRegions_fResidue$uniprot_pos <- paste(hotspotAbsoluteRegions_fResidue$protein, hotspotAbsoluteRegions_fResidue$position, sep = "_")
domainSites$in_hotspot_fResidue <- domainSites$uniprot_pos %in% hotspotAbsoluteRegions_fResidue$uniprot_pos
table(domainSites$in_hotspot_fResidue, domainSites$in_hotspot)
```


## Subset for hotspot sites
```{r}
#Make a table just for sites in hotspots
hotspotSites <- domainSites %>% subset(!is.na(hotspot_id)) %>% 
  .[, c("uniprot_pos", "uniprot", "scientific_name", "AA", "position", "hotspot_id")] %>%
  mutate(site = paste0(AA, position)) %>%
  mutate(uniprot_site = paste(uniprot, site, sep = "_"))
##Map in hotspot info
hotspotSites <- separate_rows(hotspotSites, "hotspot_id", sep = ";") #First expand to the hotspot level
hotspotSites <- left_join(hotspotSites, hotspotDf, by = "hotspot_id")
##Collapse at the individual site level
hotspotSites <- distinct(hotspotSites)
hotspotSites <- hotspotSites %>% group_by(uniprot_site) %>%
  summarise_all(~paste(unique(.), collapse = ";"), .groups = "drop")
hotspotSites <- as.data.frame(hotspotSites)
```


## Make table for domains with hotspots. This includes preparing representative structures and sequences
```{r}
#Data just for hotspot domains
hotspotDomainDf <- hotspotDf[, c("domain", "pfam", "interpro_name", "interpro_nameShort", "strRep", "strNum")] %>% distinct

#For each representative structure, extract the corresponding domain sequences
##Set up which identifier to use. This is almost always Intepro, but for one domain only the pfam works
hotspotDomainDf$queryDomainID <- hotspotDomainDf$domain
hotspotDomainDf$queryDomainID[which(hotspotDomainDf$domain == "IPR002213")] <- "PF00201"
##Run APIs
hotspotDomainDf[, c("strRep_chain", "strRep_start", "strRep_end", "strRep_seq",
                    "strRep_startAuthor", "strRep_endAuthor")] <- NA
for (i in which(!is.na(hotspotDomainDf$strRep))){
  interproID <- hotspotDomainDf$queryDomainID[i]
  PDBid <- hotspotDomainDf$strRep[i]
  
  #API call for structures with the domain
  url <- paste0("https://www.ebi.ac.uk/pdbe/api/mappings/", interproID, "?format=json")
  response <- GET(url)
  apiDat <- fromJSON(httr::content(response, as = "text"), flatten = F)
  #Select one chain
  ##Only proceed if the PDB is in the corresponding interpro
  if (PDBid %in% names(apiDat[[1]]$PDB)){
    chainDat <- apiDat[[1]]$PDB[[PDBid]][1, ]
    #Extract results
    hotspotDomainDf$strRep_chain[i] <- chainDat$chain_id[1]
    hotspotDomainDf$strRep_start[i] <- chainDat$start$residue_number[1]
    hotspotDomainDf$strRep_end[i] <- chainDat$end$residue_number[1]
    hotspotDomainDf$strRep_startAuthor[i] <- chainDat$start$author_residue_number[1]
    hotspotDomainDf$strRep_endAuthor[i] <- chainDat$end$author_residue_number[1]
    
    #API call for the structure sequence
    url <- paste0("https://www.ebi.ac.uk/pdbe/api/pdb/entry/molecules/", PDBid, "?format=json")
    response <- GET(url)
    apiDat <- fromJSON(httr::content(response, as = "text"), flatten = F)
    strDat <- apiDat[[PDBid]] %>%
      subset(grepl("polypeptide", molecule_type))
    #Loop over entities and check that they are in the right chain
    for (j in 1:nrow(strDat)){
      if (chainDat$chain_id[1] %in% strDat$in_chains[[j]]){
        hotspotDomainDf$strRep_seq[i] <- strDat$sequence[j]
      }
    }
  }
}
##Subset sequences for domains
hotspotDomainDf$strRep_seqDomain <- substr(hotspotDomainDf$strRep_seq, 
                                           start = hotspotDomainDf$strRep_start,
                                           stop = hotspotDomainDf$strRep_end)

#Export representative domain sequences for MAFFT
wRep <- hotspotDomainDf %>% 
  filter(!is.na(strRep)) %>% 
  distinct(domain, .keep_all = TRUE)  # Zinc finger has multipel hotspots but slightly different PFAM domains, so I fix this
##Prepare an identifier like that recognised by hotspot analysis
wRep$id <- paste0(wRep$domain, "__", wRep$strRep, ";", wRep$strRep_start, ";", wRep$strRep_end)
repDomainSeqs <- setNames(wRep$strRep_seqDomain, wRep$id)
writeXStringSet(BStringSet(repDomainSeqs), filepath = "output/data/hotspots/domains/representativeStrs/strRep_seqDomain.fasta",
                width = 20000)
```


# 7b. Run representative hotspot script
- Move output from output/data/hotspots/domains/representativeStrs/strRep_seqDomain.fasta to the cluster 
- Run scripts/hotspot_scripts/04_repStrsMAFFT.py to align domain sequences with the representative structure 
- Run scripts/hotspot_scripts/05_repStrsHotspots.py to run hotspot analysis with the representative structure
- Move output from output/data/hotspots/domains/representativeStrs back to local 

## Inspect representative sequence results
```{r}
#load in data
hotspotOutRep <- read_csv("output/data/hotspots/domains/representativeStrs/hotspotOutput.csv")
hotspotResRep <- read_csv("output/data/hotspots/domains/representativeStrs/hotspotResidueOutput.csv")
hotspotResRep_repOnly <- read_csv("output/data/hotspots/domains/representativeStrs/hotspotResidueOutput_strRep.csv") %>%
  as.data.frame 

#Check number of hotspots
##Set up
repHotspots <- unique(hotspotOutRep$hotspot_id) %>% sort
OGhotspots <- subset(hotspotDf, domain %in% hotspotResRep$domain)$hotspot_id %>% unique %>% sort
hotspotLst <- list("OG" = OGhotspots, "rep" = repHotspots)
hotspotLst <- map(hotspotLst, function(vec){
  data.frame("hotspot_id" = vec, "domain" = gsub("_\\d+$", "", vec))
})
##Multiplicity per domain
hotspotSummLst <- map(hotspotLst, ~group_by(., domain) %>% summarise(n = n()))
hotspotSumm <- hotspotSummLst$OG %>% dplyr::rename("n_OG" = "n") %>%
  left_join(., dplyr::rename(hotspotSummLst$rep, "n_rep" = "n"), by = "domain")
subset(hotspotSumm, is.na(n_rep) | n_rep != n_OG)

#Check coverage of hotspots
hotspotResRep_repOnlySig <- hotspotResRep_repOnly %>% subset(hotspot == TRUE)
hotspotResRep_repOnlySig$domain %>% unique %>% length
subset(hotspotResRep_repOnlySig, !is.na(position_representative))$domain %>% unique %>% length
```



# 8. Export hotspot results
```{r}
save(hotspotDf, file = "output/data/hotspots/domains/hotspotDf__3.RData")
save(hotspotDomainDf, file = "output/data/hotspots/domains/hotspotDomainDf__3.RData")
write.csv(hotspotDomainDf, file = "output/data/hotspots/domains/hotspotDomainDf__3.csv")
save(domainSites, file = "output/data/hotspots/domains/domainSites__3.RData")
save(hotspotSites, file = "output/data/hotspots/domains/hotspotSites__3.RData")
write.csv(hotspotSites, file = "output/data/hotspots/domains/hotspotSites__3.csv")
```

