# Background
Here I explore and prepare the ubn proteomics data
This encompasses site identifications across multiple species, and 
quantitative data from human

# Initialise
```{r}
library(tidyverse)
library(purrr)
library(ggplot2)
library(readxl)

home_directory <- ""
source(paste0(home_directory, "scripts/essential_visualisations_JvG.R"))
source("scripts/00_objects.R")
source("scripts/01_general.R")
```


# Load data and initial tidying
## Main data
```{r}
#Site data
ubnOmicsSites <- read_csv(file = "data/features/ubnOmics/ensp_site_canonical.csv") %>%
  dplyr::rename("publication" = "publication...3") %>% 
  dplyr::select(-c("publication...8")) %>%
  mutate(enspPos = paste0(ensp, "_", position))
## Add manual publicationID-experiment maps
pubExpMap <- read_csv("data/features/ubnOmics/manualMaps/manual_publication_experiment_map.csv")
for (i in 1:nrow(pubExpMap)){
  expID <- pubExpMap$experiment_id[i]
  ubnOmicsSites$publication[which(ubnOmicsSites$experiment == expID)] <- pubExpMap$publication[i]
}

#Ubn-omics data with quantifications
ubnOmicsQuant <- read_csv(file = "data/features/ubnOmics/quanty_ubi.csv") %>%
  dplyr::select(-c("...1", "id...2", "id...9", "peptide_id...18", "id...19", "experiment")) %>%
  dplyr::rename("peptide_id" = "peptide_id...7",
                "mystery_id" = "id...14")


#Info on conditions
conditionData <- read_csv(file = "data/features/ubnOmics/conditions.csv", col_types = cols(.default = col_character()))
conditionData$id <- as.numeric(conditionData$id)

# Worm data
wormCompsDf <- read_csv("output/data/ubnOmics/wormComps_forConditions.csv") 
wormExpCondIDs <- paste(wormCompsDf$experiment_id, wormCompsDf$condition_id, sep = "_")
```




## Tidy conditions
```{r}
#Add expCond_id
ubnOmicsQuant$expCond_id <- paste(ubnOmicsQuant$experiment_id, ubnOmicsQuant$condition_id, sep = "_")
#Subset condition data for only those in the quantitative data
ubnOmicsMeta <- conditionData %>% subset(id %in% c(ubnOmicsQuant$condition_id, wormCompsDf$condition_id)) %>% #Map in worm experiments for now
  dplyr::rename("condition_id" = "id")

# Swap conditions (treatment and control are the wrong way around)
toSwap <- which(ubnOmicsMeta$swapDecisionFinal == 1)
tempDesc <- ubnOmicsMeta$description[toSwap]
ubnOmicsMeta$description[toSwap] <- ubnOmicsMeta$control_description[toSwap]
ubnOmicsMeta$control_description[toSwap] <- tempDesc

# Make combinedDesc
ubnOmicsMeta <- ubnOmicsMeta %>%
  mutate(combinedDesc = paste0(description, "_vs_", control_description, "_time(min)_", time_min))
```


# Make dataset identifiers for all datasets
```{r}
# First make experiment metadata
experiment_metaDat <- ubnOmicsSites %>%
  group_by(publication, scientific_name, experiment) %>%
  summarise(row_count = n(), .groups = "drop") 

# Get publication info
publicationMap <- read_csv(file = "data/features/ubnOmics/publications table.csv")
##Add to metadata
experiment_metaDat <- left_join(experiment_metaDat, 
                               publicationMap[, c("ID", "pubmedID")] %>% dplyr::rename("publication" = "ID") %>%
                                 mutate(publication = as.numeric(publication)) %>% subset(!is.na(publication)), 
                               by = "publication")
experiment_metaDat$hasPMID <- !is.na(experiment_metaDat$pubmedID)
experiment_metaDat$hasPMID <- factor(experiment_metaDat$hasPMID, levels = c(T, F))
levels(experiment_metaDat$hasPMID) <- c("published", "unpublished")

# Subset down to dataset level
dataset_metaDat <- experiment_metaDat %>%
  select(-experiment) %>%
  group_by(publication) %>%
  summarise(across(everything(), ~ if (cur_column() == "row_count") sum(.x, na.rm = TRUE) else first(.x)),
            .groups = "drop")

# Make dataset ID
dataset_metaDat <- dataset_metaDat %>% 
  .[order(.$pubmedID, .$scientific_name, .$hasPMID), ] %>%
  group_by(scientific_name, hasPMID) %>% 
  mutate(seq_num = row_number()) %>% 
  ungroup() %>% 
  mutate(dataset_ID = paste(hasPMID, scientific_name, seq_num, sep = "_"))
dataset_metaDat <- dataset_metaDat %>% dplyr::select(-c("seq_num"))

# Map back into experiment_metaDat
experiment_metaDat <- left_join(experiment_metaDat, dataset_metaDat[, c("publication", "dataset_ID")], by = "publication")

# Add enrichment and acquisition mode (all are diGly and DDA)
enrAcqMeta <- read_csv("data/features/ubnOmics/publishedEnrichmentMethods.csv")
enrAcqMeta <- enrAcqMeta %>% dplyr::rename("enrichment" = "Enrichment class", "acquisition" = "Acquisition mode", "pubmedID" = "PMID")
# Add metadata for published experiments
experiment_metaDat <- left_join(experiment_metaDat,
  enrAcqMeta[, c("pubmedID", "enrichment", "acquisition")], by = "pubmedID")
# Add metadata for unpublished experiments (they are all diGly and DDA)
unpubInds <- which(experiment_metaDat$hasPMID == "unpublished")
experiment_metaDat$enrichment[unpubInds] <- "diGly antibody"
experiment_metaDat$acquisition[unpubInds] <- "DDA"

# Export
save(dataset_metaDat, file = "output/data/ubnOmics/dataset_metaDat__2.RData")
```


# Prepare quantification data
## Map site identifiers
```{r}
#Map
ubnOmicsQuantMapped <- subset(ubnOmicsQuant, site_id %in% ubnOmicsSites$site_id)  %>%
  left_join(., ubnOmicsSites[, c("site_id", "publication", "scientific_name", "enspPos")], by = "site_id")
#Inspect numbers
subset(ubnOmicsQuantMapped, !is.na(log2)) %>%
  split(.$scientific_name) %>% map(function(dat){
    #dat <- distinct(dat, site_id, .keep_all = T)
    return(length(unique(dat$enspPos)))
  })
```


## Process each dataset

Process each dataset (slow)
```{r}
ubnOmicsQuantProc_lst <- ubnOmicsQuantMapped %>% split(.$expCond_id) %>% map(function(dat){
  #Add multiplicity
  dat$multiplicity <- str_count(dat$peptide_seq, "ub")
  #Collapse information at the site level
  #This will average different peptides with the same site, and replciates within the same experiment_condition
  meanCols <- c("log2")
  maxCols <- c("localization_score", "max_spectral_count", "spectral_count")
  collapseCols <- c("peptide_id", "peptide_seq", "multiplicity")
  firstCols <- setdiff(colnames(dat), c(meanCols, maxCols, collapseCols, "enspPos"))
  summariseMax <- function(x){
    x <- x[which(!is.na(x))]
    if (length(x) == 0){
      return(NA)
    } else {
      return(max(x))
    }
  }
  datColl <- dat %>% group_by(enspPos) %>%
    summarise(
      across(all_of(meanCols), ~mean(.x, na.rm = TRUE)),
      across(all_of(maxCols), ~summariseMax(.x)),
      across(all_of(collapseCols), ~paste(na.omit(.x), collapse = ";")),
      across(all_of(firstCols), ~first(na.omit(.x)))
    ) %>% 
    as.data.frame
  return(datColl)
})
save(ubnOmicsQuantProc_lst, file = "output/data/ubnOmics/ubnOmicsQuantProc_lst__1.RData")
```

```{r}
# Checkpoint
# load("output/data/ubnOmics/ubnOmicsQuantProc_lst__1.RData")
```


## Make matrix, metadata, and identifier data
```{r}
library(data.table)
#Identifier data
ubnOmicsComp_identif <- map(ubnOmicsQuantProc_lst, ~.[, c("enspPos", "residue", "scientific_name")]) %>%
  rbindlist %>% distinct %>% as.data.frame
#Matrix
ubnOmicsComp_matrix <- matrix(NA, nrow = nrow(ubnOmicsComp_identif), ncol = length(ubnOmicsQuantProc_lst),
                              dimnames = list(ubnOmicsComp_identif$enspPos,
                                              names(ubnOmicsQuantProc_lst)))
for (i in 1:length(ubnOmicsQuantProc_lst)){
  dat <- ubnOmicsQuantProc_lst[[i]]
  ubnOmicsComp_matrix[dat$enspPos, i] <- dat$log2
}
#Metadata
expCond_ids <- c(colnames(ubnOmicsComp_matrix), wormExpCondIDs) #Manually map in worm data
ubnOmicsComp_meta <- data.frame("expCond_id" = expCond_ids) %>%
  mutate(experiment_id = strsplit(expCond_id, "_") %>% map(~.[1]) %>% unlist %>% as.numeric,
         condition_id = strsplit(expCond_id, "_") %>% map(~.[2]) %>% unlist %>% as.numeric) %>%
  left_join(., ubnOmicsMeta, by = "condition_id")  %>%
  mutate(combinedDescUID = paste0(combinedDesc, "__", expCond_id))  #Add UID that includes description
##Add species and publication
speciesPub <- ubnOmicsSites[, c("experiment", "scientific_name", "publication")] %>% distinct %>% 
  dplyr::rename("experiment_id" = "experiment")
ubnOmicsComp_meta <- left_join(ubnOmicsComp_meta, speciesPub, by = "experiment_id") 

#Prep publication info and add to metadata
publicationMap <- read_csv(file = "data/features/ubnOmics/publications table.csv")
##Add to metadata
ubnOmicsComp_meta <- left_join(ubnOmicsComp_meta, 
                               publicationMap[, c("ID", "pubmedID")] %>% dplyr::rename("publication" = "ID") %>%
                                 mutate(publication = as.numeric(publication)) %>% subset(!is.na(publication)), 
                               by = "publication")

# Add in dataset identifier
ubnOmicsComp_meta <- left_join(ubnOmicsComp_meta, dataset_metaDat[, c("publication", "dataset_ID")], by = "publication")
```



## Add in experimental detail for unpublished experiments
```{r}
# Load
experimentDetail <- read_xlsx(path = "output/data/ubnOmics/ubnOmics_unpublished_EXPDETAILS.xlsx", sheet = "ubnOmics_unpublished_forDetails") %>% 
  dplyr::rename("experiment_description" = "Cell line/Organism", "quant_type" = "quant type") %>% 
  .[, c("experiment_description", "quant_type", "experiment_id", "condition_id")] %>% 
  mutate(condition_id = as.numeric(condition_id))
# Map in
ubnOmicsComp_meta <- left_join(ubnOmicsComp_meta, experimentDetail, by = c("experiment_id", "condition_id"))
# Get rid of experimental info for published datasets
ubnOmicsComp_meta$experiment_description[which(!is.na(ubnOmicsComp_meta$pubmedID))] <- NA
ubnOmicsComp_meta$quant_type[which(!is.na(ubnOmicsComp_meta$pubmedID))] <- NA
```



## Process NAs that have been misprocessed as 0 values
```{r}
#Summary
zeroValSumm <- data.frame(
  "expCond_id" = colnames(ubnOmicsComp_matrix),
  "numQuant" = colSums(!is.na(ubnOmicsComp_matrix)),
  "numZero" = colSums(ubnOmicsComp_matrix == 0, na.rm = T)
)
zeroValSumm <- left_join(zeroValSumm, ubnOmicsComp_meta, by = "expCond_id")

zeroValConds <- subset(zeroValSumm, numZero >= 20)$expCond_id
for (cond in zeroValConds){
  ubnOmicsComp_matrix[which(ubnOmicsComp_matrix[, cond] == 0), cond] <- NA
}
```


## Merge biological replicates that are flagged separately
```{r}
#Get biological replicates
ubnOmicsComp_meta <- mutate(ubnOmicsComp_meta, expDesc = paste(experiment_id, combinedDesc, sep = "__"))
tab <- table(ubnOmicsComp_meta$expDesc)
biolRepDf <- ubnOmicsComp_meta %>% subset(expDesc %in% names(tab)[which(tab > 1)]) %>%
  .[order(.$expDesc, .$expCond_id), c("expCond_id", "expDesc", "pubmedID")] 

##Flag conditions to keep
##I retain the expCond_id identifier from the first one when sorted
biolRep_expDescs <- unique(biolRepDf$expDesc)
biolRepDf$keep <- F
for (x in biolRep_expDescs){
  inds <- which(biolRepDf$expDesc == x) %>% sort
  biolRepDf$keep[inds[1]] <- T
}
biolRepDf$grouping <- biolRepDf$expDesc # Group replica by expDesc column

#Add some conditions manually
extraCollapseConds <- list(
  c("423_1970", "426_1970", "428_1970"),
  c("427_2110", "428_2110"),
  c("448_2103", "457_2103", "446_2103"),
  c("450_2105", "451_2105")
)
## Check that they are not already being collapse
table(unlist(extraCollapseConds) %in% biolRepDf$expCond_id)
## Get df
biolRepDf_manual <- ubnOmicsComp_meta %>% 
  subset(expCond_id %in% unlist(extraCollapseConds)) %>%
  .[order(.$expDesc, .$expCond_id), c("expCond_id", "expDesc", "pubmedID")] 
## Flag conditions to keep
biolRepDf_manual$keep <- F 
for (x in extraCollapseConds){
  biolRepDf_manual$keep[which(biolRepDf_manual$expCond_id == x[1])] <- T
}
biolRepDf_manual$grouping <- map(biolRepDf_manual$expCond_id, function(x){
  boolVec <- map(extraCollapseConds, ~(x %in% .)) %>% unlist
  return(paste0("manualGroup_", which(boolVec == T)))
}) %>% unlist 

# Combine
biolRepDf <- rbind(biolRepDf, biolRepDf_manual)

#Average replicates
for (x in split(biolRepDf, biolRepDf$grouping)){
  conds <- x$expCond_id
  condKeep <- x$expCond_id[which(x$keep == T)]
  if (condKeep %in% colnames(ubnOmicsComp_matrix)){ # Error handling for conditions that are in metadata but not quant data
    ubnOmicsComp_matrix[, condKeep] <- apply(ubnOmicsComp_matrix[, conds], 1, FUN = "mean", na.rm = T)
  }
}
ubnOmicsComp_matrix <- ubnOmicsComp_matrix[, -which(colnames(ubnOmicsComp_matrix) %in% 
                                                    subset(biolRepDf, keep == F)$expCond_id)]


#Update metadata
ubnOmicsComp_meta <- subset(ubnOmicsComp_meta, expCond_id %in% subset(biolRepDf, keep == F)$expCond_id == F)
```


## QC and filtering
```{r}
#Number of quantified sites per conditions
##Focus on human
humanConds <- subset(ubnOmicsComp_meta, scientific_name == "Homo sapiens")$expCond_id
quantNumsHs <- colSums(!is.na(ubnOmicsComp_matrix[, humanConds]))
hist(quantNumsHs, breaks = 20)
sort(quantNumsHs)
```

Filtering
```{r}
# Filter out MRN experiment due to outlier behaviour and incorrect biological signal
ubnOmicsComp_meta <- subset(ubnOmicsComp_meta, experiment_id != 443)
ubnOmicsComp_matrix <- ubnOmicsComp_matrix[, intersect(colnames(ubnOmicsComp_matrix), ubnOmicsComp_meta$expCond_id)]

# Filter out arsenite experiment due to insufficient metadata
ubnOmicsComp_meta <- subset(ubnOmicsComp_meta, experiment_id != 441)
ubnOmicsComp_matrix <- ubnOmicsComp_matrix[, intersect(colnames(ubnOmicsComp_matrix), ubnOmicsComp_meta$expCond_id)]

#Filter by quantification
quantNums <- colSums(!is.na(ubnOmicsComp_matrix))
filtCols <- names(quantNums)[which(quantNums >= 500)]
##Filter meta data
ubnOmicsComp_meta_unfilt <- ubnOmicsComp_meta # Retain a dataset without quantificaiton filtering
ubnOmicsComp_meta <- subset(ubnOmicsComp_meta, expCond_id %in% filtCols) 
##Filter matrix
ubnOmicsComp_matrix <- ubnOmicsComp_matrix[, filtCols]
ubnOmicsComp_matrix <- ubnOmicsComp_matrix[which(rowSums(!is.na(ubnOmicsComp_matrix)) > 0), ] #Remove sites no longer quantified
##Filter identif data
ubnOmicsComp_identif <- subset(ubnOmicsComp_identif, enspPos %in% rownames(ubnOmicsComp_matrix))
```


# Export
```{r}
ubnOmicsComp_lst <- list(
  "meta" = ubnOmicsComp_meta,
  "identif" = ubnOmicsComp_identif,
  "matrix" = ubnOmicsComp_matrix
)
save(ubnOmicsComp_lst, file = "output/data/ubnOmics/ubnOmicsComp_lst__8.RData")

load("output/data/ubnOmics/ubnOmicsComp_lst__8.RData")
```

## Prepare and export metadata of datasets and sites
```{r}
# Start from dataset meta, because this covers all datasets
experiment_master_metaData <- experiment_metaDat %>% 
  dplyr::rename("experiment_id" = "experiment")

# Add in condition metadata where possible
experiment_master_metaData <- full_join(experiment_master_metaData, 
  ubnOmicsComp_meta_unfilt[, c("experiment_id", setdiff(colnames(ubnOmicsComp_meta_unfilt), colnames(experiment_master_metaData)))], 
  by = "experiment_id") 
experiment_master_metaData <- experiment_master_metaData[!(grepl("unpublished", experiment_master_metaData$dataset_ID) & is.na(experiment_master_metaData$experiment_description)), ]

# Save
save(experiment_master_metaData, file = "output/data/ubnOmics/experiment_master_metaData__3.RData")

# Format and export for supplementary
## Format
renameList <- c(
  "scientific_name" = "Organism",
  "description" = "Treatment_Description",
  "time_min" = "Time_Minutes",
  "control_description" = "Control_Description",
  "pubmedID" = "PubMed_ID",
  "dataset_ID" = "Dataset_ID",
  "experiment_description" = "Experiment description (unpublished data)",
  "quant_type" = "Quantification method (unpublished data)",
  "enrichment" = "Enrichment method",
  "acquisition" = "Acquisition method"
)
experiment_master_metaData_Exp <- experiment_master_metaData[, names(renameList)] %>%
  subset(scientific_name != "Homo sapiens") %>% 
  dplyr::rename(!!!setNames(names(renameList), renameList)) %>% 
  distinct 

## Order
orderDf <- strsplit(unique(experiment_master_metaData_Exp$Dataset_ID), "_") %>% 
  purrr::reduce(rbind) %>% as.data.frame
colnames(orderDf) <- c("pub", "species", "num")
orderDf <- orderDf %>% 
  mutate(num = as.numeric(num), pub = factor(pub, levels = c("unpublished", "published"))) %>% 
  .[order(.$pub, .$species, .$num), ] %>% 
  mutate(Dataset_ID = paste(pub, species, num, sep = "_"))
experiment_master_metaData_Exp$Dataset_ID <- factor(experiment_master_metaData_Exp$Dataset_ID, levels = orderDf$Dataset_ID)
experiment_master_metaData_Exp <- experiment_master_metaData_Exp[order(experiment_master_metaData_Exp$Dataset_ID), ]

## Write
library(openxlsx)
wb <- createWorkbook()
addWorksheet(wb, "All_Experiments_nonHs")
writeData(wb, "All_Experiments_nonHs", experiment_master_metaData_Exp)

saveWorkbook(wb, "output/data/ubnOmics/SuppTable_experimentMaster_meta.xlsx", overwrite = TRUE)
```


# Overview figures
Number of studies, perturbations, sites, and quantifications
```{r}
#For human
ubnOmicsComp_hs_lst <- ubnOmicsComp_lst
ubnOmicsComp_hs_lst$meta <- subset(ubnOmicsComp_hs_lst$meta, scientific_name == "Homo sapiens")
ubnOmicsComp_hs_lst$identif <- subset(ubnOmicsComp_hs_lst$identif, scientific_name == "Homo sapiens")
ubnOmicsComp_hs_lst$matrix <- ubnOmicsComp_hs_lst$matrix[ubnOmicsComp_hs_lst$identif$enspPos, 
                                                         ubnOmicsComp_hs_lst$meta$expCond_id]

##Numbers
print("Publications")
ubnOmicsComp_hs_lst$meta %>% subset(scientific_name == "Homo sapiens" & !is.na(pubmedID)) %>%
      .[, "pubmedID"] %>% unique %>% length
print("Perturbations")
nrow(ubnOmicsComp_hs_lst$meta)
print("Quantified sites")
quantM <- !is.na(ubnOmicsComp_hs_lst$matrix)
length(which(rowSums(quantM) > 0))
print("Quantifications")
print(sum(quantM))
```

Overview of number of perturbations in different groups
```{r}
# Compute fraction of total and cumulative boundaries for each slice
groupSumm <- ubnOmicsComp_meta %>% group_by(group) %>% summarise(count = n())
groupSumm <- mutate(groupSumm, group = factor(group, levels = names(omicsTreatColours))) %>% 
  .[order(.$group), ]
groupSumm <- groupSumm %>%
  mutate(fraction = count / sum(count),
         ymax = cumsum(fraction),
         ymin = lag(ymax, default = 0))

#Plot
ggplot(groupSumm, aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3, fill = group, colour = group)) +
  geom_rect(linewidth = 0.235) +
  coord_polar(theta = "y") +
  # Place group names around the outside
  geom_text(
    x = 4,  # text radius (further out)
    aes(y = (ymax + ymin) / 2, label = group),
    size = 1.875*7/5, colour = "#353535"
  ) +
  # Place the counts slightly inward
  geom_text(
    x = 3.5,  # text radius (slightly in from the outer edge)
    aes(y = (ymax + ymin) / 2, label = count),
    size = 1.875*7/5, color = "black"
  ) +
  xlim(c(2.5, 4)) + # Limit x so we only see the donut ring (3 to 4)
  comfy_theme(include_xaxis = F, include_yaxis = F) +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.text.y = element_blank()) +
  scale_fill_manual(values = alpha(omicsTreatColours, 0.8), name = "Perturbation") +
  scale_colour_manual(values = omicsTreatColours, name = "Perturbation")
ggsave_pdfpng(file = "output/images/ubnOmics/overview/pie_group_nums",
  width = 3, height = 1.75)
```