# Background
Here I analysis ubn omics data

# Initialise
```{r}
library(tidyverse)
library(purrr)
library(ggplot2)

home_directory <- ""
source(paste0(home_directory, "scripts/essential_visualisations_JvG.R"))
source(paste0(home_directory, "scripts/enrichment_analysis_JvG.R"))
source("scripts/00_objects.R")
source("scripts/01_general.R")
source("scripts/03_visualisations.R")

load("data/features/ubnOmics/ubnOmicsComp_lst__8.RData")
ubnOmicsComp_meta <- ubnOmicsComp_lst$meta
ubnOmicsComp_identif <- ubnOmicsComp_lst$identif
ubnOmicsComp_matrix <- ubnOmicsComp_lst$matrix
```

## Set up omics data
```{r}
#Subset for human
ubnOmicsComp_hs_lst <- ubnOmicsComp_lst
ubnOmicsComp_hs_lst$meta <- subset(ubnOmicsComp_hs_lst$meta, scientific_name == "Homo sapiens")
ubnOmicsComp_hs_lst$identif <- subset(ubnOmicsComp_hs_lst$identif, scientific_name == "Homo sapiens")
ubnOmicsComp_hs_lst$matrix <- ubnOmicsComp_hs_lst$matrix[ubnOmicsComp_hs_lst$identif$enspPos, 
                                                         ubnOmicsComp_hs_lst$meta$expCond_id]
write.csv(ubnOmicsComp_hs_lst$meta, file = "output/data/ubnOmics/ubnOmicsComp_hs_meta.csv",
          na = "")
```


### Map to uniprot_pos and PSP
```{r}
#Map to uniprot_pos
load("data/features/ubnOmics/ubnOmicsSites_unipMap_hs__3.RData")
ubnOmicsSites_unipMap_hs <- ubnOmicsSites_unipMap_hs %>% 
  mutate(enspPos = gsub("_[A-Z]", "_", IDpos)) %>%
  distinct(.)
if ("uniprot_pos" %in% colnames(ubnOmicsComp_hs_lst$identif) == F){
  ubnOmicsComp_hs_lst$identif <- left_join(ubnOmicsComp_hs_lst$identif,
                                           ubnOmicsSites_unipMap_hs, by = "enspPos")
}
##Filter for unique uniprot_pos
ubnOmicsComp_hs_lst$identif <- distinct(ubnOmicsComp_hs_lst$identif,
                                        uniprot_pos, .keep_all = T) %>%
  subset(!is.na(uniprot_pos))
##Assign uniprot_pos as names
ubnOmicsComp_hs_lst$matrix <- ubnOmicsComp_hs_lst$matrix[ubnOmicsComp_hs_lst$identif$enspPos, ]
rownames(ubnOmicsComp_hs_lst$matrix) <- ubnOmicsComp_hs_lst$identif$uniprot_pos
rownames(ubnOmicsComp_hs_lst$identif) <- ubnOmicsComp_hs_lst$identif$uniprot_pos

#Map to PSP
load("data/PSP/PSP_reg_ubCorr__3.RData")
PSP_reg_ubCorr <- mutate(PSP_reg_ubCorr, uniprot_pos = gsub("_[A-Z]", "_", uniprot_site))
PSP_reg_ubSub <- subset(PSP_reg_ubCorr, uniprot_pos %in% ubnOmicsComp_hs_lst$identif$uniprot_pos)
##Split for different functions
PSP_reg_ub_FUNC <- split_ON_FUNCTION(PSP_reg_ubSub) 
##Flag in omics data
protDegSites <- subset(PSP_reg_ub_FUNC, ON_FUNCTION_strp == "protein degradation")$uniprot_pos
otherFuncSites <- subset(PSP_reg_ub_FUNC, ON_FUNCTION_strp != "protein degradation")$uniprot_pos
ubnOmicsComp_hs_lst$identif$PSP_reg <- ubnOmicsComp_hs_lst$identif$uniprot_pos %in% PSP_reg_ubSub$uniprot_pos
ubnOmicsComp_hs_lst$identif$PSP_reg_ProtDeg <- ubnOmicsComp_hs_lst$identif$uniprot_pos %in% protDegSites
ubnOmicsComp_hs_lst$identif$PSP_reg_otherFunc <- ubnOmicsComp_hs_lst$identif$uniprot_pos %in% otherFuncSites
ubnOmicsComp_hs_lst$identif$PSP_reg_notProtDeg <- ubnOmicsComp_hs_lst$identif$uniprot_pos %in% setdiff(otherFuncSites, protDegSites)
ubnOmicsComp_hs_lst$identif$PSP_reg_ProtDegOnly <- ubnOmicsComp_hs_lst$identif$uniprot_pos %in% setdiff(protDegSites, otherFuncSites)
##Overview
for (col in colnames(ubnOmicsComp_hs_lst$identif)){
  if (grepl("PSP", col)){
    print(col)
    print(table(ubnOmicsComp_hs_lst$identif[, col]))
  }
}


#Map to the PRIDE build
load("data/features/ubiFeatProc__1.RData")
ubnOmicsComp_hsPRIDE_lst <- ubnOmicsComp_hs_lst
ubnOmicsComp_hsPRIDE_lst$identif <- subset(ubnOmicsComp_hsPRIDE_lst$identif,
                                           uniprot_pos %in% ubiFeatProc$uniprot_pos)
ubnOmicsComp_hsPRIDE_lst$matrix <- ubnOmicsComp_hsPRIDE_lst$matrix[ubnOmicsComp_hsPRIDE_lst$identif$uniprot_pos, ]
##Overview
for (col in colnames(ubnOmicsComp_hsPRIDE_lst$identif)){
  if (grepl("PSP", col)){
    print(col)
    print(table(ubnOmicsComp_hsPRIDE_lst$identif[, col]))
  }
}


#Export
save(ubnOmicsComp_hsPRIDE_lst, file = "output/data/ubnOmics/ubnOmicsComp_hsPRIDE_lst__8.RData")
```


### Export quantifications for a supp table
```{r}
# Pivot matrix to longer
ubnOmicsComp_hsPRIDE_matrixLong <- ubnOmicsComp_hsPRIDE_lst$matrix %>% as.data.frame %>% 
  rownames_to_column("uniprot_position") %>% pivot_longer(cols = colnames(ubnOmicsComp_hsPRIDE_lst$matrix), names_to = "condition_id", values_to = "log2FC") %>% 
  subset(!is.na(log2FC))
 
# Make cleaner experiment description
ubnOmicsComp_hsPRIDE_lst$meta <- mutate(ubnOmicsComp_hsPRIDE_lst$meta,
  combinedDescClean = paste0(description, " ", time_min, "min_vs_", control_description)
) %>% mutate(combinedDescClean = gsub(" NAmin", "", combinedDescClean))

# Export data
path <- "output/data/ubnOmics/SuppTable_panSpeciesOmics.xlsx"
library(openxlsx)
wb <- createWorkbook()

# Meta data
sheet <- "HumanConditionMeta"
renameList <- c(
  "expCond_id" = "condition_id",
  "combinedDescClean" = "Treatment_vs_Control",
  "group" = "Condition_group",
  "dataset_ID" = "Dataset_ID",
  "pubmedID" = "PubMed_ID"
)
metaExp <- ubnOmicsComp_hsPRIDE_lst$meta[, names(renameList)] %>% dplyr::rename(!!!setNames(names(renameList), renameList))
## Write
removeWorksheet(wb, sheet)
addWorksheet(wb, sheet)
writeData(wb, sheet, metaExp)
saveWorkbook(wb, path, overwrite = TRUE)

# Identif data
sheet <- "HumanSiteMeta"
renameList <- c(
  "uniprot_pos" = "uniprot_position",
  "sequence_window" = "position_+-7AA"
)
identifExp <- ubnOmicsComp_hsPRIDE_lst$identif[, names(renameList)] %>% dplyr::rename(!!!setNames(names(renameList), renameList))
## Write
removeWorksheet(wb, sheet)
addWorksheet(wb, sheet)
writeData(wb, sheet, identifExp)
saveWorkbook(wb, path, overwrite = TRUE)

# Quant data
sheet <- "HumanQuantifications"
removeWorksheet(wb, sheet)
addWorksheet(wb, sheet)
writeData(wb, sheet, ubnOmicsComp_hsPRIDE_matrixLong)
saveWorkbook(wb, path, overwrite = TRUE)
```


# Correlation analysis of conditions
## Run correlation analysis
```{r}
#Functions
library(ComplexHeatmap)
library(circlize)
# Function to make colours
makeHmapGrad <- function(midpoint = 0.5){
  breaks <- c(-1, -midpoint, 0, midpoint, 1)
  # Generate evenly spaced points between each consecutive breakpoint
  num_intervals <- 9  # Same as the original Brewer palette length
  expanded_breaks <- sort(unique(c(seq(breaks[1], breaks[2], length.out = num_intervals / 4),
                                   seq(breaks[2], breaks[3], length.out = num_intervals / 4),
                                   seq(breaks[3], breaks[4], length.out = num_intervals / 4),
                                   seq(breaks[4], breaks[5], length.out = num_intervals / 4))))
  # Select colors from the "RdBu" palette, matching the number of expanded breaks
  palette_colors <- rev(brewer.pal(length(expanded_breaks), "RdBu"))
  # Make colour ramp
  ramp <- colorRamp2(expanded_breaks, palette_colors)
  return(ramp)
}
# Function to compute correlation matrix and visualize with heatmap
correlation_heatmap <- function(mat, annotation_df, annotation_colors, midpoint = 0.5,
                                numClusters = 5) {
  # Compute correlation matrix
  cor_mat <- cor(mat, use = "pairwise.complete.obs")
  
  # Perform hierarchical clustering on the columns
  #dist_mat <- dist(t(mat))
  dist_mat <- dist(t(cor_mat)) #Do it on correlation matrix instead
  hc <- hclust(dist_mat)
  column_order <- hc$order
  
  #Get clusters
  clusters_k <- cutree(hc, k = numClusters)
  #annotation_df$cluster <- as.character(clusters_k)
  
  # Create column annotation
  ha <- HeatmapAnnotation(
    df = annotation_df,
    col = annotation_colors,
    annotation_legend_param = list(
      annotation_name_gp = gpar(fontsize = 7),
      title_gp = gpar(fontsize = 7),
      labels_gp = gpar(fontsize = 7)
    )
  )
  
  # Generate heatmap
  Heatmap(cor_mat, name = "Correlation", 
          #row_order = column_order, column_order = column_order,
          cluster_rows = T, cluster_columns = T,
          clustering_distance_rows = "pearson", clustering_distance_columns = "pearson",
          top_annotation = ha, 
          col = makeHmapGrad(midpoint = midpoint),
          show_row_names = TRUE, show_column_names = FALSE, row_names_gp = gpar(fontsize = 2),
          show_row_dend = F, show_column_dend = T,
          heatmap_legend_param = list(title_gp = gpar(fontsize = 7), labels_gp = gpar(fontsize = 7))
          )
}

#Pick data
preCorrMat <- ubnOmicsComp_hsPRIDE_lst$matrix %>% 
  .[, order(colnames(.))]

#Set up labelling of condition groups
annotation_df <- ubnOmicsComp_hsPRIDE_lst$meta %>% 
  subset(expCond_id %in% colnames(preCorrMat)) %>% .[order(.$expCond_id), ] %>% 
  dplyr::select(c("group"))
annotation_colors <- list(group = omicsTreatColours)

#Run correlation
pdf("output/images/ubnOmics/correlation/hmap_condCorr_all.pdf",
    width = 4.75, height = 3)
correlation_heatmap(mat = preCorrMat, annotation_df = annotation_df, annotation_colors = annotation_colors,
                    midpoint = 0.35, numClusters = 10)
dev.off()
```


## Correlate the same condition across publications
### Set up treatment subgroups
```{r}
# Load in grouping
metaDatSubgroups <- read_xlsx(path = "output/data/ubnOmics/metaDat_forSubgroups.xlsx", sheet = "data")
```

### Extract correlations within subgroups
```{r}
# Set up 
subgroups <- metaDatSubgroups %>% subset(!is.na(subgroup)) %>% 
  .[, c("subgroup", "publication")] %>% distinct %>% 
  group_by(subgroup) %>% summarise(num = n()) %>% 
  subset(num > 1)
subgroups <- unique(subgroups$subgroup)

# Extract correlations
cor_mat <- cor(preCorrMat, use = "pairwise.complete.obs")
cor_mat[lower.tri(cor_mat)] <- NA
cor_mat_long <- cor_mat %>% 
  as.data.frame %>% 
  rownames_to_column("row") %>% 
  pivot_longer(cols = -row, names_to = "col", values_to = "r") %>% 
  subset(!is.na(r))
cor_mat_long_subgroups <- map(subgroups, function(sg){
  metaDatGroup <- subset(metaDatSubgroups, subgroup == sg)
  cor_mat_long_sub <- subset(cor_mat_long, row %in% metaDatGroup$expCond_id & col %in% metaDatGroup$expCond_id)
  cor_mat_long_sub <- cor_mat_long_sub %>% 
    left_join(., dplyr::rename("row" = "expCond_id", "row_pub" = "publication", "row_combinedDescUID" = "combinedDescUID", metaDatGroup[, c("expCond_id", "publication", "combinedDescUID")])) %>% 
    left_join(., dplyr::rename("col" = "expCond_id", "col_pub" = "publication", "col_combinedDescUID" = "combinedDescUID", metaDatGroup[, c("expCond_id", "publication", "combinedDescUID")])) %>% 
    mutate(samePub = row_pub == col_pub, sameExp = row == col, subgroup = sg)
  return(cor_mat_long_sub)
}) %>% purrr::reduce(rbind) %>% 
  mutate(samePubLabel = factor(samePub, levels = c(F, T)))
levels(cor_mat_long_subgroups$samePubLabel) <- c("Different study", "Same study")

# visualise
## Each side by side
cor_mat_long_subgroups %>% 
  subset(!sameExp) %>%
  ggplot(aes(x = subgroup, y = r, colour = samePubLabel)) +
  geom_boxplot(position = position_dodge(width = 0.9), outlier.shape = NA, linewidth = 0.235) +
  geom_point(position = position_dodge(width = 0.9), shape = 16, size = 1, alpha = 0.7) +
  comfy_theme(rotate_x_text = T)
## Two panels
# First, compute counts per facet & subgroup
counts <- cor_mat_long_subgroups %>% 
  subset(!sameExp) %>%
  group_by(samePubLabel, subgroup) %>%
  summarise(n = n())

cor_mat_long_subgroups %>% 
  subset(!sameExp) %>%
  ggplot(aes(x = subgroup, y = r, colour = samePubLabel)) +
  geom_boxplot(outlier.shape = NA, linewidth = 0.235) +
  geom_point(shape = 16, size = 1, alpha = 0.5) +
  geom_text(data = counts, 
            aes(x = subgroup, y = 1, label = n), 
            inherit.aes = FALSE, 
            vjust = -0.5, 
            size = 1.875) +
  comfy_theme(rotate_x_text = T) +
  facet_wrap(~samePubLabel, scales = "free_x") + 
  labs(y = "Pearson's r") +
  ggtitle("Correlation of conditions")
ggsave_pdfpng(file = "output/images/ubnOmics/correlation/box_sameCond_diffStudy", width = 4.5, height = 2.5)

# Summary numbers
corrMeans <- cor_mat_long_subgroups %>% 
  subset(!sameExp) %>% 
  group_by(subgroup, samePubLabel) %>% 
  summarise(rMean = mean(r))
corrMeans %>% group_by(samePubLabel) %>% summarise(rMeanTotal = mean(rMean))
```



## Viusalise individual correlations
```{r}
prefix <- "output/images/ubnOmics/correlation/examples/corrplot_"
width = 1.75 
height = 1.75
#Set up correlations
corM <- cor(ubnOmicsComp_hsPRIDE_lst$matrix, use = "pairwise.complete.obs")
corDf <- as.data.frame(corM)
corDf$Var1 <- rownames(corDf)
corDfLong <- corDf %>%
  pivot_longer(cols = -Var1, names_to = "Var2", values_to = "correlation")

# Set up correlator function
correlator <- function(ubnOmicsLst, cond1, cond2){
  plotDat <- ubnOmicsLst$matrix[, c(cond1, cond2)] %>%
    as.data.frame %>%
    mutate(uniprot_pos = rownames(.))
    test <- cor.test(plotDat[[cond1]], plotDat[[cond2]], use = "pairwise.complete.obs")
    return(test)
}


#Pick correlations
###DNA damage vs proteasome inhib
corrplotConds(ubnOmicsLst = ubnOmicsComp_hsPRIDE_lst, cond1 = "381_2025", cond2 = "425_1973")  
ggsave_pdfpng(file = paste0(prefix, "UVprtsm"), width = width, height = height)
print(correlator(ubnOmicsComp_hsPRIDE_lst, cond1 = "381_2025", cond2 = "425_1973"))

###DNA damage (prtsm inhib bkd) vs proteasome inhib
corrplotConds(ubnOmicsLst = ubnOmicsComp_hsPRIDE_lst, cond1 = "440_2164", cond2 = "423_1970")
ggsave_pdfpng(file = paste0(prefix, "UVprtsmbkd_prtsm"), width = width, height = height)
print(correlator(ubnOmicsComp_hsPRIDE_lst, cond1 = "440_2164", cond2 = "423_1970")$p.value)

###DNA damage vs ER stress
corrplotConds(ubnOmicsLst = ubnOmicsComp_hsPRIDE_lst, cond1 = "440_2162", cond2 = "436_2136",
              sites_to_label = c("P23396_214", "P15880_275", "P15880_58"))
ggsave_pdfpng(file = paste0(prefix, "UV_ERstress"), width = width, height = height)
print(correlator(ubnOmicsComp_hsPRIDE_lst, cond1 = "440_2162", cond2 = "436_2136")$p.value)
```



# Export metadata and annotations for supplementary info
```{r}
# Load master metadata
load("data/features/ubnOmics/experiment_master_metaData__3.RData")

# Subset for quantitative conditions
metaDat <- experiment_master_metaData %>% subset(expCond_id %in% ubnOmicsComp_hsPRIDE_lst$meta$expCond_id)

# Order
x1 <- gsub("_.+$", "", metaDat$dataset_ID)
x2 <- gsub("^.+_", "", metaDat$dataset_ID) %>% as.numeric
metaDat <- metaDat[order(x1, x2), ] 

# Get summary statistics
expCond_quants <- data.frame("expCond_id" = colnames(ubnOmicsComp_hsPRIDE_lst$matrix), "numQuant" = colSums(!is.na(ubnOmicsComp_hsPRIDE_lst$matrix)))
metaDat <- left_join(metaDat, expCond_quants, by = "expCond_id")
View(metaDat)

# Export condition metadata
renameList <- c(
  "description" = "Treatment_Description",
  "time_min" = "Time_Minutes",
  "control_description" = "Control_Description",
  "group" = "Treatment_Group",
  "pubmedID" = "PubMed_ID",
  "dataset_ID" = "Dataset_ID",
  "numQuant" = "Quantified_UbiSites",
  "experiment_description" = "Experiment description (unpublished data)",
  "quant_type" = "Quantification method (unpublished data)",
  "enrichment" = "Enrichment method",
  "acquisition" = "Acquisition method"
)
metaDat_Exp <- metaDat[, names(renameList)] %>%
  dplyr::rename(!!!setNames(names(renameList), renameList))

# Write to excel for supp table
library(openxlsx)
outDir = "output/data/ubnOmics/SuppTable_experimentMaster_meta.xlsx"
sheet = "Quantitative_Conditions"
wb <- loadWorkbook(outDir)
removeWorksheet(wb, sheet)
addWorksheet(wb, sheet)
writeData(wb, sheet = sheet, x = metaDat_Exp)
saveWorkbook(wb, outDir, overwrite = TRUE)
```



# Prepare and define regulation under conditions
## Functions
```{r}
#Function: Pivot matrix to long, add metadata, and make psite groups
matrixToLongAndGrouped <- function(matrix){
  #Pivot to long
  matrixLong <- matrix %>% as.data.frame 
  pivCols <- colnames(matrixLong)
  matrixLong <- matrixLong %>%
    rownames_to_column("uniprot_pos") %>%
    pivot_longer(cols = pivCols, names_to = "expCond_id", values_to = "log2FC") %>%
    subset(!is.na(log2FC))
  #Add metadata
  matrixLong <- left_join(matrixLong, ubnOmicsComp_hsPRIDE_lst$meta, by = "expCond_id") %>%
    .[order(.$group), ] %>% 
    mutate(expCond_id = factor(expCond_id, levels = unique(expCond_id))) %>%
    mutate(combinedDescUID = factor(combinedDescUID, levels = unique(combinedDescUID))) %>%
    left_join(., ubnOmicsComp_hsPRIDE_lst$identif, by = "uniprot_pos")
  #Add groups
  matrixLongGrouped <- matrixLong %>%
  dplyr::rename("groupCond" = "group") %>%
  makeCombDat(data = ., args = list(
  "notReg" = list("PSP_reg", F),
  "reg" = list("PSP_reg", T),
  "protDeg" = list("PSP_reg_ProtDeg", T),
  "notProtDeg" = list("PSP_reg_notProtDeg", T)
))
  return(matrixLongGrouped)
}
```




## Classify regulated sites
Using top/bottom x%
```{r}
##Using top/bottom x%
fracCutoff <- 0.05
lower_cutoff <- apply(ubnOmicsComp_hsPRIDE_lst$matrix, 2, function(x) quantile(x, fracCutoff, na.rm = TRUE))
upper_cutoff <- apply(ubnOmicsComp_hsPRIDE_lst$matrix, 2, function(x) quantile(x, 1 - fracCutoff, na.rm = TRUE))
regBoolMLst <- list(
  "down" = sweep(ubnOmicsComp_hsPRIDE_lst$matrix, 2, lower_cutoff, FUN = "<"),
  "up" = sweep(ubnOmicsComp_hsPRIDE_lst$matrix, 2, upper_cutoff, FUN = ">")
)
regBoolMLst$either <- regBoolMLst$down | regBoolMLst$up
regBoolMLst <- map(regBoolMLst, function(m){
  m[is.na(m)] <- F
  return(m)
})
```

```{r}
#Assign systematically 
##Make a list of conditions to work with
condsLst <- list(
  "all" = ubnOmicsComp_hsPRIDE_lst$meta$expCond_id,
  "notPrtsm" = subset(ubnOmicsComp_hsPRIDE_lst$meta, group != "Proteasome inhibition")$expCond_id,
  "Bort" = subset(ubnOmicsComp_hsPRIDE_lst$meta, group == "Proteasome inhibition" &
                    grepl("Bortezomib", combinedDesc))$expCond_id,
  "BortMG" = subset(ubnOmicsComp_hsPRIDE_lst$meta, group == "Proteasome inhibition" &
                    (grepl("Bortezomib", description) |
                       grepl("MG-132", description)))$expCond_id
)
for (grp in unique(ubnOmicsComp_hsPRIDE_lst$meta$group)){
  conds <- subset(ubnOmicsComp_hsPRIDE_lst$meta, group == grp)$expCond_id
 condsLst[[grp]] <- conds
}
##Get # regulated and fraction regulated
###Loop over conditions
for (i in 1:length(condsLst)){
  name <- names(condsLst)[i]
  quantNumCol <- paste0("numQuant_", name)
  ubnOmicsComp_hsPRIDE_lst$identif[, quantNumCol] <- rowSums(!is.na(ubnOmicsComp_hsPRIDE_lst$matrix[ubnOmicsComp_hsPRIDE_lst$identif$uniprot_pos, 
                                                  intersect(colnames(regBoolMLst[[1]]), condsLst[[i]])]))
  ###Loop over directions
  for (j in 1:length(regBoolMLst)){
    suffix <- paste0("_", names(regBoolMLst)[j])
    if (suffix == "_either"){suffix <- ""}
    boolM <- regBoolMLst[[j]]
    regNumCol <- paste0("numReg_", name, suffix)
    
    regFracCol <- paste0("fracReg_", name, suffix)
    ubnOmicsComp_hsPRIDE_lst$identif[, regNumCol] <- rowSums(boolM[ubnOmicsComp_hsPRIDE_lst$identif$uniprot_pos, 
                                                                  intersect(colnames(boolM), condsLst[[i]])])
    ubnOmicsComp_hsPRIDE_lst$identif[, regFracCol] <- ubnOmicsComp_hsPRIDE_lst$identif[, regNumCol] / ubnOmicsComp_hsPRIDE_lst$identif[, quantNumCol] * 100 #Make percentage
    ubnOmicsComp_hsPRIDE_lst$identif[which(is.na(ubnOmicsComp_hsPRIDE_lst$identif[, regFracCol])), regFracCol] <- 0
  }
}
```


## Proteasome inhibitor conditions - Aggregate distributions
I normalise and aggregate the FC distributions within each proteasome inhibitor type
```{r}
# Get doses of epoxomicin in nM
ubnOmicsComp_hsPRIDE_lst$meta$epoxDose <- sapply(ubnOmicsComp_hsPRIDE_lst$meta$description, function(desc) {
  if (grepl("^Epoxomicin", desc)) {
    m <- regexpr("^Epoxomicin\\s+(\\d+(?:\\.\\d+)?)(nm|um)", desc, perl = TRUE)
    if(m[1] != -1) {
      dose <- as.numeric(sub("^Epoxomicin\\s+(\\d+(?:\\.\\d+)?)(nm|um).*", "\\1", desc, perl = TRUE))
      unit <- sub("^Epoxomicin\\s+(\\d+(?:\\.\\d+)?)(nm|um).*", "\\2", desc, perl = TRUE)
      if(unit == "um") dose <- dose * 1000
      return(dose)
    }
  }
  return(NA)
})

#Set up conditions and treatment types
conds <- subset(ubnOmicsComp_hsPRIDE_lst$meta, group == "Proteasome inhibition")$expCond_id
treatTypes <- list(
  "Bortezomib" = subset(ubnOmicsComp_hsPRIDE_lst$meta, group == "Proteasome inhibition" &
                    grepl("Bortezomib", combinedDesc) & 
                    !grepl("MLN", combinedDesc))$expCond_id,  # Remove the MLN combo condition because it is not pure Bort treatment
  "MG-132" = subset(ubnOmicsComp_hsPRIDE_lst$meta, group == "Proteasome inhibition" &
                    grepl("MG-132", description) )$expCond_id,
  "Epoxomicin" = subset(ubnOmicsComp_hsPRIDE_lst$meta, group == "Proteasome inhibition" &
                    epoxDose > 10 &   # Only use high doses of epoxomicin, because lower doses do not have proteome-wide effects
                    grepl("Epoxomicin", description) 
                    )$expCond_id
)

subset(ubnOmicsComp_hsPRIDE_lst$meta, group == "Proteasome inhibition") %>% View


#Get matrix and transform
matrixConds <- ubnOmicsComp_hsPRIDE_lst$matrix[, conds] %>% as.data.frame 
##Quantile-transform
normalize_median_iqr <- function(mat) {
  # Compute medians and IQRs per column
  col_medians <- apply(mat, 2, median, na.rm = TRUE)
  col_iqrs <- apply(mat, 2, function(x) IQR(x, na.rm = TRUE))

  # Compute the target median and IQR
  target_median <- median(col_medians, na.rm = TRUE)
  target_iqr <- median(col_iqrs, na.rm = TRUE)

  # Normalize each column
  normalized_mat <- sweep(mat, 2, col_medians, FUN = "-")  # Center by median
  normalized_mat <- sweep(normalized_mat, 2, col_iqrs / target_iqr, FUN = "/") # Scale by IQR
  normalized_mat <- sweep(normalized_mat, 2, target_median, FUN = "+")  # Shift to target median

  return(normalized_mat)
}
matrixConds <- normalize_median_iqr(matrixConds)

#Pivot to long
matrixLongConds <- matrixToLongAndGrouped(matrixConds)
#Group by treatment types
matrixLongConds$treatType <- NA
for (i in 1:length(treatTypes)){
  matrixLongConds$treatType[which(matrixLongConds$expCond_id %in% treatTypes[[i]])] <- names(treatTypes)[i]
}
matrixLongConds <- subset(matrixLongConds, !is.na(treatType))
matrixLongConds$treatType <- factor(matrixLongConds$treatType, levels = names(treatTypes))
#Average
condDatAvg <- matrixLongConds %>%
  group_by(uniprot_pos, group, treatType) %>% 
  summarise(log2FC = mean(log2FC, na.rm = T))
prtsmFCagg <-   condDatAvg
prtsmFCaggWide <- prtsmFCagg %>% .[, c("uniprot_pos", "treatType", "log2FC")] %>% distinct %>%
  mutate(treatType = paste(treatType, "meanlog2FC", sep = "_")) %>%
  pivot_wider(names_from = "treatType", values_from = "log2FC")
#Add to data
## Remove if already 
removeCols <- setdiff(colnames(prtsmFCaggWide), "uniprot_pos")
ubnOmicsComp_hsPRIDE_lst$identif <- ubnOmicsComp_hsPRIDE_lst$identif %>% .[, which(colnames(.) %in% removeCols == F)]
ubnOmicsComp_hsPRIDE_lst$identif <- left_join(ubnOmicsComp_hsPRIDE_lst$identif, prtsmFCaggWide, by = "uniprot_pos")

#Plot
condDatAvg %>%
  subset(group != "reg") %>%
  subset(abs(log2FC) <= 5) %>% 
ViolBoxplot(., vars = list(x = "treatType", y = "log2FC", colour = "group", fill = "group")) +
  geom_hline(yintercept = 0, size = 0.235, alpha = 0.3, linetype = "dashed") +
  scale_colour_manual(name = "Site group",
                    labels = c("Unannotated", "Protein deg.", "Other functions"),
                    values = PSPcoloursLine) +
  scale_fill_manual(name = "Site group",
                    labels = c("Unannotated", "Protein deg.", "Other functions"),
                    values = alpha(PSPcoloursLine, 0.05)) +
  labs(y = "Avg. log2FC (inhibitor/control)") + theme(axis.title.x = element_blank()) +
  ggtitle("Proteasome inhibition")
ggsave_pdfpng(file = "output/images/ubnOmics/PSP/prtsmInhib/viol_avgFC_PSPreg",
              width = 3.5, height = 2)

#Plot distributions to check normalisation
matrixLongConds %>%
  subset(group != "reg") %>%
ViolBoxplot(., vars = list(x = "expCond_id", y = "log2FC", colour = "group"),
            violScale = "width") +
  geom_hline(yintercept = c(-1, 1), size = 0.235, alpha = 0.5, linetype = "dashed") +
  theme(axis.text.x = element_text(colour = "black",
                                    size = 7,
                                    hjust = 1, vjust = 0.5, angle = 90)) 
```

# Export modified datasets for conservation analysis
```{r}
#Export omics data with PSP annotations and proteasome inhibitor averages
save(ubnOmicsComp_hsPRIDE_lst, file = "output/data/ubnOmics/ubnOmicsComp_hsPRIDE_lst__forCons__8.RData")
```



# Export features for model training
```{r}
#Get regulation features
ubnOmicsFeaturesAll <- ubnOmicsComp_hsPRIDE_lst$identif

#Export
save(ubnOmicsFeaturesAll, file = "output/data/ubnOmics/ubnOmicsFeauresAll__4.RData")
```






















