# Background
Here I compare ubn omics data to conservation data
This is a modular notebook containing analyses of how ubiquitin modification rates relate to evolutionary conservation, 
following initial data preparation in 08_ubnOmicsAnalysis.Rmd.

# Initialise
```{r}
library(tidyverse)
library(purrr)
library(ggplot2)
library(ComplexHeatmap)
library(circlize)
library(readxl)
library(openxlsx)
library(limma)

home_directory <- ""
source(paste0(home_directory, "scripts/essential_visualisations_JvG.R"))
source(paste0(home_directory, "scripts/enrichment_analysis_JvG.R"))
source("scripts/00_objects.R")
source("scripts/01_general.R")
source("scripts/03_visualisations.R")
source("scripts/17_linear_models.R")
```

## Load dependencies from 08_ubnOmicsAnalysis and elsewhere
```{r}
#Load processed omics data with PSP annotations and proteasome inhibitor averages
load("output/data/ubnOmics/ubnOmicsComp_hsPRIDE_lst__forCons__8.RData")

#Load conservation features
load("data/features/ubiFeatProc__1.RData")
```

## Functions
```{r}
#Function: Heatmap of box enrichment results
hmap_boxEnr <- function(data, val = "nlog10p_sign", yaxisName = "combinedDescUID", maxVal = NA,
                        cluster = FALSE, order_column = NA,
                        ...){
  ##Cap values
  if (!is.na(maxVal)){
    data[which(data[, val] > maxVal), val] <- maxVal
    data[which(data[, val] < -maxVal), val] <- -maxVal
  }
  ##Prep data
  plotDatWide <- data[, c("box", yaxisName, "padj", val)] %>%
    pivot_longer(cols = c("padj", val)) %>%
    mutate(boxName = paste(box, name, sep = "_")) %>%
    dplyr::select(-c("box", "name")) %>%
    pivot_wider(names_from = "boxName", values_from = "value")
  ##Get columns
  fc_cols <- colnames(plotDatWide) %>% .[grep(val, .)]
  p_cols <- colnames(plotDatWide) %>% .[grep("padj$", .)]
  xnames <- strsplit(fc_cols, "_") %>% map(~.[1]) %>% unlist
  ##Assign y-axis name
  plotDatWide <- as.data.frame(plotDatWide)
  rownames(plotDatWide) <- plotDatWide[, yaxisName]
  ##Order column
  if (is.na(order_column)){order_column <- yaxisName}
  ##Cluster
  if (cluster){
    clust_cols <- fc_cols
  } else {
    clust_cols <- NULL
  }
  ##Plot
  plt <- fc_sig_heatmap(plotDatWide, 
                 fc_cols_w_pvals = fc_cols,
                 pval_cols = p_cols,
                 x_axis_names = xnames,
                 order_column = order_column,
                 yaxis_naming = "keepall",
                 clust_cols = clust_cols,
                 ...)
  return(plt)
}

# Function: Calculate median of logFCs across several factors and plot
hmap_FCmed <- function(
  dat,
  logFCcol = "log2FC",
  xCol = "consBox_LysCont",
  yCol = "protAbundance_log_bins",
  facetCol = "treatType",
  maxFC = NULL,
  showNum = FALSE
  ){
  # Summarise data
  summDat <- dat %>% 
    .[rowSums(is.na(.[, c(logFCcol, xCol, yCol)])) == 0,  ] %>% 
    group_by(!!sym(xCol), !!sym(yCol), !!sym(facetCol)) %>% 
    summarise(log2FC_med = median(!!sym(logFCcol)), num = n()) %>% 
    ungroup
  summDat$facetCol <- summDat[, facetCol]
  # Cap FC values if desired
  if (!is.null(maxFC)){
    summDat$log2FC_med[which(summDat$log2FC_med >= maxFC)] <- maxFC
    summDat$log2FC_med[which(summDat$log2FC_med <= -maxFC)] <- -maxFC
  }
  # Plot
  maxVal <- max(abs(summDat$log2FC_med), na.rm = T)
  plt <- summDat %>% 
    ggplot(., aes(x = !!sym(xCol), y = !!sym(yCol), fill = log2FC_med)) + 
    geom_tile() + 
    scale_fill_gradientn(colors = rev(brewer.pal(9, "RdBu")), limits = c(-maxVal, maxVal)) +
    coord_fixed() + comfy_theme(rotate_x_text = T, include_xaxis = F, include_yaxis = F) +
    facet_wrap(facetCol)
  if (showNum){
    plt <- plt + geom_text(aes(label = num), colour = "#7a7a7a", size = 1.875 * 3 / 5)
  }
  return(plt)
}

# Function: Plot box plots for FC data
# xCol and colourCol are factors that are combined for the x-axis
# Colour is done in blocks, so colourCol specifies the groups and xCol specifies the order within groups
box_FCs <- function(
  dat,
  logFCcol = "log2FC",
  xCol = "consBox_LysCont",
  colourCol = "protAbundance_log_binsLab",
  yLims = c(-2, 3),
  hlineInts = c(-1, 0, 1)
){
  # Prep data
  dat <- dat %>% .[rowSums(is.na(.[, c(logFCcol, xCol, colourCol)])) == 0,  ] %>% 
    as.data.frame
  ## Combine columns to make one x-axis
  dat <- dat %>% .[order(.[, colourCol], .[, xCol]), ] %>% 
    mutate(xFact = paste(!!sym(colourCol), !!sym(xCol), sep = "__")) %>% 
    mutate(xFact = factor(xFact, levels = unique(xFact)))

  ## Calculate numbers to show above violins
  dataNums <- dat %>% 
    group_by(xFact) %>%
    summarise(num = n(), .groups = "drop")
  ## Calculate y value for number annotations
  dataNums$y <- max(yLims) * 0.9
  
  # Plot
  plt <- ggplot(dat, aes(x = xFact, y = !!sym(logFCcol), color = !!sym(colourCol), fill = !!sym(colourCol))) + 
    geom_boxplot(outlier.shape = NA, size = 0.235, alpha = 0.2, coef = 0, width = 0.7) + 
    geom_hline(yintercept = hlineInts, size = 0.235, linetype = "dashed", alpha = 0.5) +
    geom_text(data = dataNums, aes(x = xFact, y = y, label = num),
                          size = 1.875 * 3 / 5, colour = "black",
                          position = position_dodge2(width = 0.9),
                          inherit.aes = F) +
    comfy_theme(rotate_x_text = T) + 
    theme(axis.text.x = element_text(colour = "black",size = 5, hjust = 1, vjust = 1, angle = 45),
          axis.title.x = element_blank()) +
    lims(y = yLims)
  return(plt)
}

# Function: Plot mean +- IQR for FC data
# xCol and colourCol are factors that are combined for the x-axis
# Colour is done in blocks, so colourCol specifies the groups and xCol specifies the order within groups
meanIQR_FCs <- function(
  dat,
  logFCcol = "log2FC",
  xCol = "consBox_LysCont",
  colourCol = "protAbundance_log_binsLab",
  yLims = c(-2, 3),
  hlineInts = c(-1, 0, 1),
  pointSize = 2.5, errorBarLinewidth = 0.235*2, errorBarWidth = 0
){
  # Prep data
  dat <- dat %>% .[rowSums(is.na(.[, c(logFCcol, xCol, colourCol)])) == 0,  ] %>%
    as.data.frame
  ## Combine columns to make one x-axis
  dat <- dat %>% .[order(.[, colourCol], .[, xCol]), ] %>%
    mutate(xFact = paste(!!sym(colourCol), !!sym(xCol), sep = "__")) %>%
    mutate(xFact = factor(xFact, levels = unique(xFact)))

  ## Calculate mean, IQR and numbers to show per group
  datSummary <- dat %>%
    group_by(xFact, !!sym(colourCol)) %>%
    summarise(mean_y = mean(!!sym(logFCcol), na.rm = TRUE),
              Q1 = quantile(!!sym(logFCcol), 0.25, na.rm = TRUE),
              Q3 = quantile(!!sym(logFCcol), 0.75, na.rm = TRUE),
              num = n(), .groups = "drop")
  ## Calculate y value for number annotations
  datSummary$y <- max(yLims) * 0.9

  # Plot
  plt <- ggplot(datSummary, aes(x = xFact, y = mean_y, color = !!sym(colourCol))) +
    geom_errorbar(aes(ymin = Q1, ymax = Q3), width = errorBarWidth, linewidth = errorBarLinewidth, alpha = 0.8) +
    geom_point(size = pointSize) +
    geom_hline(yintercept = hlineInts, size = 0.235, linetype = "dashed", alpha = 0.5) +
    geom_text(data = datSummary, aes(x = xFact, y = y, label = num),
                          size = 1.875 * 3 / 5, colour = "black",
                          position = position_dodge2(width = 0.9),
                          inherit.aes = F) +
    comfy_theme(rotate_x_text = T) +
    theme(axis.text.x = element_text(colour = "black",size = 5, hjust = 1, vjust = 1, angle = 45),
          axis.title.x = element_blank()) +
    lims(y = yLims)
  return(plt)
}
```

# Regulation vs conservation
## Visualise regulation of each condition at each conservation level
```{r}
# Prepare data by pivoting to long and adding conservation
matrixLong <- ubnOmicsComp_hsPRIDE_lst$matrix %>% as.data.frame 
  pivCols <- colnames(matrixLong)
matrixLong <- matrixLong %>%
    rownames_to_column("uniprot_pos") %>%
    pivot_longer(cols = pivCols, names_to = "expCond_id", values_to = "log2FC") %>%
    subset(!is.na(log2FC))
  #Add metadata
matrixLong <- left_join(matrixLong, ubnOmicsComp_hsPRIDE_lst$meta, by = "expCond_id") %>%
    .[order(.$group), ] %>% 
    mutate(expCond_id = factor(expCond_id, levels = unique(expCond_id))) %>%
    mutate(combinedDescUID = factor(combinedDescUID, levels = unique(combinedDescUID))) %>%
    left_join(., ubnOmicsComp_hsPRIDE_lst$identif, by = "uniprot_pos")
# Add conservation levels
matrixLong <- left_join(matrixLong, ubiFeatProc[, c("uniprot_pos", "consBox_LysCont")])
matrixLong <- subset(matrixLong, !is.na(consBox_LysCont))
```

```{r}
#Plot distributions
##Set up colours
violCols <- c("Level1" = "#2472b6",
              "Level2" = "#2472b6",
              "Level3" = "#2472b6",
              "Level4" = "#2472b6",
              "Level5" = "#2472b6")
###Mean +- IQR
for (expCond_idSel in unique(matrixLong$expCond_id)){
    desc <- as.character(subset(ubnOmicsComp_hsPRIDE_lst$meta, expCond_id == expCond_idSel)$description)
    group <- as.character(subset(ubnOmicsComp_hsPRIDE_lst$meta, expCond_id == expCond_idSel)$group)

    matrixLong %>% 
        subset(expCond_id == expCond_idSel) %>%
    MeanIQRPlot(., vars = list(x = "consBox_LysCont", y = "log2FC", colour = "consBox_LysCont"), includeLegend = F, pointSize = 3) +
            geom_hline(yintercept = 0, linewidth = 0.235, linetype = "dashed") +
    scale_colour_manual(name = "Site group",
                        values = violCols) +
    labs(y = "log2 fold change (mean ± sd)") + theme(axis.title.x = element_blank()) +
    ggtitle(desc)
    ggsave_pdfpng(file = paste0("output/images/ubnOmics/regVsCons/allCondsPlots/viol_cons_vs_", group, "_", expCond_idSel),
                    width = 2.1, height = 2.1)
}
```


## Enrichment of regulation at conservation levels 
### Gene set test
```{r}
#Set up
quantM <- !is.na(ubnOmicsComp_hsPRIDE_lst$matrix)
boxes <- as.list(paste0("Level", 1:5))
names(boxes) <- unlist(boxes)
enrResMat <- matrix(NA, nrow = length(boxes), ncol = ncol(ubnOmicsComp_hsPRIDE_lst$matrix),
                 dimnames = list(names(boxes), colnames(ubnOmicsComp_hsPRIDE_lst$matrix)))
gstResDf <- enrResMat %>% as.data.frame %>% rownames_to_column("box") %>%
  pivot_longer(cols = colnames(.)[-1], names_to = "expCond_id", values_to = "pval") %>%
  as.data.frame
gstResDf$logFC_mean_diff <- NA
#Run enrichment
for (i in 1:nrow(gstResDf)){
  box <- gstResDf$box[i]
  id <- gstResDf$expCond_id[i]
  #Get data
  FCvec <- ubnOmicsComp_hsPRIDE_lst$matrix[, id] %>% .[!is.na(.)]
  boxSites <- subset(ubiFeatProc, consBox_LysCont %in% box)$uniprot_pos
  boxSitesSub <- intersect(boxSites, names(FCvec))
  #Run overrep
  gstRes <- geneSetTest(index = boxSitesSub, statistics = FCvec, alternative = "either", type = "t")
  logFC_mean_diff <- mean(FCvec[boxSitesSub]) - mean(FCvec)
  gstResDf$pval[i] <- gstRes
  gstResDf$logFC_mean_diff[i] <- logFC_mean_diff
}
#Adjust p-values
gstResDf$padj <- p.adjust(gstResDf$pval, method = "fdr")
#Add values
gstResDf$box <- factor(gstResDf$box, levels = names(boxes))
gstResDf$nlog10p <- -log10(gstResDf$padj)
gstResDf$sign <- sign(gstResDf$logFC_mean_diff)
gstResDf$nlog10p_sign <- gstResDf$nlog10p * gstResDf$sign

#Add metadata
gstResDf <- left_join(gstResDf, ubnOmicsComp_hsPRIDE_lst$meta, by = "expCond_id")
```


### Visualise
```{r}
#Plot
sigConds <- subset(gstResDf, padj < 0.05)$expCond_id %>% unique
plotDat <- subset(gstResDf, expCond_id %in% sigConds)

hmap_boxEnr(data = plotDat,
            val = "nlog10p_sign",
            yaxisName = "combinedDescUID", maxVal = 10,
            scale_colours_manually = F, mid_colour_threshold = 3,
            legend_title = "-log10 adj. p (signed)",
            cluster = T)
ggsave_pdfpng(file = "output/images/ubnOmics/regVsCons/hmap_boxGST_allsig",
              width = 10, height = 10)


#Aggregate results at the level of groups
geomean <- function(x){exp(mean(log(x)))}
gstResDfGroupsumm <- gstResDf %>% group_by(box, group) %>%
  summarise(nlog10p_sign = mean(nlog10p_sign),
            padj = geomean(padj)) %>%
  as.data.frame
hmap_boxEnr(data = gstResDfGroupsumm,
            val = "nlog10p_sign",
            yaxisName = "group", maxVal = 10,
            scale_colours_manually = T, mid_colour_threshold = 3,
            legend_title = "-log10 adj. p (signed)",
            order_column = "Level5_nlog10p_sign") +
  theme(axis.title.y = element_text(colour = "black", size = 7)) + labs(y = "Condition group") +
  ggtitle("Averaged within groups")
ggsave_pdfpng(file = "output/images/ubnOmics/regVsCons/hmap_boxGST_groupMean",
              width = 3.75, height = 3)

# Just significant (in a consistent direction)
gstResDfGroupsumm %>% 
  subset(group %in% unique(group[abs(nlog10p_sign) > -log10(0.05)])) %>%
hmap_boxEnr(data = .,
            val = "nlog10p_sign",
            yaxisName = "group", maxVal = 10,
            scale_colours_manually = T, mid_colour_threshold = 3,
            legend_title = "-log10 adj. p (signed)",
            order_column = "Level5_nlog10p_sign") +
  theme(axis.title.y = element_text(colour = "black", size = 7)) + labs(y = "Condition group") 
ggsave_pdfpng(file = "output/images/ubnOmics/regVsCons/hmap_boxGST_groupMean_sig",
              width = 3.35, height = 2.1)
```


## Proteasome inhibition
### Prepare data
This is the normalised and aggregated proteasome inhibition data combined with conservation scores
```{r}
#Prep data
prtsmInhibCols <- c("Bortezomib_meanlog2FC", "MG-132_meanlog2FC", "Epoxomicin_meanlog2FC")
prtsmInhibDat <- ubnOmicsComp_hsPRIDE_lst$identif %>%
  pivot_longer(cols = prtsmInhibCols,
               names_to = "treatType", values_to = "log2FC") %>%
  mutate(treatType = factor(treatType, levels = prtsmInhibCols)) %>% 
  left_join(., distinct(ubiFeatProc[, c("uniprot_pos", "consBox_LysCont")]), by = "uniprot_pos")  %>% 
  mutate(dummy = T) #Dummy column for all sites
levels(prtsmInhibDat$treatType) <- gsub("_meanlog2FC", "", levels(prtsmInhibDat$treatType))
prtsmInhibDatComb <- makeCombDat(data = prtsmInhibDat, args = list(
  "Level1" = list("consBox_LysCont", "Level1"),
  "Level2" = list("consBox_LysCont", "Level2"),
  "Level3" = list("consBox_LysCont", "Level3"),
  "Level4" = list("consBox_LysCont", "Level4"),
  "Level5" = list("consBox_LysCont", "Level5")
  )
)
```


### Plot distributions
```{r}
#Plot distributions
##Set up colours
violCols <- c("Level1" = "#2472b6",
              "Level2" = "#2472b6",
              "Level3" = "#2472b6",
              "Level4" = "#2472b6",
              "Level5" = "#2472b6")
##Plot
###Violin
plotLst <- prtsmInhibDatComb %>% split(.$treatType) %>% map(function(dat){
  treat <- dat$treatType[1]
  dat %>%
  subset(abs(log2FC) <= 5) %>% 
ViolBoxplot(., vars = list(x = "group", y = "log2FC", colour = "group", fill = "group"), 
            includeLegend = F) +
  geom_hline(yintercept = 0, size = 0.235, alpha = 0.3, linetype = "dashed") +
  scale_colour_manual(name = "Site group",
                    values = violCols) +
  scale_fill_manual(name = "Site group",
                    values = alpha(violCols, 0.15)) +
  labs(y = "Avg. log2FC (vs control)") + theme(axis.title.x = element_blank()) +
  ggtitle(treat)
})
ggarrange(plotlist = plotLst, nrow = 1)
ggsave_pdfpng(file = "output/images/ubnOmics/regVsCons/viol_prtmsInhib_avgFC_Box",
            width = 6, height = 2)
###Mean +- IQR
plotLst <- prtsmInhibDatComb %>% split(.$treatType) %>% map(function(dat){
  treat <- dat$treatType[1]
  dat %>%
  MeanIQRPlot(., vars = list(x = "group", y = "log2FC", colour = "group"), includeLegend = F, pointSize = 3) +
        geom_hline(yintercept = 0, linewidth = 0.235, linetype = "dashed") +
  scale_colour_manual(name = "Site group",
                    values = violCols) +
  labs(y = "log2 fold change (mean ± sd)") + theme(axis.title.x = element_blank()) +
  ggtitle(treat)
})
ggarrange(plotlist = plotLst, nrow = 1)
ggsave_pdfpng(file = "output/images/ubnOmics/regVsCons/meanIQR_prtmsInhib_avgFC_Box",
            width = 5, height = 1.9)
```



# Comparing conservation and regulation to protein/site abundance
These are potential confounders, so I inspect how they influence protein and site abundance
## Prep data
```{r}
load("data/features/abundDat__1.RData")
abundDat$uniprot_pos <- gsub("_K", "_", rownames(abundDat))
abundFeatsLog <- colnames(abundDat) %>% .[grep("log$", .)]
prtsmInhibDat <- left_join(prtsmInhibDat, abundDat, by = "uniprot_pos")
```

## Compare conservation to abundance features
```{r}
pltLst <- list()
k <- 1
treat <- unique(prtsmInhibDat$treatType)[1]
for (abundVar in abundFeatsLog){
  subDat <- subset(prtsmInhibDat, treatType == treat & !is.na(consBox_LysCont))
  plt <- ViolBoxplot(subDat, vars = list(x = "consBox_LysCont", colour = 'consBox_LysCont', y = abundVar)) 
  pltLst[[k]] <- plt 
  k <- k + 1
}
ggarrange(plotlist = pltLst, nrow = 1, ncol = 3)
ggsave_pdfpng(file = "output/images/ubnOmics/regVsCons/abundance/viol_abundFeats_consBox", width = 8, height = 2)
```

## Compare proteasomal regulation to abundance features
```{r}
# Set up
library(viridis)
colScale <- c(seq(0, 0.25, length.out = 6 + 1)[-1], 
            seq(0.25, 0.7, length.out = 5 + 1)[-1], 1) 
hexScale <- scale_fill_gradientn(colours = viridis(10), 
                        values = rescale(colScale, c(0, 1)), 
                        limits = c(1, NA), name = "# sites")

# Plot
hexPltLst <- list()
pltLst <- list()
k <- 1 
for (treat in unique(prtsmInhibDat$treatType)){
  hexPltLst[[treat]] <- list()
  for (abundVar in abundFeatsLog){
    subDat <- subset(prtsmInhibDat, treatType == treat & !is.na(log2FC)) %>% 
    subset(abs(log2FC) < 10) %>%
      .[!is.na(.[, abundVar]), ] %>% 
      as.data.frame
    plt <- ggplot(subDat, aes(y = log2FC, x = !!sym(abundVar))) +
              geom_hex(bins = 50) +
              geom_hline(yintercept = 0, colour = "black", size = 0.235, alpha = 0.5) +
              geom_smooth(method = "lm", formula = y ~ splines::ns(x, 3), se = TRUE, colour = "red", fill = "red", alpha = 0.15, size = 0.235) +
              comfy_theme() + hexScale + ggtitle(treat) 
    pltLst[[k]] <- plt 
    hexPltLst[[treat]][[abundVar]] <- plt
    k <- k + 1
  }
}
ggarrange(plotlist = pltLst, nrow = length(abundFeatsLog), ncol = length(abundFeatsLog))
ggsave_pdfpng(file = "output/images/ubnOmics/regVsCons/abundance/hex_abundFeats_prtsmi", width = 8, height = 6)
```

Plot with abundance or site quantiles as lines
```{r}
for (var in abundFeatsLog){
  quants <- quantile(prtsmInhibDat[, var], seq(0, 10, len = 11) / 10, na.rm = T)
  plt <- hexPltLst[["MG-132"]][[var]] + 
    geom_vline(xintercept = quants, size = 0.235, colour = "red", alpha = 0.5)
  plot(plt)
  ggsave_pdfpng(file = paste0("output/images/ubnOmics/regVsCons/abundance/hex_", var,  "_wQuants_MG-132"), width = 3, height = 2)
}
```

## Compare conservation to proteasomal regulation, controling for protein abundance
```{r}
# Bin protein abundance
binN <- 10
prtsmInhibDat$protAbundance_log_bins <- cut_number(prtsmInhibDat$protAbundance_log, binN)
prtsmInhibDat$protAbundance_log_binsLab <- prtsmInhibDat$protAbundance_log_bins
levels(prtsmInhibDat$protAbundance_log_binsLab) <- paste0("Abun", 1:binN)

#Boxplot
# Split by treatments
lst <- prtsmInhibDat %>% split(.$treatType) %>% map(~subset(., !is.na(protAbundance_log_binsLab)))
pltLst <- map2(lst, names(lst), function(dat, treat){
  plt <- box_FCs(
                dat = dat,
                logFCcol = "log2FC",
                xCol = "consBox_LysCont",
                colourCol = "protAbundance_log_binsLab",
                yLims = c(-2, 3),
                hlineInts = c(-1, 0, 1)) + 
      ggtitle(treat)
  return(plt)
})
ggarrange(plotlist = pltLst, nrow = 3)
ggsave_pdfpng(file = "output/images/ubnOmics/regVsCons/abundance/box_abundBin_consVsPrtsmi", width = 0.6*binN, height = 2*3)

# Heatmap of medians
hmap_FCmed(
  dat = prtsmInhibDat,
  logFCcol = "log2FC",
  xCol = "consBox_LysCont",
  yCol = "protAbundance_log_binsLab",
  facetCol = "treatType",
  showNum = T
  )
ggsave_pdfpng(file = "output/images/ubnOmics/regVsCons/abundance/hmap_abundBin_consVsPrtsmi", width = 6, height = 4)
```


### Inspect above and below the top x %
```{r}
quantSel <- 0.1
prtsmInhibDat$topAbund <- prtsmInhibDat$protAbundance_log >= quantile(prtsmInhibDat$protAbundance_log, 1 - quantSel, na.rm = T)
prtsmInhibDat$topAbund <- factor(prtsmInhibDat$topAbund, levels = c(F, T), labels = c(paste0("0BelowTop", 100*quantSel), paste0("1AboveTop", 100*quantSel)))

#Boxplot
# Split by treatments
lst <- prtsmInhibDat %>% subset(!is.na(topAbund)) %>% split(.$treatType)
pltLst <- map2(lst, names(lst), function(dat, treat){
  plt <- box_FCs(
                dat = dat,
                logFCcol = "log2FC",
                xCol = "consBox_LysCont",
                colourCol = "topAbund",
                yLims = c(-2, 3),
                hlineInts = c(-1, 0, 1)) + 
      ggtitle(treat)
  return(plt)
})
ggarrange(plotlist = pltLst, nrow = 3)
ggsave_pdfpng(file = "output/images/ubnOmics/regVsCons/abundance/box_abundTop10_consVsPrtsmi", width = 3, height = 2*3)

# Mean +- sd plot
pltLst <- map2(lst, names(lst), function(dat, treat){
  plt <- meanIQR_FCs(
                dat = dat,
                logFCcol = "log2FC",
                xCol = "consBox_LysCont",
                colourCol = "topAbund",
                yLims = c(-2, 3),
                hlineInts = c(-1, 0, 1)) + 
      ggtitle(treat)
  return(plt)
})
ggarrange(plotlist = pltLst, nrow = 3)
ggsave_pdfpng(file = "output/images/ubnOmics/regVsCons/abundance/meanIQR_abundTop10_consVsPrtsmi", width = 3, height = 2*3)
```



#### Plot mean +- SD plots
```{r}
# Prep data
prtsmInhibDatComb <- makeCombDat(data = prtsmInhibDat, args = list(
  "Level1" = list("consBox_LysCont", "Level1"),
  "Level2" = list("consBox_LysCont", "Level2"),
  "Level3" = list("consBox_LysCont", "Level3"),
  "Level4" = list("consBox_LysCont", "Level4"),
  "Level5" = list("consBox_LysCont", "Level5")
  )
)

#Plot distributions
##Set up colours
violCols <- c("Level1" = "#2472b6",
              "Level2" = "#2472b6",
              "Level3" = "#2472b6",
              "Level4" = "#2472b6",
              "Level5" = "#2472b6")
##Plot
###Mean +- IQR
plotLst <- prtsmInhibDatComb %>% split(.$treatType) %>% map(function(dat){
  treat <- dat$treatType[1]
  dat %>%
  subset(topAbund == "0BelowTop10") %>%  # Subset for bottom 90% by protein abundance
  MeanIQRPlot(., vars = list(x = "group", y = "log2FC", colour = "group"), includeLegend = F, pointSize = 3) +
        geom_hline(yintercept = 0, linewidth = 0.235, linetype = "dashed") +
  scale_colour_manual(name = "Site group",
                    values = violCols) +
  labs(y = "log2 fold change (mean ± sd)") + theme(axis.title.x = element_blank()) +
  ggtitle(treat)
})
ggarrange(plotlist = plotLst, nrow = 1)
ggsave_pdfpng(file = "output/images/ubnOmics/regVsCons/meanIQR_prtmsInhib_avgFC_Box_bottom90Abun",
            width = 5, height = 1.9)
```


## Compare conservation to proteasomal regulation, controling for site abundance
```{r}
# Bin site abundance
binN <- 10
prtsmInhibDat$PSMcounts_log_bins <- cut_number(prtsmInhibDat$PSMcounts_log, binN)
prtsmInhibDat$PSMcounts_log_binsLab <- prtsmInhibDat$PSMcounts_log_bins
levels(prtsmInhibDat$PSMcounts_log_binsLab) <- paste0("PSM", 1:binN)


#Boxplot
# Split by treatments
lst <- prtsmInhibDat %>% split(.$treatType)
pltLst <- map2(lst, names(lst), function(dat, treat){
  plt <- box_FCs(
                dat = dat,
                logFCcol = "log2FC",
                xCol = "consBox_LysCont",
                colourCol = "PSMcounts_log_binsLab",
                yLims = c(-2, 3),
                hlineInts = c(-1, 0, 1)) + 
      ggtitle(treat)
  return(plt)
})
ggarrange(plotlist = pltLst, nrow = 3)
ggsave_pdfpng(file = "output/images/ubnOmics/regVsCons/abundance/box_sitePSMBin_consVsPrtsmi", width = 0.6*binN, height = 2*3)

# Heatmap of medians
hmap_FCmed(
  dat = prtsmInhibDat,
  logFCcol = "log2FC",
  xCol = "consBox_LysCont",
  yCol = "PSMcounts_log_binsLab",
  facetCol = "treatType",
  showNum = T
  )
ggsave_pdfpng(file = "output/images/ubnOmics/regVsCons/abundance/hmap_sitePSMBin_consVsPrtsmi", width = 6, height = 4)

```


## Linear models to quantify the influence of confounders
### Main: test protein abundance or site abundance as cofounders
```{r}
# Run linear models for abundance and PSM counts, additive or interaction, linear and nonlinear


coefDatLst <- list()
j <- 1
# Either all data or just bottom 90% by protein abundance
datLst <- list("allData" = prtsmInhibDat, "bottom90Abund" = subset(prtsmInhibDat, topAbund == "0BelowTop10"))
for (datName in names(datLst)){
  dat <- datLst[[datName]]
  for (confCol in c("protAbundance_log", "PSMcounts_log")){
    for (treat in unique(prtsmInhibDat$treatType)){
      # Set up dat
      subDat <- dat %>%
        subset(treatType == treat) %>%
        subset(!is.na(consBox_LysCont) & !is.na(log2FC))
      subDat$confCol <- subDat[[confCol]]
      subDat$confCol <- scale(subDat$confCol)
      ## Set base level
      baseLev <- "Level2"
      levs <- sort(unique(subDat$consBox_LysCont))
      subDat$consBox_LysCont <- factor(subDat$consBox_LysCont, levels = c(baseLev, setdiff(levs, baseLev)))

      #Linear effects
      lmBase <- lm(log2FC ~ consBox_LysCont, data = subDat)
      lmAdd <- lm(log2FC ~ consBox_LysCont + confCol, data = subDat)
      lmInt <- lm(log2FC ~ consBox_LysCont * confCol, data = subDat)
      #Non-linear effects
      gamBase <- gam(log2FC ~ consBox_LysCont, data = subDat)
      gamAdd <- gam(log2FC ~ consBox_LysCont + s(confCol), data = subDat)
      gamInt <- gam(log2FC ~ consBox_LysCont + s(confCol, by = consBox_LysCont), 
                    data = subDat)


      # Combine all coefficient tables into one dataframe
      coefDat <- format_model_coefs(list(lmBase = lmBase, lmAdd = lmAdd, lmInt = lmInt,
                                          gamBase = gamBase, gamAdd = gamAdd, gamInt = gamInt))
      coefDat$treatType <- treat
      coefDat$confCol <- confCol
      coefDat$data <- datName
      coefDatLst[[j]] <- coefDat
      j <- j + 1
    }
  }
}
coefDatComb <- purrr::reduce(coefDatLst, rbind)
View(coefDatComb)
```

Export results
```{r}
prepModelResults_prtmsi <- function(x){
  x %>% subset(model %in% c("gamBase", "gamInt") & termType == "main effect") %>% 
    dplyr::relocate(c("treatType", "confCol")) %>% 
    dplyr::select(-c("data", "termType")) %>% 
    mutate(treatType = factor(treatType, levels = c("MG-132", "Bortezomib", "Epoxomicin"))) %>% 
    .[order(.$treatType, .$confCol), ] %>% 
    mutate(confCol = ifelse(model == "gamBase", "", confCol)) %>% 
    mutate(confCol = factor(confCol, levels = c("protAbundance_log", "PSMcounts_log"), labels = c("log10 protein abundance", "log10 peptide PSM counts"))) %>%
    mutate(model = factor(model, levels = c("gamBase", "gamInt"), labels = c("Base (no confounder)", "Interaction")))  %>%  
    mutate(term = gsub("consBox_LysCont", "", term)) %>% 
    distinct %>% 
    dplyr::rename(
      "Treatment" = "treatType",
      "Confounder" = "confCol",
      "Main effect term" = "term",
      "Term estimate" = "Estimate"
    )
}
# Base or interaction models for all data
# Format for audience consumption
coefDat_mainExp <- subset(coefDatComb, data == "allData") %>% prepModelResults_prtmsi
# Export
outDir = "output/data/ubnOmics/regVsCons/SuppTable_consPrstmi_abund_linMods.xlsx"
sheet = "generalised_additive_models"
if (file.exists(outDir)){
  wb <- loadWorkbook(outDir)
} else {
  wb <- createWorkbook(outDir)
}
removeWorksheet(wb, sheet)
addWorksheet(wb, sheet)
writeData(wb, sheet = sheet, x = coefDat_mainExp)
saveWorkbook(wb, outDir, overwrite = TRUE)


# Base or interaction models on bottom 90% by protein abudance
coefDat_bottom90Exp <- subset(coefDatComb, data == "bottom90Abund") %>% prepModelResults_prtmsi
# Export
outDir = "output/data/ubnOmics/regVsCons/SuppTable_consPrstmi_abund_linMods_bottom90protabund.xlsx"
sheet = "generalised_additive_models"
if (file.exists(outDir)){
  wb <- loadWorkbook(outDir)
} else {
  wb <- createWorkbook(outDir)
}
removeWorksheet(wb, sheet)
addWorksheet(wb, sheet)
writeData(wb, sheet = sheet, x = coefDat_bottom90Exp)
saveWorkbook(wb, outDir, overwrite = TRUE)
```