# Background
Here I train models to predict regulatory Ubn sites. 
These form the ubi-site positional importance score

# Initialise
```{r}
library(tidyverse)
library(purrr)
library(ggplot2)
library(caret)
library(openxlsx)

home_directory <- ""
source(paste0(home_directory, "scripts/essential_visualisations_JvG.R"))
source(paste0(home_directory, "scripts/caret_JvG.R"))
source("scripts/00_objects.R")
source("scripts/01_general.R")
source("scripts/02_model_training.R")

#Load data
load("data/features/ubiFeatProc__1.RData")
ubiFeatProc <- as.data.frame(ubiFeatProc)
rownames(ubiFeatProc) <- ubiFeatProc$uniprot_site
```



# Extra processing of features
## Tidy features from ubiFeatProc
```{r}
ubiFeatProc$PRIDE_gold <- ubiFeatProc$PRIDE_set == "Gold"

ubiFeatProc$FinalBoxProcNum <- factor(ubiFeatProc$consBox_LysCont)
levels(ubiFeatProc$FinalBoxProcNum) <- gsub("Level", "", levels(ubiFeatProc$FinalBoxProcNum))
ubiFeatProc$FinalBoxProcNum <- as.numeric(as.character(ubiFeatProc$FinalBoxProcNum))
table(ubiFeatProc$FinalBoxProcNum)

ubiFeatProc$NSP_p_q3_H <- ubiFeatProc$`NSP_p[q3_H]`
ubiFeatProc$NSP_p_q3_C <- ubiFeatProc$`NSP_p[q3_C]`
```


## Add omics features
```{r}
#Load
load("data/features/ubnOmics/ubnOmicsFeauresAll__4.RData")


#Preapre mean of Bort and MG-132
ubnOmicsFeaturesAll$BortMG_meanlog2FC <- apply(ubnOmicsFeaturesAll[, c("Bortezomib_meanlog2FC", "MG-132_meanlog2FC")],
                                               1, FUN = "mean", na.rm = T)
##Cap the bottom and top 0.5% of values at the respective quantile, to avoid extreme values
FCs <- ubnOmicsFeaturesAll$BortMG_meanlog2FC 
FCPercs <- percent_rank(FCs)*100
extremeRangeCutoff <- 0.5
FCs[which(FCPercs < extremeRangeCutoff)] <- FCs[which.min(abs(FCPercs - extremeRangeCutoff))]
FCs[which(FCPercs > 100 - extremeRangeCutoff)] <- FCs[which.min(abs(FCPercs - (100 - extremeRangeCutoff)))]
hist(FCs, breaks = 20)
ubnOmicsFeaturesAll$BortMG_meanlog2FC <- FCs

#Add features
omicsFeatures <- c(
  "numQuant_all",
  "numReg_notPrtsm", "fracReg_notPrtsm",
  "BortMG_meanlog2FC"
)
ubiFeatProc <- left_join(ubiFeatProc, 
                           ubnOmicsFeaturesAll[, c("uniprot_pos", omicsFeatures)],
                           by = "uniprot_pos") %>% as.data.frame
rownames(ubiFeatProc) <- ubiFeatProc$uniprot_site

#Deal with NAs when it is straightforward
ubiFeatProc$numQuant_all[which(is.na(ubiFeatProc$numQuant_all))] <- 0
ubiFeatProc$numReg_notPrtsm[which(is.na(ubiFeatProc$numReg_notPrtsm))] <- 0
ubiFeatProc$fracReg_notPrtsm[which(is.na(ubiFeatProc$fracReg_notPrtsm))] <- 0
```




## Add hotspots 
```{r}
load("data/hotspots/hotspotSites__3.RData")
ubiFeatProc$HotspotNew_bool <- ubiFeatProc$uniprot_site %in% subset(hotspotSites, !is.na(hotspot_id))$uniprot_site
```


## Add adjacent PTMs in different windows
This is generated by running scripts/15_PSPadjRes.R
```{r}
load("data/features/PSPwindRes.RData")
ubiFeatProc <- left_join(ubiFeatProc, PSPwindRes, by = "uniprot_site") %>% as.data.frame
rownames(ubiFeatProc) <- ubiFeatProc$uniprot_site
```

# Select and prepare features
## Select features
Here you can select more features than you want to train on, just so that they are processed
```{r}
#All features to use
selFeats <- c(
  "PRIDE_gold", #One-hot for PRIDE set being gold
  
  "FinalBoxProcNum", #Conservation box as integers
  
  "HotspotNew_bool", #Whether site is in a hotspot, usign my re-calculated hotspots
  
  "mutR_am_pathogenicity_all", #AlphaMissense pathogenicity score for K->R mutation. Uses my implementation of AlphaMissense, which is more comprehensive
  
  "interface_label", #Whether site is at an AF2 multimer interface

  "NSP_p_q3_H", #Predicted probability of alpha helix, from NetSurfP3.0
  "NSP_p_q3_C", #Predicted probability of coil, from NetSurfP3.0
  "NSP_rsa", #Predicted releative solvent accessibility, from NetSurfP3.0
  "NSP_disorder", #Predicted disorder, from NetSurfP3.0
  
  "PSP_boolOtherLysMod", #One-hot for whether the Lys has a non-ubn modification in PhosphositePlus
  "PSP_numAdj_5", #Number of PTMs in a 5-residue window
  
  "numQuant_all", #Number of quantifications in omics data
  "numReg_notPrtsm", #Number of regulations in non-proteasome inhibition data
  
  "UNIPROT_in_region", #One-hot for whether the Lys is in a uniprot "region"
  "UNIPROT_in_domain" #One-hot for whether the Lys is in a uniprot "domain"
  

 
)


#Features to log-transform
logFeats <- c(
  "mutR_am_pathogenicity_all",


  "PSP_numAdj_5",
  
  "numQuant_all",
  
  "numReg_notPrtsm"

  
)


#Create feature data
datFeats <- ubiFeatProc[, selFeats]
setdiff(selFeats, colnames(ubiFeatProc))
```




## Deal with missing values
```{r}
#Inspect NA values
colSums(is.na(datFeats))

#Process conservation 
##Impute mean
boxMean <- datFeats$FinalBoxProcNum %>% .[which(. != 0)] %>% mean
datFeats$FinalBoxProcNum[which(is.na(datFeats$FinalBoxProcNum))] <- boxMean

#Filter out rows with missing values
datFeats <- datFeats[which(rowSums(is.na(datFeats)) == 0), ]

# Save data without processing
datFeats_noProc <- datFeats
```


## Process features
```{r}
#Binarise boolean features
for (col in colnames(datFeats)){
  if (is_logical(datFeats[, col])){
    datFeats[, col] <- 1*datFeats[, col]
  }
}

#Log-transform
for (col in logFeats){
  if (sum(datFeats[, col] == 0, na.rm = TRUE) > 0){
    datFeats[, col] <- log(1 + datFeats[, col]) #Avoid NA for 0 values
  } else {
    datFeats[, col] <- log(datFeats[, col])
  }
}

#Center and scale
func <- preProcess(datFeats, method = c("center", "scale"))
datFeats <- predict(func, datFeats)

# export
save(datFeats, file = "output/data/functional_scores/datFeats__7.RData")
```


## QC on features
```{r}
#Distributions
datFeats %>% mutate("uniprot_site" = rownames(.)) %>% 
      pivot_longer(cols = colnames(.)[-ncol(.)], names_to = "feature", values_to = "val") %>% 
      ggplot(., aes(x = val)) +
      geom_histogram() +
      comfy_theme() +
      facet_wrap(~feature, scales = "free")  
```

## Select only the features used for training
```{r}
selFeatsFinal <- c(
  "PRIDE_gold", 
  
  "FinalBoxProcNum",
  
  "HotspotNew_bool", 
  
  "mutR_am_pathogenicity_all", 
  
  "interface_label", 

  "NSP_p_q3_H", 
  "NSP_p_q3_C", 
  "NSP_rsa", 
  "NSP_disorder", 
  
  "PSP_boolOtherLysMod", 
  "PSP_numAdj_5", 
  
  "numQuant_all", 
  "numReg_notPrtsm", 
  
  "UNIPROT_in_region", 
  "UNIPROT_in_domain"
  

 
)
```

## Prep response
```{r}
respCol <- "PSP_reg"

#Set up as factor
resp <- ubiFeatProc[rownames(datFeats), respCol]
resp <- factor(make.names(1*resp), levels = c("X1", "X0"))

#Combine with features
datFeatsResp <- datFeats
datFeatsResp$response <- resp
```


## Export PSP annotations for supplementary table 
```{r}
# Load PSP info
load("data/PSP/PSP_reg_ubCorr__3.RData")

# Subset for regulatory sites only
regSiteDat <- ubiFeatProc[rownames(datFeats), ] %>%  
subset(PSP_reg_ProtDeg == T | PSP_reg_notProtDeg == T) %>%   
  rownames_to_column("ubi-site") %>%  
  .[, c("ubi-site", "PSP_reg_ProtDeg", "PSP_reg_notProtDeg")] %>% 
  # Add PSP info 
  left_join(., PSP_reg_ubCorr[, c("uniprot_site", "ON_FUNCTION", "ON_PROCESS", "ON_PROT_INTERACT", "ON_OTHER_INTERACT")] %>%  
                dplyr::rename("ubi-site" = "uniprot_site"), 
                by = "ubi-site") %>% 
  # Rename
  dplyr::rename("regulates protein degradation" = "PSP_reg_ProtDeg",
                "regulates non-degradation functions" = "PSP_reg_notProtDeg") %>%  
  .[order(.$`ubi-site`), ]
# Export
outDir = "output/data/functional_scores/SuppTable_degNonDeg.xlsx"
sheet = "regulatory_annotations"
if (file.exists(outDir)){
  wb <- loadWorkbook(outDir)
} else {
  wb <- createWorkbook(outDir)
}
removeWorksheet(wb, sheet)
addWorksheet(wb, sheet)
writeData(wb, sheet = sheet, x = regSiteDat)
saveWorkbook(wb, outDir, overwrite = TRUE)
```

# Train models
## Split protein degradation from other functions
```{r}
PSPsiteLst <- list(
  "nonReg" = subset(ubiFeatProc, PSP_reg == F),
  "reg" = subset(ubiFeatProc, PSP_reg == T),
  "protDeg" = subset(ubiFeatProc, PSP_reg_ProtDeg == T),
  "protDegOnly" = subset(ubiFeatProc, PSP_reg_ProtDegOnly == T),
  "notProtDeg" = subset(ubiFeatProc, PSP_reg_notProtDeg == T),
  "otherFunc" = subset(ubiFeatProc, PSP_reg_otherFunc == T)
) %>% map(~.$uniprot_site)
PSPsiteLst_datFeatsResp <- map(PSPsiteLst, ~datFeatsResp[intersect(rownames(datFeatsResp), .), ])
#Numbers
map(PSPsiteLst_datFeatsResp, ~table(.$response))

datFeatsResp_funcSpltLst <- list(
  "protDeg" = rbind(PSPsiteLst_datFeatsResp$nonReg, PSPsiteLst_datFeatsResp$protDeg),
  "protDegOnly" = rbind(PSPsiteLst_datFeatsResp$nonReg, PSPsiteLst_datFeatsResp$protDegOnly),
  "notProtDeg" = rbind(PSPsiteLst_datFeatsResp$nonReg, PSPsiteLst_datFeatsResp$notProtDeg),
  "otherFunc" = rbind(PSPsiteLst_datFeatsResp$nonReg, PSPsiteLst_datFeatsResp$otherFunc)
)

#Remove P13010, because it has many sites and biases the analysis
datFeatsResp_funcSpltLst <- map(datFeatsResp_funcSpltLst, ~.[!grepl("P13010", rownames(.)), ])

#Set up arguments for training regimes
trainArgs <- list(
  "protDeg" = list(trainDat = datFeatsResp_funcSpltLst$protDeg),
  "protDegOnly" = list(trainDat = datFeatsResp_funcSpltLst$protDegOnly),
  "notProtDeg" = list(trainDat = datFeatsResp_funcSpltLst$notProtDeg),
  "otherFunc" = list(trainDat = datFeatsResp_funcSpltLst$otherFunc)
)
```


## Pre-score screening: Train different model architectures
### Set up data 
```{r}
suffix <- "attempt11_revisions"

#Set up data (just non-prot deg sites for now)
args <- trainArgs$notProtDeg
trainDat_preSplit <- args$trainDat
save(trainDat_preSplit, file = "output/data/functional_scores/trainDat_preSplit__7.RData")
trainDat_preSplit <- trainDat_preSplit[, c(selFeatsFinal, "response")] # Select final features

#Parameters
subsampSeed <- 66 
subsampNum <- 5 
subsampMult <- 10
reorder <- TRUE 
seed <- 80
```

### Split train-test 
```{r}
#Use protein stratification
trainTestSplits <- makeFoldsProtstrat(seed = 123, k = 5, data = trainDat_preSplit)
```

### Train
```{r}
#Loop over train-test splits
mods_trainTest <- list()
for (j in 1:length(trainTestSplits)){
  splt <- trainTestSplits[[j]]
  trainDat <- splt$trainDat
  mods <- list()
  #logreg
  mods$logreg <- subsample_train(feats = trainDat[, -which(colnames(trainDat) == "response")], 
                        resp = trainDat$response, 
                        subsampSeed = subsampSeed + j, #Make sure I use a different seed for each split
                        subsampNum = subsampNum,
                        subsampMult = subsampMult,
                        reorder = reorder,
                        modFunc = modTrain_LogReg_unweighted,
                        fitControl = fitControl_CV,
                        seed = seed + j)
  #GBM
  mods$gbm <- subsample_train(feats = trainDat[, -which(colnames(trainDat) == "response")], 
                        resp = trainDat$response, 
                        subsampSeed = subsampSeed + j,
                        subsampNum = subsampNum,
                        subsampMult = subsampMult,
                        reorder = reorder,
                        modFunc = modTrain_GBM_unweighted,
                        fitControl = fitControl_CV,
                        seed = seed + j)

  #RF
  mods$RF <- subsample_train(feats = trainDat[, -which(colnames(trainDat) == "response")], 
                        resp = trainDat$response, 
                        subsampSeed = subsampSeed + j,
                        subsampNum = subsampNum,
                        subsampMult = subsampMult,
                        reorder = reorder,
                        modFunc = modTrain_RF_unweighted,
                        fitControl = fitControl_CV,
                        seed = seed + j)
  mods_trainTest[[j]] <- mods
}
names(mods_trainTest) <- names(trainTestSplits)
##Transpose
mods_trainTest_trans <- purrr::transpose(mods_trainTest)
save(mods_trainTest_trans, file = paste0("output/data/functional_scores/mods_trainTest_trans__", suffix, ".RData"))
# Checkpoint
#load(paste0("output/data/functional_scores/mods_trainTest_trans__", suffix, ".RData"))

#Plot subsampling and CV results for one split
plotROC_fromMods(mods = mods_trainTest[[1]])

#Apply to test set
##Make predictions
modsPreds_trainTest <- map(mods_trainTest_trans, function(mods){ #Loop over models
      modsPreds <- map2(mods, names(mods), function(mod, fold){ #Loop over train-test splits
        testDat <- trainTestSplits[[fold]]$testDat
        predDat <- map_dfr(mod$Model_subsamps, function(tempMod){ #Loop over subsampled models
              predictions <- predict(tempMod, newdata = testDat[, -which(colnames(testDat) == "response")], type = "prob")
              predDf <- data.frame("X1" = predictions[, "X1"], "obs" = testDat$response)
            }, .id = "subsamp")
        predDat$fold <- fold 
        return(predDat)
      })
      modsPredsDf <- purrr::reduce(modsPreds, rbind) %>% 
          mutate(Resample = paste(subsamp, fold, sep = "__"))
      return(modsPredsDf)
})
modsPreds_trainTest_df <- map_dfr(modsPreds_trainTest, function(x){x}, .id = "model")
save(modsPreds_trainTest_df, file = paste0("output/data/functional_scores/modsPreds_trainTest_df__", suffix, ".RData"))
subset(modsPreds_trainTest_df, obs == "X1")$Resample %>% table

##Calculate AUCs
modsPredsAUCs <- map_dfr(modsPreds_trainTest, function(predsDf){
  AUCs <- predsDf %>% split(.$Resample) %>% 
    map(~ROCAUC_fromPredictor(data = ., obsCol = "obs", predCol = "X1")) %>% unlist
  data.frame("Resample" = names(AUCs), "AUC" = AUCs)
}, .id = "model")
###Plot
auc_summary <- modsPredsAUCs %>%
  group_by(model) %>%
  summarize(meanAUC = mean(AUC, na.rm = TRUE),
            sdAUC = sd(AUC, na.rm = TRUE)) %>% 
  mutate(model = factor(model, levels = names(mods_trainTest[[1]])))
ggplot(auc_summary, aes(x = model, y = meanAUC, fill = model)) +
  geom_col() +
  geom_errorbar(aes(ymin = meanAUC - sdAUC, ymax = meanAUC + sdAUC), width = 0.2, linewidth = 0.235) +
  labs(title = "Mean AUC by Model", x = "Model", y = "Mean AUC") +
  comfy_theme(include_xaxis = F, rotate_x_text = T)
ggsave_pdfpng(file = paste0("output/images/functional_scores/repeatedTrainTest/bplotAUC_trainTest_", suffix), 
              width = 2, height = 2)

##Plot ROC
ROCdat <- prep_ROC_meansd_fromLst(modsPreds_trainTest)
plotROC_fromData(data = ROCdat, 
                 MeanSd = T, linewidth = 0.235*1.5) +
                 ggtitle("Repeated test sets")
ggsave_pdfpng(file = paste0("output/images/functional_scores/repeatedTrainTest/ROC_trainTest_", suffix), 
              width = 2.5, height = 2)
```

#### Make predictions on all data
```{r}
finalModel <- mods_trainTest_trans$logreg
allPreds_trainTest_lst <- map2(finalModel, names(finalModel), function(mod, fold){ #Loop over train-test splits
  predDat <- map_dfr(mod$Model_subsamps, function(tempMod){ #Loop over subsampled models
        predictions <- predict(tempMod, newdata = datFeats, type = "prob")
        predDf <- data.frame("X1" = predictions[, "X1"], "uniprot_site" = rownames(datFeats))
      }, .id = "subsamp")
  predDat$fold <- fold 
  return(predDat)
})
allPreds_trainTest_df <- purrr::reduce(allPreds_trainTest_lst, rbind) %>% 
    mutate(Resample = paste(subsamp, fold, sep = "__"))
save(allPreds_trainTest_df, file = paste0("output/data/functional_scores/allPreds_trainTest_df__", suffix, ".RData"))
```


## Train final ubi-site score
### Set up data 
```{r}
#Set up data (just non-prot deg sites)
args <- trainArgs$notProtDeg
trainDat_preSplit <- args$trainDat

#Parameters
seed <- 80
```

### Split train-test 
```{r}
#Use protein stratification, with 5-time repeated 5-fold split
numReps <- 5
trainTestforPredSplits <- list()
for (j in 1:numReps){
  splts <- makeFoldsProtstrat(seed = 123 + j, k = 5, data = trainDat_preSplit)
  names(splts) <- paste0("Rep", j, names(splts))
  trainTestforPredSplits <- c(trainTestforPredSplits, splts)
} 
save(trainTestforPredSplits, file = paste0("output/data/functional_scores/trainTestforPredSplits__", suffix, ".RData"))
#load(paste0("output/data/functional_scores/trainTestforPredSplits__", suffix, ".RData"))
```

Check splits
```{r}
# Check numbers of sites in each
splitTestNums <- data.frame(
  "Resample" = names(trainTestforPredSplits), 
  "numTest" = map(trainTestforPredSplits, ~nrow(.$testDat)) %>% unlist)
```



### Train
```{r}
#Loop over train-test splits
start <- Sys.time()
mods_trainTestforPred <- list()
for (j in 1:length(trainTestforPredSplits)){
  splt <- trainTestforPredSplits[[j]]
  trainDat <- splt$trainDat
  mods <- list()
  #logreg
  mods$logreg <- modTrain_LogReg_weighted(feats = trainDat[, -which(colnames(trainDat) == "response")], 
                                          resp = trainDat$response, 
                                          fitControl = fitControl_CV, seed = seed + j,
                                          weightFunc = caseWeights_balanced)
    mods_trainTestforPred[[j]] <- mods
}
names(mods_trainTestforPred) <- names(trainTestforPredSplits)
end <- Sys.time()
timing <- end - start
print(timing)
##Transpose
mods_trainTestforPred_trans <- purrr::transpose(mods_trainTestforPred)
save(mods_trainTestforPred_trans, file = paste0("output/data/functional_scores/mods_trainTestforPred_trans__", suffix, ".RData"))
# Checkpoint
#load(paste0("output/data/functional_scores/mods_trainTestforPred_trans__", suffix, ".RData"))

#Plot subsampling and CV results for one split
plotROC_fromMods(mods = mods_trainTestforPred[[1]])

#Apply to test set
##Make predictions
modsPreds_trainTestforPred <- map(mods_trainTestforPred_trans, function(mods){ #Loop over models
      modsPreds <- map2(mods, names(mods), function(tempMod, fold){ #Loop over train-test splits
        testDat <- trainTestforPredSplits[[fold]]$testDat
        predictions <- predict(tempMod, newdata = testDat[, -which(colnames(testDat) == "response")], type = "prob")
        predDf <- data.frame("X1" = predictions[, "X1"], "obs" = testDat$response)
        predDf$fold <- fold
        predDf$uniprot_site <- rownames(testDat)
        return(predDf)
      })
      modsPredsDf <- purrr::reduce(modsPreds, rbind) %>% 
          mutate(Resample = fold)
      return(modsPredsDf)
})
modsPreds_trainTestforPred_df <- map_dfr(modsPreds_trainTestforPred, function(x){x}, .id = "model")
## Export
save(modsPreds_trainTestforPred_df, file = paste0("output/data/functional_scores/modsPreds_trainTestforPred_df__", suffix, ".RData"))
# Checkpoint
#load(paste0("output/data/functional_scores/modsPreds_trainTestforPred_df__", suffix, ".RData"))
modsPreds_trainTestforPred <- modsPreds_trainTestforPred_df %>% split(.$model)

##Calculate AUCs
modsPredsAUCs <- map_dfr(modsPreds_trainTestforPred, function(predsDf){
  AUCs <- predsDf %>% split(.$Resample) %>% 
    map(~ROCAUC_fromPredictor(data = ., obsCol = "obs", predCol = "X1")) %>% unlist
  data.frame("Resample" = names(AUCs), "AUC" = AUCs)
}, .id = "model")
###Plot
auc_summary <- modsPredsAUCs %>%
  group_by(model) %>%
  summarize(meanAUC = mean(AUC, na.rm = TRUE),
            sdAUC = sd(AUC, na.rm = TRUE)) %>% 
  mutate(model = factor(model, levels = unique(model)))
ggplot(auc_summary, aes(x = model, y = meanAUC, fill = model)) +
  geom_col() +
  geom_errorbar(aes(ymin = meanAUC - sdAUC, ymax = meanAUC + sdAUC), width = 0.2, linewidth = 0.235) +
  labs(title = "Mean AUC by Model", x = "Model", y = "Mean AUC") +
  comfy_theme(include_xaxis = F, rotate_x_text = T)
ggsave_pdfpng(file = paste0("output/images/functional_scores/repeatedtrainTestforPred/bplotAUC_trainTestforPred_", suffix), 
              width = 2, height = 2)

##Plot ROC
ROCdat <- prep_ROC_meansd_fromLst(modsPreds_trainTestforPred)
plotROC_fromData(data = ROCdat, 
                 MeanSd = T, linewidth = 0.235*1.5) +
                 ggtitle("Repeated test sets")
ggsave_pdfpng(file = paste0("output/images/functional_scores/repeatedtrainTestforPred/ROC_trainTestforPred_", suffix), 
              width = 2.5, height = 2)
```


### Make predictions
```{r}
#Select model
finalModel <- mods_trainTestforPred_trans$logreg

#Save the individual logistic regression models coefficients
finalModel_logregs_coeffs <- map(finalModel, ~.$finalModel$coefficients) 
save(finalModel_logregs_coeffs, file = paste0("output/data/functional_scores/finalModel_logregs_coeffs__", suffix, ".RData"))

#Predict with all subsampled models
allPredsLst <- map2(finalModel, names(finalModel), function(tempMod, fold){ #Loop over train-test splits
        predictions <- predict(tempMod, newdata = datFeats, type = "prob")
        predDf <- data.frame("X1" = predictions[, "X1"], "uniprot_site" = rownames(datFeats))
        return(predDf)
      })
allPredsLstDf <- map_dfr(allPredsLst, function(x){x}, .id = "fold")
save(allPredsLstDf, file = paste0("output/data/functional_scores/allPredsLstDf__", suffix, ".RData"))

#Take median of predictions across sub-sampled models
allPredsDf <- allPredsLstDf %>% group_by(uniprot_site) %>% summarise(X1 = median(X1, na.rm = T))

#Combine with features
ubiFeatPreds <- left_join(ubiFeatProc, dplyr::rename(allPredsDf, "functional_score" = "X1"), by = "uniprot_site")

#Flag training data class
ubiFeatPreds$trainingClass <- NA
ubiFeatPreds$trainingClass[which(ubiFeatPreds$uniprot_site %in% rownames(subset(trainDat_preSplit, response == "X1")))] <- "Positive"
ubiFeatPreds$trainingClass[which(ubiFeatPreds$uniprot_site %in% rownames(subset(trainDat_preSplit, response == "X0")))] <- "Negative"
```



# Export 
```{r}
save(ubiFeatPreds, file = "output/data/functional_scores/ubiFeatPreds__7.RData")
```



## Export for supp table
```{r}
renameCols <- c(
        "uniprot_site"              = "Ubi-site",
        "functional_score"          = "Functional score",
        "trainingClass"             = "Training Class",
        "FinalBoxProcNum"           = "Conservation level",
        "HotspotNew_bool"           = "In hotspot",
        "PRIDE_gold"                = "Gold identification confidence",
        "numReg_notPrtsm"           = "# regulations",
        "numQuant_all"              = "# quantifications",
        "UNIPROT_in_domain"         = "In domain",
        "UNIPROT_in_region"         = "In region",
        "NSP_disorder"              = "Disorder probability",
        "NSP_p_q3_H"                = "Helix probability",
        "NSP_p_q3_C"                = "Coil probability",
        "NSP_rsa"                   = "Relative surface accessibility",
        "interface_label"           = "At interface",
        "PSP_boolOtherLysMod"       = "Site contains other PTMs",
        "PSP_numAdj_5"              = "# Nearby PTMs",
        "mutR_am_pathogenicity_all" = "AlphaMissense pathogenicity"
                
        )
ubiFeatPreds_exp <- ubiFeatPreds[, names(renameCols)] %>% 
  dplyr::rename(!!!setNames(names(renameCols), renameCols)) %>% 
  subset(!is.na(`Functional score`)) %>%
  .[order(.$`Functional score`, decreasing = T), ]
# Export
outDir = "output/data/functional_scores/SuppTable_ScoresFeats.xlsx"
sheet = "scores_and_features"
if (file.exists(outDir)){
  wb <- loadWorkbook(outDir)
} else {
  wb <- createWorkbook(outDir)
}
removeWorksheet(wb, sheet)
addWorksheet(wb, sheet)
writeData(wb, sheet = sheet, x = ubiFeatPreds_exp)
saveWorkbook(wb, outDir, overwrite = TRUE)
```