# Background
In 11_training_models.Rmd, I train the major model used throughout the project.
Here I train additional models. For instance, I do:
- Different negative class subsampling
- Feature ablation

I typically evaluate the models trained here in 12b_iterate_accuracy_and_major_results.Rmd

# Initialise
```{r}
library(tidyverse)
library(purrr)
library(ggplot2)
library(caret)
library(openxlsx)

home_directory <- ""
source(paste0(home_directory, "scripts/essential_visualisations_JvG.R"))
source(paste0(home_directory, "scripts/caret_JvG.R"))
source("scripts/00_objects.R")
source("scripts/01_general.R")
source("scripts/02_model_training.R")
source("scripts/03_visualisations.R")
```

# Load data
```{r}
load("data/functional_scores/datFeats__7.RData")  # Features for all sites, used for final predictions
load("data/functional_scores/trainDat_preSplit__7.RData")  # Features with "response" - the class column
```


# Retraining 1B: Individual feature ablation
Here I remove each feature one-by-one and retrain the model
```{r}
suffix <- "retrain1BFeatAblate"
dir.create(paste0("output/data/functional_scores/", suffix))
resDir <- paste0("output/data/functional_scores/", suffix, "/")
```


## Set up features to use
```{r}
finalFeats <- c(
    "PRIDE_gold", 
    "FinalBoxProcNum",
    "HotspotNew_bool", 
    "mutR_am_pathogenicity_all", 
    "interface_label", 
    "NSP_p_q3_H", 
    "NSP_p_q3_C", 
    "NSP_rsa", 
    "NSP_disorder", 
    "PSP_boolOtherLysMod", 
    "PSP_numAdj_5",
    "numQuant_all", 
    "numReg_notPrtsm", 
    "UNIPROT_in_region", 
    "UNIPROT_in_domain"
  )
finalFeatsLst <- map(finalFeats, ~setdiff(finalFeats, .))
names(finalFeatsLst) <- finalFeats
```


## Make train-test split models
```{r}
#Use protein stratification, with 5-time repeated 5-fold split
numReps <- 5
trainTestforPredSplits <- list()
trainDat_preSplit_sub <- trainDat_preSplit
for (j in 1:numReps){
  splts <- makeFoldsProtstrat(seed = 123 + j, k = 5, data = trainDat_preSplit_sub)
  names(splts) <- paste0("Rep", j, names(splts))
  trainTestforPredSplits <- c(trainTestforPredSplits, splts)
} 
save(trainTestforPredSplits, file = paste0(resDir, "trainTestforPredSplits__", suffix, ".RData"))

# Subset columns for each model
trainTestforPredSplits_lst <- map(finalFeatsLst, function(selFeats){
  lst1 <- map(trainTestforPredSplits, function(lst2){
    map(lst2, function(dat){
      dat <- dat[, c(selFeats, "response")]
      return(dat)
    })
  })
  return(lst1)
})
```


## Train models and make predictions 
```{r}
# Train
resLst_lst <- map2(trainTestforPredSplits_lst, names(trainTestforPredSplits_lst), function(trainTestSplits_temp, modName){
  # Get features
  selFeats <- finalFeatsLst[[modName]]
  # Train model and get results
  resLst <- trainAndMakePreds_logreg(
    trainTestSplits = trainTestSplits_temp,  
    datFeats = datFeats[, selFeats],
    seed = 80
  )
  return(resLst)
})

# Export
map2(resLst_lst, names(resLst_lst), function(resLst, modName){
  coeffs <- map(resLst$mods, ~.$finalModel$coefficients) 
  save(coeffs, file = paste0(resDir, "coeffs_", modName, ".RData"))
  modsPredsTest_df <- resLst$modsPredsTest_df
  save(modsPredsTest_df, file = paste0(resDir, "modsPredsTest_df_", modName, ".RData"))
  allPredsDf <- resLst$allPredsDf
  save(allPredsDf, file = paste0(resDir, "allPredsDf_", modName, ".RData"))
})
```



# Subsampling the negative class at different ratios to positive
```{r}
suffix <- "retrain1Csubsamps"
dir.create(paste0("output/data/functional_scores/", suffix))
resDir <- paste0("output/data/functional_scores/", suffix, "/")
```


## Set up features to use
```{r}
finalFeats <- c(
    "PRIDE_gold", 
    "FinalBoxProcNum",
    "HotspotNew_bool", 
    "mutR_am_pathogenicity_all", 
    "interface_label", 
    "NSP_p_q3_H", 
    "NSP_p_q3_C", 
    "NSP_rsa", 
    "NSP_disorder", 
    "PSP_boolOtherLysMod", 
    "PSP_numAdj_5",
    "numQuant_all", 
    "numReg_notPrtsm", 
    "UNIPROT_in_region", 
    "UNIPROT_in_domain"
  )
```

## Make train-test splits
I  do one five-fold split
```{r}
#Use protein stratification, with 5-time repeated 5-fold split
numReps <- 1
trainTestforPredSplits <- list()
trainDat_preSplit_sub <- trainDat_preSplit[, c(finalFeats, "response")]
for (j in 1:numReps){
  splts <- makeFoldsProtstrat(seed = 123 + j, k = 5, data = trainDat_preSplit_sub)
  names(splts) <- paste0("Rep", j, names(splts))
  trainTestforPredSplits <- c(trainTestforPredSplits, splts)
} 
save(trainTestforPredSplits, file = paste0(resDir, "trainTestforPredSplits__", suffix, ".RData"))
```

## Do subsampling for train-test splits
```{r}
subsamp_args <- list(
  "1x" = list("subsampMult" = 1, "subsampNum" = 10),
  "5x" = list("subsampMult" = 5, "subsampNum" = 5),
  "10x" = list("subsampMult" = 10, "subsampNum" = 5),
  "100x" = list("subsampMult" = 100, "subsampNum" = 5)
)
# Make subsamples
# These are indices that subset trainDat from the corresponding train-test split
subsamp_lst <- map(subsamp_args, function(args){
  # Loop over train-test splits
  subsampsLst <- map(trainTestforPredSplits, function(lst){
    subsamps <- subsample_negatives(resp = lst$trainDat$response,
                        subsampSeed = 70,
                        subsampNum = args$subsampNum,
                        subsampMult = args$subsampMult,
                        reorder = TRUE)
    return(subsamps)
  })
  return(subsampsLst)
})
# Add arguments for the main model (no subsampling)
subsamp_lst$noSubsamp <- map(trainTestforPredSplits, function(lst){
  list("Subsamp1" = 1:nrow(lst$trainDat))
})
save(subsamp_lst, file = paste0(resDir, "subsamp_lst__", suffix, ".RData"))
```


## Train models and make predictions 
Train one subsampling regime at a time to save memory
```{r}
trainSeed <- 80
startTime <- Sys.time()
# Loop over subsampling and +- class weights
for (subsamp_name in names(subsamp_lst)){
  for (useClassWeights in c(T, F)){

    subsampsLst <- subsamp_lst[[subsamp_name]]

    #  subsampsLst <- map(subsampsLst[1], ~.[1])  # Just do one for testing

    # Get train-test splits
    trainTestSplits_subsamp <- map2(subsampsLst, names(subsampsLst), function(subsamps, sampName){
      trainTest_native <- trainTestforPredSplits[[sampName]]
      # Loop over subsamples
      trainTest_subsamp <- map(subsamps, function(inds){
        lst <- trainTest_native
        lst$trainDat <- lst$trainDat[inds, ]
        return(lst)
      })
      names(trainTest_subsamp) <- paste(sampName, names(trainTest_subsamp), sep = "__")
      return(trainTest_subsamp)
    })
    trainTestSplits_subsampUnrav <- purrr::reduce(trainTestSplits_subsamp, function(x, y){c(x, y)})

    # Train with or without class weights
    if (useClassWeights){
      modelName <- paste(subsamp_name, "WWeights")
      mods <- list()
      for (j in 1:length(trainTestSplits_subsampUnrav)){
        splt <- trainTestSplits_subsampUnrav[[j]]
        trainDat <- splt$trainDat
        # Train logistic regression
        mod <- modTrain_LogReg_weighted(feats = trainDat[, -which(colnames(trainDat) == "response")], 
                                                resp = trainDat$response, 
                                                fitControl = fitControl_CV, seed = trainSeed + j,
                                                weightFunc = caseWeights_balanced)
          mods[[j]] <- mod
      }
      names(mods) <- names(trainTestSplits_subsampUnrav)
    } else {
      modelName <- paste(subsamp_name, "NoWeights")
      mods <- list()
      for (j in 1:length(trainTestSplits_subsampUnrav)){
        splt <- trainTestSplits_subsampUnrav[[j]]
        trainDat <- splt$trainDat
        # Train logistic regression
        mod <- modTrain_LogReg_unweighted(feats = trainDat[, -which(colnames(trainDat) == "response")], 
                                          resp = trainDat$response, fitControl = fitControl_CV, seed = trainSeed + j)
        mods[[j]] <- mod
      }
      names(mods) <- names(trainTestSplits_subsampUnrav)
    }

    # Make predictions and get ROC results 
    resLst <- makePreds_fromMods(
      mods = mods,  
      trainTestSplits = trainTestSplits_subsampUnrav,  
      datFeats = datFeats[, finalFeats],
      returnAllPredsLst = T
    )
    # Export
    coeffs <- map(mods, ~.$finalModel$coefficients) 
    save(coeffs, file = paste0(resDir, "coeffs_", modelName, ".RData"))
    modsPredsTest_df <- resLst$modsPredsTest_df
    save(modsPredsTest_df, file = paste0(resDir, "modsPredsTest_df_", modelName, ".RData"))
    allPredsLst <- resLst$allPredsLst
    save(allPredsLst, file = paste0(resDir, "allPredsLst_", modelName, ".RData"))

  }
}
endTime <- Sys.time()
print(endTime - startTime)
```

## Train one model on all data
```{r}
# Train
trainSeed <- 80
modAllDat <- modTrain_LogReg_weighted(feats = trainDat_preSplit_sub[, -which(colnames(trainDat_preSplit_sub) == "response")], 
                                                resp = trainDat_preSplit_sub$response, 
                                                fitControl = fitControl_CV, seed = trainSeed + 1,
                                                weightFunc = caseWeights_balanced)
# Predict on all data
predictions <- predict(modAllDat, newdata = datFeats, type = "prob")
predDf <- data.frame("X1" = predictions[, "X1"], "uniprot_site" = rownames(datFeats))
# Export
modelName <- "allDat"
coeffs <- modAllDat$finalModel$coefficients 
save(coeffs, file = paste0(resDir, "coeffs_", modelName, ".RData"))
allPredsDf <- predDf
save(allPredsDf, file = paste0(resDir, "allPredsDf_", modelName, ".RData"))
```



# Retrain with filtered negative classes
Here I filter the negative class in different ways

## Filter for not regulated in proteomics conditions
### Generate negative classes
```{r}
source("scripts/04b_ubnOmics_filtering.R")
load("output/data/ubnOmics/ubnOmicsComp_hsPRIDE_lst__forCons__8.RData")

# Define number of regulations at cutoffs
fracCutoff <- 0.1
ubnOmics_reg_lst <- classifyRegulatedSites(ubnOmicsComp_hsPRIDE_lst, fracCutoff)
ubnOmics_reg_dat <- ubnOmics_reg_lst$identif
regSites <- subset(ubnOmics_reg_dat, numReg_notPrtsm > 0)$uniprot_pos %>% 
  gsub("_", "_K", .)

# Use this to filter training data
trainDat_preSplit_filt <- trainDat_preSplit
removeBool <- trainDat_preSplit_filt$response == "X0" & rownames(trainDat_preSplit_filt) %in% regSites
trainDat_preSplit_filt <- trainDat_preSplit_filt[!removeBool, ]

# Remove regulations as a column
trainDat_preSplit_filt <- trainDat_preSplit_filt %>% dplyr::select(-c("numReg_notPrtsm"))
table(trainDat_preSplit_filt$response)
```

### Make train-test splits
```{r}
#Use protein stratification, with 1-time repeated 5-fold split
numReps <- 1
trainTestforPredSplits <- list()
for (j in 1:numReps){
  splts <- makeFoldsProtstrat(seed = 123 + j, k = 5, data = trainDat_preSplit_filt)
  names(splts) <- paste0("Rep", j, names(splts))
  trainTestforPredSplits <- c(trainTestforPredSplits, splts)
} 
save(trainTestforPredSplits, file = paste0("output/data/functional_scores/trainTestforPredSplits__", suffix, ".RData"))
# Checkpoint
#load(paste0("output/data/functional_scores/trainTestforPredSplits__", suffix, ".RData"))
```

### Train models 
```{r}
resLst_filt <- trainAndMakePreds_logreg(
  trainTestSplits = trainTestforPredSplits,  
  datFeats = datFeats %>% dplyr::select(-c("numReg_notPrtsm")),
  seed = 80
)
coeffs_filt <- map(resLst_filt$mods, ~.$finalModel$coefficients) 
```

### Inspect results
```{r}
# Inspect AUCs
resLst_filt$AUCdf
mean(resLst_filt$AUCdf$AUC)
sd(resLst_filt$AUCdf$AUC)

# Plot coefficients
bplot_coeffs(coeffs_filt)
ggsave_pdfpng(file = "output/images/functional_scores/negClassFilt/bplot_featImp_negClassFilt_regNum",
                width = 2.75, height = 2.25)
```


## Use ClinVar benign as the negative class
### Set up negative class
```{r}
# Load clinvar
load("output/data/genetics/clinvarDatSub__2.RData")
clinvarDatSub <- subset(clinvarDatSub, grepl("missense_variant", MC))
clinvarDatSub <- clinvarDatSub %>% mutate(clinvar_label = factor(clinvar_label, levels = c("Benign", "VUS", "Pathogenic")))
levels(clinvarDatSub$clinvar_label)[2] <- "Uncertain" 
# Filter for benign sites
benignSites <- subset(clinvarDatSub, clinvar_label == "Benign")$uniprot_site

# Use this to filter training data
trainDat_preSplit_filt <- trainDat_preSplit
keepBool <- trainDat_preSplit_filt$response == "X1" | rownames(trainDat_preSplit_filt) %in% benignSites
trainDat_preSplit_filt <- trainDat_preSplit_filt[keepBool, ]
table(trainDat_preSplit_filt$response)
```


### Make train-test splits
```{r}
#Use protein stratification, with 1-time repeated 5-fold split
numReps <- 1
trainTestforPredSplits <- list()
for (j in 1:numReps){
  splts <- makeFoldsProtstrat(seed = 123 + j, k = 5, data = trainDat_preSplit_filt)
  names(splts) <- paste0("Rep", j, names(splts))
  trainTestforPredSplits <- c(trainTestforPredSplits, splts)
} 
save(trainTestforPredSplits, file = paste0("output/data/functional_scores/trainTestforPredSplits__", suffix, ".RData"))
#load(paste0("output/data/functional_scores/trainTestforPredSplits__", suffix, ".RData"))
```

### Train models 
```{r}
resLst_filt <- trainAndMakePreds_logreg(
  trainTestSplits = trainTestforPredSplits,  
  datFeats = datFeats,
  seed = 80
)
coeffs_filt <- map(resLst_filt$mods, ~.$finalModel$coefficients) 
```

### Inspect results
```{r}
# Inspect AUCs
resLst_filt$AUCdf
mean(resLst_filt$AUCdf$AUC)
sd(resLst_filt$AUCdf$AUC)

# Plot coefficients
bplot_coeffs(coeffs_filt)
ggsave_pdfpng(file = "output/images/functional_scores/negClassFilt/bplot_featImp_negClassFilt_CVbenign",
                width = 2.75, height = 2.25)
```