# Background
Here I perform analysis comparing models trained under different class balance regimes, 
to address the class imbalance in our data


# Initialise
```{r}
library(tidyverse)
library(purrr)
library(ggplot2)
library(ggridges)
library(readxl)
library(openxlsx)
library(data.table)
library(caret)

home_directory <- ""
source(paste0(home_directory, "scripts/essential_visualisations_JvG.R"))
source(paste0(home_directory, "scripts/enrichment_analysis_JvG.R"))
source(paste0(home_directory, "scripts/caret_JvG.R"))
source("scripts/01_general.R")
source("scripts/00_objects.R")
source("scripts/03_visualisations.R")
source("scripts/04_ubnOmics.R")
source("scripts/02_model_training.R")
source("scripts/16_model_accuracy.R") # Auxiliary functions for assessing accuracy
```

## Functions
```{r}
# Function: Plotting correlation matrix as heatmap
hmap_corm <- function(corM, fillLims = c(0.8, 1), fontSize = 4){
  # Convert correlation matrix to data frame and pivot longer
  corM_long <- as.data.frame(corM) %>%
    rownames_to_column(var = "row") %>%
    pivot_longer(
      cols = -row,
      names_to = "column",
      values_to = "correlation"
    ) %>%
    mutate(
      row = factor(row, levels = rownames(corM)),
      column = factor(column, levels = colnames(corM))
    )

  # Plot with ggplot
  plt <- ggplot(corM_long, aes(x = column, y = row, fill = correlation)) +
    geom_tile() +
    scale_fill_gradientn(colours = rev(brewer.pal(9, "RdBu")), limits = fillLims) +
    comfy_theme(rotate_x_text = T) + coord_fixed()  + 
    theme(
      axis.text.y = element_text(colour = "black", hjust = 1, size = fontSize),
      axis.text.x = element_text(colour = "black", hjust = 1, size = fontSize, angle = 90))
  return(plt)
}

```

# Compare different subsampling procedures (retrain1Csubsamps)
These models are trainined in 11c_extra_model_training.Rmd
## Load results
```{r}
# Get model names in the right order
resDir <- "output/data/functional_scores/retrain1Csubsamps/"
mods <- dir(resDir) %>% .[grep("modsPredsTest_df_", .)] %>% gsub("modsPredsTest_df_", "", .) %>% gsub(".RData", "", ., fixed = T)
modsDf <- data.frame("mods" = mods) %>% mutate(x = as.numeric(gsub("x.+", "", mods)))
modsDf$x[which(is.na(modsDf$x))] <- 100000
modsDf <- modsDf[order(modsDf$x, decreasing = F), ]
mods <- modsDf$mods
# Load data
testSetPreds_lst <- list()
allPreds_lst <- list()
for (modelName in mods){
  load(paste0(resDir, "modsPredsTest_df_", modelName, ".RData"))
  testSetPreds_lst[[modelName]] <- modsPredsTest_df
  load(file = paste0(resDir, "allPredsLst_", modelName, ".RData"))
  allPredsDf <- map_dfr(allPredsLst, function(x){x}, .id = "Resample")
  allPreds_lst[[modelName]] <- allPredsDf
}

# Load model trained on all data
load(paste0(resDir, "allPredsDf_allDat.RData"))
allPredsDf_allDat <- allPredsDf
```


## Compare AUC
```{r}
AUCdf <- map_dfr(testSetPreds_lst, function(predsDf){
    AUCs <- predsDf %>% split(.$Resample) %>% 
        map(~ROCAUC_fromPredictor(data = ., obsCol = "obs", predCol = "X1")) %>% unlist
    data.frame("Resample" = names(AUCs), "AUC" = AUCs)
}, .id = "Regime")  %>% mutate(Regime = factor(Regime, levels = mods)) %>% .[order(.$Regime), ]
AUCsumm <- AUCdf %>% group_by(Regime) %>% summarise(AUC_mean = mean(AUC), AUC_sd = sd(AUC)) %>% ungroup 

# Plot
ggplot(AUCsumm, aes(x = Regime, y = AUC_mean)) +
            geom_bar(stat = "identity", width = 0.8, linewidth = 0.235, fill = "grey") +
            geom_errorbar(aes(ymin = AUC_mean - AUC_sd, ymax = AUC_mean + AUC_sd), 
                                        width = 0.2, color = "black", size = 0.235) +
            comfy_theme(include_legend = F, rotate_x_text = T) + 
            theme(axis.title.y = element_blank(), axis.text.y = element_text(colour = "black", size = 6, hjust = 1)) + 
            ggtitle("Test-set AUC")
ggsave_pdfpng(file = "output/images/functional_scores/classImbalance/bar_retrain1Csubsamps_AUC", width = 2, height = 2)
```


## Compare predictions
```{r}
# Combine all predictions
allPreds_comb <- map_dfr(allPreds_lst, function(x){
    if ("Resample" %in% colnames(x) == F){colnames(x)[which(colnames(x) == "fold")] <- "Resample"}
    return(x)
}, .id = "Regime") %>% 
  mutate(Regime = factor(Regime, levels = mods)) %>% .[order(.$Regime), ]

## Pivot to wider
allPreds_combWide <- allPreds_comb %>% mutate(id = paste(Regime, Resample, sep = "__")) %>% 
    dplyr::select(c("uniprot_site", "X1", "id")) %>% 
    pivot_wider(names_from = "id", values_from = "X1", id_cols = "uniprot_site") %>% 
    as.data.frame %>% column_to_rownames("uniprot_site") 

# Spearman's correlation
corM <- cor(allPreds_combWide, method = "spearman")
# Statistics
mean(corM)
min(corM)
max(corM)

# Plot
hmap_corm(corM, fillLims = c(0.65, 1), fontSize = 2) +
  ggtitle("Correlating predictions between models")
ggsave_pdfpng(file = "output/images/functional_scores/classImbalance/hmap_retrain1Csubsamps_scoreCorrs", width = 10, height = 10)
```

Take median of predictions across resamples and compare these
```{r}
# Take median of predictions
allPreds_combMedians <- allPreds_comb %>% group_by(Regime, uniprot_site) %>% 
    summarise(X1 = median(X1)) %>% ungroup
allPreds_combMediansWide <- allPreds_combMedians %>% pivot_wider(names_from = "Regime", values_from = "X1") %>% 
    as.data.frame %>% column_to_rownames("uniprot_site")
corM_medians <- cor(allPreds_combMediansWide, method = "spearman")

# Statistics
mean(corM_medians)
min(corM_medians)
max(corM_medians)

#Plot
hmap_corm(corM_medians, fillLims = c(0.65, 1), fontSize = 4) +
  ggtitle("Median prediction overall")
ggsave_pdfpng(file = "output/images/functional_scores/classImbalance/hmap_retrain1Csubsamps_TotalMedianCorrs", width = 3, height = 3)

# Correlate just two conditions
ggplot(allPreds_combMediansWide, aes(x = `10x NoWeights`, y = `10x WWeights`)) + 
  geom_point(shape = 16, size = 0.5, alpha = 0.3) + 
  comfy_theme() + 
  ggtitle("Median prediction across folds")
ggsave_pdfpng(file = "output/images/functional_scores/classImbalance/pplot_retrain1Csubsamps_TotalMedianCorrs_10x", width = 2.5, height = 2.5)
```

Take median of predictions across resamples within each fold and compare these
```{r}
# Take median of predictions
allPreds_comb <- allPreds_comb %>% mutate(Resample = factor(Resample)) %>% mutate(Fold = Resample)
levels(allPreds_comb$Fold) <- gsub("__.+$", "", levels(allPreds_comb$Fold))
allPreds_combMedians_Fold <- allPreds_comb %>% 
group_by(Regime, uniprot_site, Fold) %>% 
    summarise(X1 = median(X1)) %>% ungroup
allPreds_combMedians_FoldWide <- allPreds_combMedians_Fold %>% 
  .[order(.$Fold, .$Regime), ] %>% 
  mutate(RegimeFold = paste(Regime, Fold, sep = "__")) %>% 
  dplyr::select(c("RegimeFold", "X1", "uniprot_site")) %>%
  pivot_wider(names_from = "RegimeFold", values_from = "X1") %>% 
    as.data.frame %>% column_to_rownames("uniprot_site")
corM_FoldMedians <- cor(allPreds_combMedians_FoldWide, method = "spearman")

# Plot
hmap_corm(corM_FoldMedians, fillLims = c(0.65, 1), fontSize = 4) +
  ggtitle("Median of predictions within folds")
ggsave_pdfpng(file = "output/images/functional_scores/classImbalance/hmap_retrain1Csubsamps_FoldMedianCorrs", width = 5, height = 5)
```



# Analysis of similarity of final models
Here I look at the final models trainined in 11_training_models.Rmd

## Load and prepare final predictions
```{r}
# Load
suffix <- "attempt11_revisions"
load(paste0("output/data/functional_scores/allPredsLstDf__", suffix, ".RData")) # Load predictions

# Pivot to wider
allPredsLstDf_Wide <- allPredsLstDf %>% mutate(id = fold) %>% 
    dplyr::select(c("uniprot_site", "X1", "id")) %>% 
    pivot_wider(names_from = "id", values_from = "X1", id_cols = "uniprot_site") %>% 
    as.data.frame %>% column_to_rownames("uniprot_site") 
# Get median across models
allPredsLstDf_Wide$final_score <- apply(allPredsLstDf_Wide[, which(colnames(allPredsLstDf_Wide) %in% allPredsLstDf$fold)], 1, "median", na.rm = T)
# Map in scores from a model trained on all data
load(paste0(resDir, "allPredsDf_allDat.RData"))
allPredsDf_allDat <- allPredsDf
allPredsLstDf_Wide <- left_join(
  rownames_to_column(allPredsLstDf_Wide, "uniprot_site"),
  dplyr::rename(allPredsDf_allDat, "scoreAllData" = "X1"), 
  by = "uniprot_site") %>%  
  column_to_rownames("uniprot_site")
```

## Plot different comparisons
```{r}
##### Everything #####
corM <- cor(allPredsLstDf_Wide, method = "spearman")
# Statistics
mean(corM)
min(corM)
max(corM)
# Correlation between final predictions and each model
mean(corM["final_score", grep("^Rep\\d", colnames(corM))])
colnames(corM)
# Plot
hmap_corm(corM, fillLims = c(0.65, 1), fontSize = 4) +
  ggtitle("Final predictions")
ggsave_pdfpng(file = "output/images/functional_scores/classImbalance/hmap_finalScoresAll_scoreCorrs", width = 5, height = 5)

##### Just 25 models #####
corM <- cor(allPredsLstDf_Wide[, grep("^Rep\\d", colnames(allPredsLstDf_Wide))], method = "spearman")
# Statistics
mean(corM)
min(corM)
max(corM)
# Plot
hmap_corm(corM, fillLims = c(0.65, 1), fontSize = 4) +
  ggtitle("Final predictions")
ggsave_pdfpng(file = "output/images/functional_scores/classImbalance/hmap_finalScores25mods_scoreCorrs", width = 5, height = 5)
```