# Background
Here I analyse the positional importance score

# Initialise
```{r}
library(tidyverse)
library(purrr)
library(ggplot2)
library(ggridges)
library(readxl)
library(openxlsx)
library(data.table)

home_directory <- ""
source(paste0(home_directory, "scripts/essential_visualisations_JvG.R"))
source(paste0(home_directory, "scripts/enrichment_analysis_JvG.R"))
source("scripts/01_general.R")
source("scripts/00_objects.R")
source("scripts/03_visualisations.R")
source("scripts/04_ubnOmics.R")
source("scripts/17_linear_models.R")

#Load data
load("data/functional_scores/ubiFeatPreds__7.RData")
load("data/features/ubnOmics/ubnOmicsFeauresAll__4.RData")
load("data/features/ubnOmics/ubnOmicsComp_hsPRIDE_lst__8.RData")
load("data/hotspots/hotspotSites__3.RData")

#Set up bin breaks for the functional score, and other aesthetic paramters
funcScoreBreaks <- c(0, 0.2, 0.4, 0.6, 0.8, 1)
funcScoreBinCols <- brewer.pal(9, "Blues")[c(1, 3, 5, 7, 9)]
funcScoreFillAlpha <- 0.3

#Get gene identifiers for our ubi sites
library(AnnotationDbi)
library(org.Hs.eg.db)
collapseFunc <- function(x){
    x <- unique(x)
    if (is.na(x[1])){
        return(NA)
    } else {
        return(paste(x, collapse = ";"))
    }
}
geneMap <- AnnotationDbi::select(org.Hs.eg.db, keys = unique(ubiFeatPreds$uniprot), keytype = "UNIPROT",
                                    columns = c("SYMBOL")) %>% 
                                group_by(UNIPROT) %>% 
                                summarise(gene = collapseFunc(SYMBOL))
ubiFeatPreds <- left_join(ubiFeatPreds, dplyr::rename(geneMap, "uniprot" = "UNIPROT"), by = "uniprot")
##Expand out by gene when they are not unique
ubiFeatPreds_byGene <- separate_rows(ubiFeatPreds, "gene", sep = ";") %>% 
                        mutate(gene_site = paste(gene, site, sep = "_")) 
```

```{r}
## Add quantile-normalised score and rank within protein
ubiFeatPreds <- ubiFeatPreds %>%
    group_by(uniprot) %>%
    mutate(
        numUbiSites = sum(!is.na(functional_score)),
        funcScoreRank = if_else(numUbiSites > 0, rank(-functional_score, ties.method = "first"), NA_integer_),
        funcScoreQuant = to_quantiles(functional_score),
        numUbiSites = if_else(numUbiSites > 0, numUbiSites, NA_integer_)
    ) %>%
    ungroup() %>%
    as.data.frame()
```

# Overview
## Score distributions and number of sites
```{r}
#Plot non-protdeg vs unannotated sites
plotDat <- ubiFeatPreds %>%
    subset(!is.na(trainingClass)) %>% 
    subset(!is.na(functional_score))
median_lines <- plotDat %>%
  group_by(trainingClass) %>%
  summarize(
    med = median(functional_score),
    density_at_median = {
      d <- density(functional_score)
      approx(d$x, d$y, xout = median(functional_score))$y
    }
  )
ggplot(plotDat, aes(x = functional_score, y = trainingClass, fill = trainingClass)) +
    geom_density_ridges(linewidth = 0.235) +
    geom_vline(xintercept = median_lines$med, color = "black", size = 0.235) +
    comfy_theme() +
    scale_fill_manual(values = setNames(c(PSPcolours["reg"], "#e5e5e5"), c("Positive", "Negative")),
                      labels = c("Regulatory", "Unannotated")) +
    lims(x = c(0, 1)) +
    labs(x = "Functional score")
ggsave_pdfpng(file = "output/images/functional_scores/overview/ridge_funcScoreDist",
                width = 3.25, height = 1.5)

#How many sites at different cutoffs?
##Score cutoff that recovers 50% of nondeg sites
scoreCutoff <- subset(ubiFeatPreds, trainingClass == "Positive" & !is.na(functional_score))$functional_score %>% median
print(scoreCutoff)
scoreCutoff <- 0.65
table(ubiFeatPreds$trainingClass == "Positive", ubiFeatPreds$functional_score > scoreCutoff)

fishers_DE_pathway(background_genes = ubiFeatPreds$uniprot_site,
                   pathway_genes = subset(ubiFeatPreds, trainingClass == "Positive")$uniprot_site, 
                   DE_genes = subset(ubiFeatPreds, functional_score > scoreCutoff)$uniprot_site)

# How many sites in each class?
table(ubiFeatPreds$trainingClass)
```


# Feature importance 
```{r}
# Load model
load("data/functional_scores/finalModel_logregs_coeffs__attempt11_revisions.RData")
bplot_coeffs(finalModel_logregs_coeffs)
ggsave_pdfpng(file = "output/images/functional_scores/featureImportance/bplot_featImp_featGroups",
                width = 2.75, height = 2.25)
```


# Regulation in omics data
## Proteasome inhibition
```{r}
# Add prtsm inhib data
prtsmInhibCols <- c("Bortezomib_meanlog2FC", "MG-132_meanlog2FC", "Epoxomicin_meanlog2FC")
if (length(intersect(prtsmInhibCols, colnames(ubiFeatPreds))) == 0){
    ubiFeatPreds <- left_join(ubiFeatPreds, ubnOmicsFeaturesAll[, c("uniprot_pos", prtsmInhibCols)],
                                by = "uniprot_pos")
}

# Bin score
binBreaks <- funcScoreBreaks
binN <- length(binBreaks)
binDat <- binner(data = subset(ubiFeatPreds, !is.na(functional_score)), 
                valCol = "functional_score", breakType = "float",
                bin_breaks = binBreaks)

# Pivot by inhibitor type
binDat <- pivot_longer(binDat, cols = prtsmInhibCols, names_to = "treatType", values_to = "logFC") %>% 
            mutate(treatType = factor(treatType, levels = prtsmInhibCols))
levels(binDat$treatType) <- gsub("_meanlog2FC$", "", levels(binDat$treatType))


# Plot
## Violin
pltLst <- map(levels(binDat$treatType), function(lev){
    plt <- binDat %>% subset(treatType == lev) %>%
        subset(abs(logFC) < 5) %>%
        ViolBoxplot(., vars = list(x = "Bin", y = "logFC", colour = "Bin", fill = "Bin"), includeLegend = F) +
        geom_hline(yintercept = 0, linewidth = 0.235, linetype = "dashed")  +
        scale_colour_manual(values = rep(brewer.pal(9, "Blues")[9], binN)) +
        scale_fill_manual(values = alpha(funcScoreBinCols, funcScoreFillAlpha)) +
        labs(x = "Functional score", y = "Avg. log2FC (vs control)") +
        ggtitle(lev)
    return(plt)
})
ggarrange(plotlist = pltLst, nrow = 1)
ggsave_pdfpng(file = "output/images/functional_scores/regulation/viol_funcScore_prtsmInhib",
                width = 6, height = 2)

## Dot with IQR
pltLst <- map(levels(binDat$treatType), function(lev){
    plt <- binDat %>% subset(treatType == lev) %>%
        MeanIQRPlot(., vars = list(x = "Bin", y = "logFC", colour = "Bin"), includeLegend = F, pointSize = 3) +
        geom_hline(yintercept = 0, linewidth = 0.235, linetype = "dashed")  +
        scale_colour_manual(values = rep(ubiSiteColour, binN)) +
        labs(x = "Functional score", y = "Avg. log2FC (vs control)") +
        ggtitle(lev)
 plt
})
ggarrange(plotlist = pltLst, nrow = 1)
ggsave_pdfpng(file = "output/images/functional_scores/regulation/meanIQR_funcScore_prtsmInhib",
                width = 5.25, height = 2)
```

## Linear models to test significance
```{r}
prepModelResults_prtmsi <- function(x){
  x %>% subset(termType == "main effect") %>%
    dplyr::relocate("model") %>%
    dplyr::select(-"termType") %>%
    mutate(model = factor(model, levels = c("MG-132", "Bortezomib", "Epoxomicin"))) %>%
    .[order(.$model), ] %>%
    mutate(term = gsub("Bin", "", term)) %>%
    distinct %>%
    dplyr::rename(
      "Treatment" = "model",
      "Main effect term" = "term",
      "Term estimate" = "Estimate"
    )
}

lmLst <- map(levels(binDat$treatType), function(lev){
  subDat <- subset(binDat, treatType == lev)
  lm(logFC ~ Bin, data = subDat)
})
names(lmLst) <- levels(binDat$treatType)
coefDat <- format_model_coefs(lmLst)

coefDat_prtmsi <- prepModelResults_prtmsi(coefDat)
write.csv(coefDat_prtmsi, file = "output/data/functional_scores/coefDat_prtmsi_linearModels.csv")
```


# Nuclear localisation signals
## Extract and process from uniprot API search
```{r}
load("output/data/structure/domainDat_cache.RData")  # This loads domainDatRaw

#Process domain annotations + similar
domainProcessArgs <-  c("Motif", "MOTIF", "motif")
domainDat <- filter(domainDatRaw, !is.na(!!sym(domainProcessArgs[1]))) %>% as.data.frame
#Process annotations
annot <- domainDat[, domainProcessArgs[1]]
annot <- gsub(paste0("^", domainProcessArgs[2], " "), "", annot)
##Split multiple annotations per uniprot
annotSplt <- strsplit(annot, paste0("; ", domainProcessArgs[2], " "))
##Add in uniprot
annotSplt <- map2(annotSplt, rownames(domainDat), function(x, y){paste(y, x, sep = "; ")})
##Unlist and process individually
annotSpltUnlst <- unlist(annotSplt)
##Extract info
annotProcLst <- list()
for (i in 1:length(annotSpltUnlst)){
    annotChar <- annotSpltUnlst[i]
    annotCharSplt <- strsplit(annotChar, "; ")[[1]]
    #Remove matches that have ? or are isoforms
    if (grepl("?", annotCharSplt[2], fixed = T) | 
        grepl("-", annotCharSplt[2], fixed = T)){
        annotProcLst[[i]] <- NULL
    } else {
        resLst <- list()
        ##Get sequence positions
        vecSplt <- strsplit(annotCharSplt[2], "..", fixed = T)
        sequence <- map(vecSplt, function(y){
            if (length(y) == 1){
            return(as.numeric(y))
            } else {
            return(y[1]:y[2])
            }
        }) %>% unlist
        resLst$sequence <- sequence
        ##Get other info
        resLst$note <- gsub("/note=", "", annotCharSplt[3], fixed = T)
        resLst$evidence <- gsub("/evidence=", "", annotCharSplt[4], fixed = T)
        resLst$uniprot <- annotCharSplt[1]
        annotProcLst[[i]] <- resLst
    }
}
nullBool <- map(annotProcLst, ~is.null(.)) %>% unlist
annotProcLst <- annotProcLst[!nullBool]
#Make motif UID
for (i in 1:length(annotProcLst)){
    annotProcLst[[i]]$UID <- paste0("motif", i)
}

#Map ubi-sites into motifs
relevantUnips <- map(annotProcLst, ~.$uniprot) %>% unlist %>% unique
ubnSitesLst <- ubiFeatPreds %>% subset(uniprot %in% relevantUnips) %>% split(.$uniprot) %>% map(~.$pos)
for (i in 1:length(annotProcLst)){
    annotProcLst[[i]]$ubiSites <- NA
    x <- annotProcLst[[i]]
    unip <- x$uniprot 
    motifPosns <- x$sequence
    sites <- ubnSitesLst[[unip]]
    int <- intersect(sites, motifPosns)
    if (length(int) != 0){
      sites <- paste(unip, int, sep = "_")
      annotProcLst[[i]]$ubiSites <- sites
    } 
  }
annotProcLst_wUbi <- annotProcLst[map(annotProcLst, ~!is.na(.$ubiSites[1])) %>% unlist]

#Extract info for ubi-sites in motifs
ubisitesMotifsDf <- map(annotProcLst_wUbi, function(x){
    data.frame("motifID" = x$UID, "motifNote" = x$note, "motifEvidence" = x$evidence, 
               "uniprot_pos" = paste(x$ubiSites, sep = ";"))
}) %>% 
purrr::reduce(rbind) %>% 
separate_rows("uniprot_pos", sep = ";") %>% 
group_by(uniprot_pos) %>% 
summarise(
    motifID = paste(motifID, collapse = ";"),
    motifNote = paste(motifNote, collapse = ";"),
    motifEvidence = paste(motifEvidence, collapse = ";")
)

#Add to feats df
ubiFeatPreds <- left_join(ubiFeatPreds, ubisitesMotifsDf, by = "uniprot_pos")
ubiFeatPreds$inNLS <- grepl("nuclear localization signal", tolower(ubiFeatPreds$motifNote))

table(ubiFeatPreds$inNLS)

#Export
ubiFeatPreds_NLS <- ubiFeatPreds[, c("uniprot_site", "inNLS", colnames(ubiFeatPreds)[grep("^motif", colnames(ubiFeatPreds))])]
save(ubiFeatPreds_NLS, file = "output/data/functional_scores/ubiFeatPreds_NLS__7.RData")
```


## Inspect NLSs
```{r}
# Functional scores
## Visualise
NLSCols <- setNames(c("#6356a5", PSPcoloursLine["notReg"]), c("Yes", "No"))
ubiFeatPreds <- mutate(ubiFeatPreds, inNLSfact = factor(make.names(1*inNLS), levels = c("X0", "X1")))
levels(ubiFeatPreds$inNLSfact) <- c("No", "Yes")
### Label sites
sitesToLabel <- c("P38936_K154", "Q01469_K24", "Q01469_K34", "Q9UBU8_K143")
labelDat <- subset(ubiFeatPreds, uniprot_site %in% sitesToLabel) %>% 
            mutate(gene_site_nice = paste(gene, site, sep = " "))

### Plot
ubiFeatPreds %>%
ViolBoxplot(., vars = list(x = "inNLSfact", y = "functional_score", colour = "inNLSfact", fill = "inNLSfact"),
                includeLegend = F) + 
    geom_point(data = labelDat, aes(x = inNLSfact, y = functional_score),
                shape = 21, size = 1, stroke = 0.71*0.25, colour = "black", fill = alpha(NLSCols[1], 0.8), 
                inherit.aes = F) + 
    geom_text(data = labelDat, aes(x = inNLSfact, y = functional_score, label = gene_site_nice), 
              colour = "#3f3f3e", size = 1.875*5/5) +
    scale_colour_manual(values = NLSCols) +
    scale_fill_manual(values = alpha(NLSCols, 0.2)) + 
    labs(x = "Ubi-site in NLS", y = "Positional importance score") +
    ggtitle("Nuclear localisation sequences")
ggsave_pdfpng(file = "output/images/functional_scores/NLS/viol_inNLS_funcScore",
                width = 1.6, height = 1.8)

### Test
wilcox.test(functional_score ~ inNLSfact, data = ubiFeatPreds)

## Filter
cutoff <- 0.65
print(paste0("NLS ubi-sites with score > ", cutoff))
print(length(which(ubiFeatPreds$inNLS == T & ubiFeatPreds$functional_score > cutoff)))
NLSfuncSites <- ubiFeatPreds[which(ubiFeatPreds$inNLS == T & ubiFeatPreds$functional_score > cutoff), ]
table("score" = ubiFeatPreds$functional_score > cutoff, "NLS" = ubiFeatPreds$inNLS == T)

# Inspect
NSLSites_summ <- ubiFeatPreds[which(ubiFeatPreds$inNLS == T & !is.na(ubiFeatPreds$functional_score)), ] %>% 
    group_by(uniprot) %>% 
    mutate(protMean = mean(functional_score, na.rm = TRUE)) %>% 
    ungroup() %>% 
    .[order(.$protMean, decreasing = T), ] %>% 
    dplyr::relocate(c("functional_score", "BortMG_meanlog2FC", "gene")) 
NSLfuncSites_summ <- NLSfuncSites %>% 
    group_by(uniprot) %>% 
    mutate(protMean = mean(functional_score, na.rm = TRUE)) %>% 
    ungroup() %>% 
    .[order(.$protMean, decreasing = T), ] %>% 
    dplyr::relocate(c("functional_score", "BortMG_meanlog2FC", "gene")) 
write.csv(NSLfuncSites_summ, file = "output/data/functional_scores/NLSfuncSites.csv", na = "")

# Export for supplementary table
renameCols <- c(
    "uniprot_site" = "ubi-site",
    "functional_score" = "Functional score",
    "gene" = "Gene",
    "motifNote" = "UNIPROT motifs"
)
NSLSites_summExp <- NSLSites_summ[, names(renameCols)] %>% 
    dplyr::rename(!!!setNames(names(renameCols), renameCols))
# Export
outDir = "output/data/functional_scores/SuppTable_predFuncSites.xlsx"
sheet = "NLS_sites"
if (file.exists(outDir)){
  wb <- loadWorkbook(outDir)
} else {
  wb <- createWorkbook(outDir)
}
removeWorksheet(wb, sheet)
addWorksheet(wb, sheet)
writeData(wb, sheet = sheet, x = NSLSites_summExp)
saveWorkbook(wb, outDir, overwrite = TRUE)

# Proteasome inhibition
ViolBoxplot(ubiFeatPreds, vars = list(x = "inNLS", y = "BortMG_meanlog2FC", colour = "inNLS"))


# Regulation
sites <- c("P38936_K154", "Q9UBU8_K143")
omicsCondsLst <- extractOmicsConds(sites, ubnOmicsComp_hsPRIDE_lst)
omicsCondsDat <- map_dfr(omicsCondsLst, function(x){data.frame("combinedDescUID" = names(x), logFC = x)}, .id = "site")
omicsCondsDat <- subset(omicsCondsDat, grepl("UV", combinedDescUID))
## Plot
ggplot(omicsCondsDat, aes(x = combinedDescUID, y = logFC)) +   
    geom_hline(yintercept = 0, linewidth = 0.235, colour = "black") +
    geom_col() + 
    comfy_theme(rotate_x_text = T, x_text_angle = 90, include_xaxis = F) + 
    facet_wrap(~site, scale = "free")
ggsave_pdfpng(file = "output/images/functional_scores/NLS/bar_UV_CDKN1A_MORF4L1", width = 3, height = 5)
```


## Check presence in subcellular proteomics 
### Elia et al
```{r}
# Load data
eliaPrtms <- list()
eliaPrtms$nucl <- read_xls(path = paste0(home_directory, "data/supplementary_tables/Elia_MolCell_2015/table_S4.xls"), 
                            sheet = "DiGly sites in SCX Nuc screens")
eliaPrtms$cell_UV <- read_xls(path = paste0(home_directory, "data/supplementary_tables/Elia_MolCell_2015/table_S2.xls"), 
                            sheet = "DiGly sites in UV screen")
eliaPrtms$cell_IR <- read_xls(path = paste0(home_directory, "data/supplementary_tables/Elia_MolCell_2015/table_S2.xls"), 
                            sheet = "DiGly sites in IR screen")
eliaPrtms <- map(eliaPrtms, function(dat){
    dat$gene_pos <- paste(dat$gene_symbol, dat$site_position, sep = "_")
    return(dat)
})
# Get site lists 
eliaSites <- list( 
    nucl = unique(eliaPrtms$nucl$gene_pos),
    cell = unique(c(eliaPrtms$cell_UV$gene_pos, eliaPrtms$cell_IR$gene_pos))
)
ggVennDiagram_nice(eliaSites)
# Map into feats
ubiFeatPreds$gene_pos <- paste(ubiFeatPreds$gene, ubiFeatPreds$pos, sep = "_")
for (name in names(eliaSites)){
    ubiFeatPreds[, paste0("eliaPresent_", name)] <- ubiFeatPreds$gene_pos %in% eliaSites[[name]]
}
table(ubiFeatPreds$eliaPresent_nucl, ubiFeatPreds$eliaPresent_cell)

# Inspect quantification of NLS sites
ubiFeatPreds$inNLS_and_func <- ubiFeatPreds$inNLS & ubiFeatPreds$functional_score > 0.65
ubiFeatPreds$inNLS_and_func[which(is.na(ubiFeatPreds$inNLS_and_func))] <- FALSE
NLSeliaSumm <- ubiFeatPreds %>% 
# subset(functional_score > 0.65) %>%
catBoolSummariser(data = ., catCol = "inNLS_and_func",
                  boolCols = c("eliaPresent_nucl", "eliaPresent_cell")) %>% as.data.frame
View(NLSeliaSumm)
subset(ubiFeatPreds, eliaPresent_nucl == T & inNLS == T)$functional_score %>% sort

# Plot enrichment of NLS sites
NLScolsBar <- setNames(rep(NLSCols[1], 2), c("Nucleus", "Whole-cell"))
NLSeliaSumm_fPlot <- data.frame("col" = colnames(NLSeliaSumm), "val" = as.numeric(NLSeliaSumm[which(NLSeliaSumm$inNLS_and_func == T), ])) %>% 
                    subset(col %in% c("inNLS_and_func", "num_rows") == F) %>% 
                    mutate(colType = strsplit(col, "_") %>% map(~rev(.)[1]) %>% unlist,
                           compartment = ifelse(grepl("nucl", col), "Nucleus", "Whole-cell")) %>% 
                    mutate(compartment = factor(compartment, levels = c("Whole-cell", "Nucleus"))) %>%
                    subset(colType %in% c("num", "enr")) %>% 
                    dplyr::select(-c("col")) %>%
                    pivot_wider(names_from = "colType", values_from = "val") %>% 
                    mutate(log2enr = log2(enr))
ggplot(NLSeliaSumm_fPlot, aes(x = compartment, y = log2enr, fill = compartment, colour = compartment)) +
    geom_bar(stat = "identity", width = 0.8, size = 0.235) +
    geom_text(aes(label = num), vjust = -0.5, color = "black", size = 1.875*6/5) +
    comfy_theme(include_xaxis = F, rotate_x_text = T, include_legend = F) +
    theme(axis.title.x = element_blank()) +
    scale_fill_manual(values = alpha(NLScolsBar, 0.4)) +
    scale_colour_manual(values = NLScolsBar) +
    labs(y = "Log2 enrichment") +
    ggtitle("Detection of functional NLS ubi-sites in subcellular ubi-omics (Elia et al.)")
ggsave_pdfpng(file = "output/images/functional_scores/NLS/bplot_NLS_elia_enr", 
                width = 1, height = 1.775)
```


# CPTAC
## Prep CPTAC ubn-ome data 
```{r}
CPTAC_ubi_list <- list()
```

### LSCC 2021
```{r}
#Read in
CPTAC_ubi_raw <- read_xlsx(paste0(home_directory, "data/supplementary_tables/Satpathy_Cell_2021/Table_S3.xlsx"), sheet = "Table S3L",
                            skip = 2)
##Correct identifiers
colnames(CPTAC_ubi_raw) <- gsub("C3N.", "C3N-", colnames(CPTAC_ubi_raw), fixed = T)   
colnames(CPTAC_ubi_raw) <- gsub("C3L.", "C3L-", colnames(CPTAC_ubi_raw), fixed = T)  
#Prep 
CPTAC_ubi_raw <- CPTAC_ubi_raw %>% 
            mutate(site = strsplit(id, "_") %>% map(~gsub("[a-z]", "", .[3])) %>% unlist) %>% 
            mutate(site = gsub(" ", "", site)) %>% 
            mutate(gene_site = paste(geneSymbol, site, sep = "_")) 
#Get numerical columns
CPTAC_ubi_numCols <- colnames(CPTAC_ubi_raw) %>% .[grep("^C3", .)]
CPTAC_ubi_numColsTum <- CPTAC_ubi_numCols[-grep("N$", CPTAC_ubi_numCols)]
write.csv(ubiFeatPreds_byGene[, c("gene", "gene_site", "uniprot_site")], file = "output/data/ubiSitesGenes.csv") #Export for TF activity estimation

# Assign to list
CPTAC_ubi_list$LSCC <- list(
    "data" = CPTAC_ubi_raw,
    "CPTAC_ubi_numCols" = CPTAC_ubi_numCols, 
    "CPTAC_ubi_numColsTum"= CPTAC_ubi_numColsTum
)
```

### LUAD 2025
```{r}
#Read in
CPTAC_ubi_raw <- read_xlsx(paste0(home_directory, "data/supplementary_tables/Satpathy_CancerCell_2025/Table_S2.xlsx"), sheet = "S2D_ubiquitylome",
                            skip = 0)
##Correct identifiers
colnames(CPTAC_ubi_raw) <- gsub("C3N.", "C3N-", colnames(CPTAC_ubi_raw), fixed = T)   
colnames(CPTAC_ubi_raw) <- gsub("C3L.", "C3L-", colnames(CPTAC_ubi_raw), fixed = T)  
#Prep 
CPTAC_ubi_raw <- CPTAC_ubi_raw %>% 
            mutate(site = sites) %>% 
            mutate(gene_site = paste(geneSymbol, site, sep = "_")) 
# For the few cases with multiple gene_sites corresponding to missed cleavages, take the highest ReferenceIntensity
CPTAC_ubi_raw <- CPTAC_ubi_raw %>%
            group_by(gene_site) %>%
            slice_max(ReferenceIntensity, n = 1, with_ties = FALSE) %>%
            ungroup()

#Get numerical columns
CPTAC_ubi_numCols <- colnames(CPTAC_ubi_raw) %>% .[grep("^C3", .)]
CPTAC_ubi_numColsTum <- CPTAC_ubi_numCols[-grep("N$", CPTAC_ubi_numCols)]

# Assign to list
CPTAC_ubi_list$LUAD <- list(
    "data" = CPTAC_ubi_raw,
    "CPTAC_ubi_numCols" = CPTAC_ubi_numCols, 
    "CPTAC_ubi_numColsTum"= CPTAC_ubi_numColsTum
)
```

### Combine
```{r}
# Overview
map(CPTAC_ubi_list, ~.$CPTAC_ubi_numColsTum)
map(CPTAC_ubi_list, ~nrow(.$data))

# Combine
CPTAC_ubi_list$combined <- list(
    "data" = map(CPTAC_ubi_list[c("LSCC", "LUAD")], function(lst){
        datTrim <- lst$data[, c("gene_site", lst$CPTAC_ubi_numColsTum)]
        return(datTrim)
    }) %>% purrr::reduce(function(x, y){full_join(x, y, by = "gene_site")}),

    "CPTAC_ubi_numColsTum" = map(CPTAC_ubi_list[c("LSCC", "LUAD")], function(lst){lst$CPTAC_ubi_numColsTum}) %>% 
        unlist %>% unique
)
```



## Load CPTAC multi-omics 
These are generated by scripts/CPTAC/01_CPTAC_assemble.py and scripts/CPTAC/02_CPTAC_activity.R
```{r}
#Phospho
CPTACphos <- read_csv("output/data/CPTAC/phosphoproteomics_merged.csv") %>% as.data.frame
rownames(CPTACphos) <- CPTACphos$gene_site
CPTACphos <- CPTACphos[, -1]
CPTACidentif <- read_csv("output/data/CPTAC/phosphoproteomics_merged_identif.csv") %>% as.data.frame %>% .[, -1]

#Clinical
CPTACclin <- read_csv("output/data/CPTAC/clin_merged.csv") %>% as.data.frame
rownames(CPTACclin) <- CPTACclin$Patient_ID
CPTACclin$hasPhospho <- CPTACclin$Patient_ID %in% colnames(CPTACphos)

#Transcriptome
CPTACtrans <- read_csv("output/data/CPTAC/transcriptomics_merged.csv") %>% as.data.frame
CPTACtransIdentif <- CPTACtrans[, c("Name", "Database_ID")] %>% dplyr::rename("gene" = "Name", "ENSG" = "Database_ID")
rownames(CPTACtrans) <- CPTACtrans$Name
CPTACtrans <- CPTACtrans[, -c(1, 2)]
#Proteome
CPTACprot <- read_csv("output/data/CPTAC/proteomics_merged.csv") %>% as.data.frame
CPTACprotIdentif <- CPTACprot[, c("Name", "Database_ID")] %>% dplyr::rename("gene" = "Name", "ENSG" = "Database_ID")
rownames(CPTACprot) <- CPTACprot$Name
CPTACprot <- CPTACprot[, -c(1, 2)]

#Activity
TFactivity <- read_csv("output/data/CPTAC/TFactivity_m.csv") %>% as.data.frame
rownames(TFactivity) <- TFactivity[, 1]
TFactivity <- TFactivity[, -1]
KinactivityViper <- read_csv("output/data/CPTAC/kinase_activity_VIPER_m.csv") %>% as.data.frame
rownames(KinactivityViper) <- KinactivityViper[, 1]
KinactivityViper <- KinactivityViper[, -1]
```


## Check identifiers agree 
```{r}
map(CPTAC_ubi_list, function(lst){
    cols <- lst$CPTAC_ubi_numColsTum
    list(
        "numCols" = length(cols),
        "numOverlap" = length(intersect(cols, colnames(TFactivity)))
    )
})
```



## Perform correlation
```{r}
#Load correlation/annotation/filtering functions
source("scripts/CPTAC/03_CPTAC_correlation.R")

#Set up
activMatLst <- list("kinase" = KinactivityViper, "TF" = TFactivity)

#Run correlation, annotation and filtering for each CPTAC dataset
CPTAC_corr_results <- map(CPTAC_ubi_list, function(lst){
    run_CPTAC_correlation_pipeline(
        CPTAC_ubi_dat = lst$data,
        CPTAC_ubi_numColsTum = lst$CPTAC_ubi_numColsTum,
        activMatLst = activMatLst,
        ubiFeatPreds_byGene = ubiFeatPreds_byGene,
        CPTACprot = CPTACprot,
        hotspotSites = hotspotSites
    )
})

```


## Export
```{r}
# Export for supp table
activCorrResLst_dat <- map_dfr(CPTAC_corr_results$combined$filtered, function(x){x}, .id = "proteinType")
activCorrResLst_exp <- activCorrResLst_dat
renameCols <- c(
    "proteinType" = "Protein type",
    "functional_score" = "Functional score",
    "gene_site" = "ubi-site",
    "r_value" = "Correlation r-value",
    "p_value_adj" = "Adjusted p-value"
)
activCorrResLst_exp <- activCorrResLst_exp[, names(renameCols)] %>% 
    dplyr::rename(!!setNames(names(renameCols), renameCols))
outDir = "output/data/functional_scores/SuppTable_predFuncSites.xlsx"
sheet = "CPTAC_sites"
if (file.exists(outDir)){
  wb <- loadWorkbook(outDir)
} else {
  wb <- createWorkbook(outDir)
}
removeWorksheet(wb, sheet)
addWorksheet(wb, sheet)
writeData(wb, sheet = sheet, x = activCorrResLst_exp)
saveWorkbook(wb, outDir, overwrite = TRUE)
```


# Turnover and occupancy 
## Load and prep data 
```{r}
geo_mean <- function(x){exp(mean(log(x)))}
#Occupancy
ubiOccDat <- read_xlsx(paste0(home_directory, "data/supplementary_tables/Prus_Cell_2024/mmc1.xlsx"),
                        skip = 23) 
ubiOccDat <- ubiOccDat %>% #Separatea alternative protein IDs
                        separate_rows(c("Proteins", "Positions.within.Proteins"), sep = ";") %>% 
                        mutate(uniprot_pos = paste(Proteins, Positions.within.Proteins, sep = "_")) %>% 
                        group_by(uniprot_pos) %>% summarise(mean_Ub.occupancy = geo_mean(mean_Ub.occupancy)) #There are very few duplicate uniprot_sites, just average them
nrow(ubiOccDat)
length(which(ubiOccDat$uniprot_pos %in% ubiFeatPreds$uniprot_pos))
ubiFeatPreds <- left_join(ubiFeatPreds, ubiOccDat[, c("uniprot_pos", "mean_Ub.occupancy")], by = "uniprot_pos")
#Turnover
ubiTurnDat <- read_xlsx(paste0(home_directory, "data/supplementary_tables/Prus_Cell_2024/mmc2.xlsx"),
                        skip = 72) 
ubiTurnDat <- ubiTurnDat %>% 
                        dplyr::rename("half.life.min" = "half.life (min)") %>%
                        #Separatea alternative protein IDs
                        separate_rows(c("Proteins", "Positions.within.proteins"), sep = ";") %>% 
                        mutate(uniprot_pos = paste(Proteins, Positions.within.proteins, sep = "_")) %>% 
                        group_by(uniprot_pos) %>% summarise(half.life.min = geo_mean(half.life.min)) #There are very few duplicate uniprot_sites, just average them
nrow(ubiTurnDat)
length(which(ubiTurnDat$uniprot_pos %in% ubiFeatPreds$uniprot_pos))
ubiFeatPreds <- left_join(ubiFeatPreds, ubiTurnDat[, c("uniprot_pos", "half.life.min")], by = "uniprot_pos")
```


## Visualise
```{r} 
# Bin score
binBreaks <- funcScoreBreaks
binN <- length(binBreaks)
binCols <- brewer.pal(9, "Blues")[(9 - binN + 1):9]
binDat <- binner(data = subset(ubiFeatPreds, !is.na(functional_score)), 
                valCol = "functional_score", breakType = "float",
                bin_breaks = binBreaks)

# Violin plots
## Occupancy
bin1median <- subset(binDat, Bin == levels(Bin)[1])$mean_Ub.occupancy %>% median(., na.rm = T)
binDat %>%
ViolBoxplot(., vars = list(x = "Bin", y = "mean_Ub.occupancy", colour = "Bin", fill = "Bin"), includeLegend = F) +
        geom_hline(yintercept = bin1median, linewidth = 0.235, linetype = "dashed")  +
        scale_colour_manual(values = rep(brewer.pal(9, "Blues")[9], binN)) +
        scale_fill_manual(values = alpha(funcScoreBinCols, funcScoreFillAlpha)) +
        scale_y_log10() +
        labs(x = "Functional score", y = "Ubi-site occupancy") +
        ggtitle("Occupancy (Prus et al.)")
ggsave_pdfpng(file = "output/images/functional_scores/OccTurnover/viol_funcScore_occupancy",
                width = 2, height = 2)

## Turnover
bin1median <- subset(binDat, Bin == levels(Bin)[1])$half.life.min %>% median(., na.rm = T)
binDat %>%
ViolBoxplot(., vars = list(x = "Bin", y = "half.life.min", colour = "Bin", fill = "Bin"), includeLegend = F) +
        geom_hline(yintercept = bin1median, linewidth = 0.235, linetype = "dashed")  +
        scale_colour_manual(values = rep(brewer.pal(9, "Blues")[9], binN)) +
        scale_fill_manual(values = alpha(funcScoreBinCols, funcScoreFillAlpha)) +
        scale_y_log10() +
        labs(x = "Functional score", y = "Ubi-site half-life (min)")  +
        ggtitle("Turnover (Prus et al.)")
ggsave_pdfpng(file = "output/images/functional_scores/OccTurnover/viol_funcScore_halflife",
                width = 2, height = 2)
```




# Clinvar genetic variants
This is generates by scripts/12_prep_genetic_variants.R
```{r}
load("output/data/genetics/clinvarDatSub__2.RData")
# Subset for missense variants
clinvarDatSub <- subset(clinvarDatSub, grepl("missense_variant", MC))
dim(clinvarDatSub)
length(unique(clinvarDatSub$uniprot_site))
# Map functinal scores in
clinvarDatSub <- left_join(clinvarDatSub, ubiFeatPreds[, c("uniprot_site", "functional_score", "funcScoreQuant", "mutR_am_pathogenicity_all", "PSP_reg", "PSP_reg_notProtDeg")], by = "uniprot_site")
clinvarDatSub <- clinvarDatSub %>% mutate(clinvar_label = factor(clinvar_label, levels = c("Benign", "VUS", "Pathogenic")))
levels(clinvarDatSub$clinvar_label)[2] <- "Uncertain" 
View(head(clinvarDatSub))

# Enrichment
fishers_DE_pathway(background_genes = unique(clinvarDatSub$uniprot_site), 
                   pathway_genes = subset(clinvarDatSub, clinvar_label == "Pathogenic")$uniprot_site %>% unique,
                   DE_genes = subset(clinvarDatSub, functional_score > 0.65)$uniprot_site %>% unique)

# Visualise
## Flag mutations to plot
mutsToLabel <- c("Q14683/K536R")
labelDat <- subset(clinvarDatSub, variant_id %in% mutsToLabel) %>% 
            mutate("uniprot" = gsub("_.+$", "", uniprot_site)) %>%
            left_join(., distinct(ubiFeatPreds_byGene[, c("uniprot", "gene")]), by = "uniprot") %>% 
            mutate(mut = strsplit(variant_id, "/") %>% map(~.[2]) %>% unlist) %>%
            mutate(geneMut = paste(gene, mut, sep = " "))
## Violin
clinvarCols <- c("#f7fcfe", "#6badd6", "#203569")
### FUnctional score
clinvarDatSub %>%
ViolBoxplot(., vars = list(x = "clinvar_label", y = "functional_score", colour = "clinvar_label", fill = "clinvar_label"),
                includeLegend = F) + 
    geom_point(data = labelDat, aes(x = clinvar_label, y = functional_score),
               shape = 21, size = 1, stroke = 0.71*0.25, colour = "black", fill = alpha(clinvarCols[3], 0.8), 
               inherit.aes = F) + 
    geom_text(data = labelDat, aes(x = clinvar_label, y = functional_score, label = geneMut), 
              colour = "#3f3f3e", size = 1.875*5/5) +
    scale_colour_manual(values = rep("#203569", 3)) +
    scale_fill_manual(values = alpha(clinvarCols, 0.25)) + 
    labs(x = "Pathogenicity", y = "Functional score") +
    ggtitle("ClinVar missense variants")
ggsave_pdfpng(file = "output/images/functional_scores/genetics/viol_clinvar_funcScore",
                width = 2, height = 2)
#### Mann-Whitney U-tests vs Benign
wilcox.test(subset(clinvarDatSub, clinvar_label == "Benign")$functional_score,
            subset(clinvarDatSub, clinvar_label == "Uncertain")$functional_score)
wilcox.test(subset(clinvarDatSub, clinvar_label == "Benign")$functional_score,
            subset(clinvarDatSub, clinvar_label == "Pathogenic")$functional_score)$p.value
### Score quantile within protein
clinvarDatSub %>%
ViolBoxplot(., vars = list(x = "clinvar_label", y = "funcScoreQuant", colour = "clinvar_label", fill = "clinvar_label"),
                includeLegend = F) + 
    geom_point(data = labelDat, aes(x = clinvar_label, y = functional_score),
               shape = 21, size = 1, stroke = 0.71*0.25, colour = "black", fill = alpha(clinvarCols[3], 0.8), 
               inherit.aes = F) + 
    geom_text(data = labelDat, aes(x = clinvar_label, y = functional_score, label = geneMut), 
              colour = "#3f3f3e", size = 1.875*5/5) +
    scale_colour_manual(values = rep("#203569", 3)) +
    scale_fill_manual(values = alpha(clinvarCols, 0.25)) + 
    labs(x = "Pathogenicity", y = "Score percentile within protein") +
    ggtitle("ClinVar missense variants")
ggsave_pdfpng(file = "output/images/functional_scores/genetics/viol_clinvar_funcScoreQuant",
                width = 2, height = 2)
#### Mann-Whitney U-tests vs Benign
wilcox.test(subset(clinvarDatSub, clinvar_label == "Benign")$funcScoreQuant,
            subset(clinvarDatSub, clinvar_label == "Uncertain")$funcScoreQuant)
wilcox.test(subset(clinvarDatSub, clinvar_label == "Benign")$funcScoreQuant,
            subset(clinvarDatSub, clinvar_label == "Pathogenic")$funcScoreQuant)

# Collapse at the site level and add to feature data
clinvarDatSub_coll <- clinvarDatSub %>%
    group_by(uniprot_site) %>%
    summarise(across(.cols = colnames(.)[-which(colnames(.) %in% c("uniprot_site"))], ~ paste(., collapse = ";")),  .groups = "drop")
ubiFeatPreds <- left_join(ubiFeatPreds, clinvarDatSub_coll[, c("uniprot_site", "Amino_acid_change", "clinvar_label", "CLNDN", "CLNREVSTAT", "CLNREVSTAT_STARS")])
## Flag arginine changes that are pathogenic
ubiFeatPreds$ArgChangePatho <- ubiFeatPreds$uniprot_site %in% subset(clinvarDatSub, clinvar_label == "Pathogenic" & grepl("Arg", Amino_acid_change))$uniprot_site

# Export for inspection
subset(ubiFeatPreds, grepl("Pathogenic", clinvar_label)) %>% .[order(.$functional_score, decreasing = T), ] %>% 
    dplyr::relocate(c("functional_score", "ArgChangePatho", "Amino_acid_change", "BortMG_meanlog2FC", "HotspotNew_bool", "gene", "interface_label", "interaction_id", "PSP_reg", "PSP_reg_ProtDeg")) %>% 
    write.csv(., file = "output/data/genetics/ClinvarPathogenicSites.csv", na = "")
```


# Lysine base editor screen (Bao et al 2023)
## Load and prep data
```{r}
LysBEdat_raw <- read_xlsx(path = paste0(home_directory, "data/supplementary_tables/Bao_MolCell_2023/Table S2-Day24_Day0_screen results.xlsx"), sheet = "Table S3")
LysBEdat <- LysBEdat_raw %>% as.data.frame

# Subset for lysines
LysBEdat <- subset(LysBEdat, type %in% c("Depleted_Lysine", "Enriched_Lysine", "Lysine"))
LysBEdat$type <- factor(LysBEdat$type, levels = c("Lysine", "Depleted_Lysine", "Enriched_Lysine"))
LysBEdat$sig <- LysBEdat$type != "Lysine"

# Get site ID
## Clean up redundant IDs for the same guide
LysBEdat$aa_id <- strsplit(LysBEdat$aa_id, ";") %>% map(~paste(unique(.), collapse = ";")) %>% unlist
## Split by aa_id
LysBEdat_perSite <- separate_rows(LysBEdat, "aa_id", sep = ";")
## Filter for multiple values of type
multTypeSites <- LysBEdat_perSite %>%
    group_by(aa_id) %>%
    filter(n_distinct(type) > 1)
multTypeSites %>% .[order(.$aa_id), ] %>% View

# Map into our data
## Overview
length(which(ubiFeatPreds$uniprot_pos %in% LysBEdat_perSite$aa_id))
## Subset their data and add our info
## This deals with the fact that they have multiple rows per site
LysBEdat_funcScore <- subset(LysBEdat_perSite, aa_id %in% ubiFeatPreds$uniprot_pos) %>% 
    left_join(., ubiFeatPreds[, c("uniprot_pos", "functional_score", "funcScoreQuant")] %>% dplyr::rename("aa_id" = "uniprot_pos"), by = "aa_id")
## How many have phenotypes?
table(LysBEdat_funcScore$type)

# Collapse to the individual site level
# When one site has multiple classes, assign in this order: Lysine -> Enriched_Lysine -> Depleted_Lysine 
LysBEdat_funcScore_sites <- LysBEdat_funcScore[, c("type", "functional_score", "funcScoreQuant", "aa_id")] %>% 
    .[order(.$type, decreasing = T), ] %>%  # This ensures the ordering above
    group_by(aa_id) %>% summarise(type = type[1], functional_score = functional_score[1], funcScoreQuant = funcScoreQuant[1]) %>% 
    ungroup %>% 
    mutate(typeLabel = factor(type, levels = c("Lysine", "Depleted_Lysine", "Enriched_Lysine"), labels = c("No effect", "Deleterious", "Beneficial")))
LysBEdat_funcScore_sites$sig <- LysBEdat_funcScore_sites$type != "Lysine"
LysBEdat_funcScore_sites$uniprot <- gsub("_.+$", "", LysBEdat_funcScore_sites$aa_id)
## Export
save(LysBEdat_funcScore_sites, file = "output/data/genetics/LysBEdat_funcScore_sites.RData")


# Add to our data
toAdd <- LysBEdat_funcScore_sites %>% 
    dplyr::rename("uniprot_pos" = "aa_id", "BE_type" = "type", "BE_typeLabel" = "typeLabel", "BE_sig" = "sig") %>% 
    dplyr::select(-c("uniprot", "functional_score", "funcScoreQuant"))
if ("BE_type" %in% colnames(ubiFeatPreds) == F){
    ubiFeatPreds <- left_join(ubiFeatPreds, toAdd, by = "uniprot_pos") %>% 
        mutate("BE_tested" = !is.na(BE_type))
}
```

## Compare functional scores across phenotype groups
```{r}
# Plot
lysBECols <- c("#f7fcfe", "#6badd6", "#203569")
## Funcitonal score
LysBEdat_funcScore_sites %>%
ViolBoxplot(., vars = list(x = "typeLabel", y = "functional_score", colour = "typeLabel", fill = "typeLabel"),
                includeLegend = F) + 
    scale_colour_manual(values = rep("#203569", 3)) +
    scale_fill_manual(values = alpha(lysBECols, 0.25)) + 
    labs(x = "Base edit effect on viability", y = "Functional score") +
    ggtitle("Lys base editor screen")
ggsave_pdfpng(file = "output/images/functional_scores/Bao_baseEdit/viol_BaoLys_funcScore",
                width = 1.75, height = 2)
#### Mann-Whitney U-tests vs No effect
wilcox.test(subset(LysBEdat_funcScore_sites, typeLabel == "No effect")$functional_score,
            subset(LysBEdat_funcScore_sites, typeLabel == "Deleterious")$functional_score)
wilcox.test(subset(LysBEdat_funcScore_sites, typeLabel == "No effect")$functional_score,
            subset(LysBEdat_funcScore_sites, typeLabel == "Beneficial")$functional_score)
## Score quantile within proteins
LysBEdat_funcScore_sites %>%
ViolBoxplot(., vars = list(x = "typeLabel", y = "funcScoreQuant", colour = "typeLabel", fill = "typeLabel"),
                includeLegend = F) + 
    scale_colour_manual(values = rep("#203569", 3)) +
    scale_fill_manual(values = alpha(lysBECols, 0.25)) + 
    labs(x = "Base edit effect on viability", y = "Score percentile within protein") +
    ggtitle("Lys base editor screen")
ggsave_pdfpng(file = "output/images/functional_scores/Bao_baseEdit/viol_BaoLys_funcScoreQuant",
                width = 1.75, height = 2)
#### Mann-Whitney U-tests vs No effect
wilcox.test(subset(LysBEdat_funcScore_sites, typeLabel == "No effect")$funcScoreQuant,
            subset(LysBEdat_funcScore_sites, typeLabel == "Deleterious")$funcScoreQuant)
wilcox.test(subset(LysBEdat_funcScore_sites, typeLabel == "No effect")$funcScoreQuant,
            subset(LysBEdat_funcScore_sites, typeLabel == "Beneficial")$funcScoreQuant)

```