# Background
Here I build a pipeline to assess model accuracy and major results.
This can be applied to different trainings of the model and individual features.


# Initialise
```{r}
library(tidyverse)
library(purrr)
library(ggplot2)
library(ggridges)
library(readxl)
library(openxlsx)
library(data.table)
library(caret)

home_directory <- ""
source(paste0(home_directory, "scripts/essential_visualisations_JvG.R"))
source(paste0(home_directory, "scripts/enrichment_analysis_JvG.R"))
source(paste0(home_directory, "scripts/caret_JvG.R"))
source("scripts/01_general.R")
source("scripts/00_objects.R")
source("scripts/03_visualisations.R")
source("scripts/04_ubnOmics.R")
source("scripts/02_model_training.R")
source("scripts/16_model_accuracy.R") # Auxiliary functions for assessing accuracy

# Load features
load("data/features/ubnOmics/ubnOmicsFeauresAll__4.RData")


# Load clinvar
load("output/data/genetics/clinvarDatSub__2.RData")
clinvarDatSub <- subset(clinvarDatSub, grepl("missense_variant", MC))
clinvarDatSub <- clinvarDatSub %>% mutate(clinvar_label = factor(clinvar_label, levels = c("Benign", "VUS", "Pathogenic")))
levels(clinvarDatSub$clinvar_label)[2] <- "Uncertain" 

# Load base editing
load("output/data/genetics/LysBEdat_funcScore_sites.RData")
LysBEdat_funcScore_sites$pos <- gsub("^.+_" ,"", LysBEdat_funcScore_sites$aa_id)

# Load NLS
load("output/data/functional_scores/ubiFeatPreds_NLS__7.RData")

## Load protein abundance data
load("data/features/abundDat__1.RData")
abundDat$uniprot_pos <- gsub("_K", "_", rownames(abundDat))
abundDat$uniprot <- gsub("_.+$", "", abundDat$uniprot_pos)
abundFeatsLog <- colnames(abundDat) %>% .[grep("log$", .)]
protAbundDat <- abundDat[, c("uniprot", "protAbundance_log")] %>% distinct
```


# Load data to test
## Load current set predictions
Here suffix indicates the latest model
```{r}
# Load
suffix <- "attempt11_revisions"
load(paste0("output/data/functional_scores/modsPreds_trainTestforPred_df__", suffix, ".RData"))
testSetPreds <- modsPreds_trainTestforPred_df 
testSetPreds$uniprot <- gsub("_[A-Z].+$", "", testSetPreds$uniprot_site)
testSetPreds$uniprot_pos <- gsub("_[A-Z]", "_", testSetPreds$uniprot_site)
```

# Prepare external data to add
This is done at the level of uniprot_site, so it can be done once and then mapped to multiple datasets more quickly
```{r}
# Set up data
siteData <- distinct(testSetPreds[, c("uniprot", "uniprot_site", "uniprot_pos")])

# Add prtsm inhib data
prtsmInhibCols <- c("Bortezomib_meanlog2FC", "MG-132_meanlog2FC", "Epoxomicin_meanlog2FC")
if (length(intersect(prtsmInhibCols, colnames(siteData))) == 0){
    siteData <- left_join(siteData, ubnOmicsFeaturesAll[, c("uniprot_pos", prtsmInhibCols)],
                                by = "uniprot_pos")
}

# Add clinvar data
useStars <- FALSE
CVdatFilt <- clinvarDatSub # Optional filtering here
siteData$clinvarClass <- NA
levels <- c("Uncertain", "Benign", "Pathogenic") # This degines the order of evidence. Pathogenic trumps all
if (useStars){
    for (lev in levels){
        siteData$clinvarClass[which(siteData$uniprot_site %in% subset(CVdatFilt, clinvar_label == lev & CLNREVSTAT_STARS == 1)$uniprot_site)] <- paste0(lev, "_1star")
        siteData$clinvarClass[which(siteData$uniprot_site %in% subset(CVdatFilt, clinvar_label == lev & CLNREVSTAT_STARS > 1)$uniprot_site)] <- paste0(lev, "_2+star")
    }
    siteData$clinvarClass <- factor(siteData$clinvarClass, levels = sort(unique(siteData$clinvarClass[which(!is.na(siteData$clinvarClass))])))
} else {
    for (lev in levels){
        siteData$clinvarClass[which(siteData$uniprot_site %in% subset(CVdatFilt, clinvar_label == lev)$uniprot_site)] <- lev
    }
    siteData$clinvarClass <- factor(siteData$clinvarClass, levels = levels)
}

# Add base editing data
siteData <- left_join(
    siteData, 
    LysBEdat_funcScore_sites[, c("aa_id", "typeLabel")] %>% dplyr::rename("uniprot_pos" = "aa_id", "BE_typeLabel" = "typeLabel")
    )

# Add abundance
siteData <- left_join(
    siteData, 
    abundDat
    )

# Add NLS data
siteData <- left_join(siteData, ubiFeatPreds_NLS)
```

## Sites of interest
```{r}
sitesOfInterest <- list(
    "ELAVL1" = "Q15717_K320"
) %>% map_dfr(function(x){data.frame("uniprot_site" = x)}, .id = "source") %>% 
    mutate(uniprot = gsub("_.+$", "", uniprot_site))
#Get gene identifiers 
library(AnnotationDbi)
library(org.Hs.eg.db)
collapseFunc <- function(x){
    x <- unique(x)
    if (is.na(x[1])){
        return(NA)
    } else {
        return(paste(x, collapse = ";"))
    }
}
geneMap <- AnnotationDbi::select(org.Hs.eg.db, keys = unique(sitesOfInterest$uniprot), keytype = "UNIPROT",
                                    columns = c("SYMBOL")) %>% 
                                group_by(UNIPROT) %>% 
                                summarise(gene = collapseFunc(SYMBOL))
sitesOfInterest <- left_join(sitesOfInterest, dplyr::rename(geneMap, "uniprot" = "UNIPROT"))
```


# Function: Run the analysis pipeline
```{r}
# Function: Get type of metric from its name
getMetricName <- function(x){
    if (length(grep("AUC", x)) > 0){
        return("AUC")
    } else if (length(grep("MAD", x)) > 0){
        return("MAD")
    } else if (length(grep("_r$", x)) > 0){
        return("r")
    } else {
        return(x)
    }
}

# Evaluate predictions across various criteria
# dataset: Must contain Resample as a column
predictionEvaluationPipeline <- function(dataset, predCol = "X1"){
    ##### Compute columns and add features to data #####
    # Add identifier columns
    dataset$uniprot <- gsub("_[A-Z].+$", "", dataset$uniprot_site)
    dataset$uniprot_pos <- gsub("_[A-Z]", "_", dataset$uniprot_site)
    # Assign prediction column
    dataset$predCol <- dataset[[predCol]]
    # Quantile-normalise within proteins
    dataset <- dataset %>% group_by(uniprot, Resample) %>% 
        mutate(
            predColQuant = to_quantiles(predCol), 
            predColRank = rank(-predCol),
            numSites = n()) %>% ungroup 
    
    # Add features
    cols_to_add <- setdiff(colnames(siteData), colnames(dataset))
    if (length(cols_to_add) > 0){
        dataset <- left_join(dataset, siteData[, c(cols_to_add, "uniprot_site")], by = "uniprot_site")
    }
    # Add column for negative trainign class and not benign
    dataset$negClassNotBenign <- dataset$obs == "X0" & (dataset$clinvarClass != "Benign" | is.na(dataset$clinvarClass))
    # Add column for pheno in base editing
    dataset$BE_hasPheno <- dataset$BE_typeLabel %in% c("Deleterious", "Beneficial")

    ##### Compute metrics #####
    # Calculate model accuracy
    AUCresLst <- calculateAUC_withinAndBetweenProts(dat = dataset, predCol = "predCol")
    AUCresSumm <- summarise_AUC_withinAndBetweenProts(resLst = AUCresLst, metric = "AUC")

    # Calculate MAD for differences within and between proteins
    MADres <- calculateMAD_withinAndBetweenProts(dataset, predCol = "predCol", removeSingles = T)
    MADresSumm <- summarise_AUC_withinAndBetweenProts(MADres, metric = "MAD")

    # Association with proteasome inhibition logFC
    prtmsiCorr <- dataset %>% group_by(Resample) %>% 
        summarise(
            Bort_r = cor(Bortezomib_meanlog2FC, predCol, use = "complete.obs"),
            MG132_r = cor(`MG-132_meanlog2FC`, predCol, use = "complete.obs"),
            Epox_r = cor(Epoxomicin_meanlog2FC, predCol, use = "complete.obs")
        ) %>% 
        ungroup

    # Association with protein abunadnce
    abundCorr <- dataset %>% group_by(Resample) %>% 
        summarise(protAbund_r = cor(protAbundance_log, predCol, use = "complete.obs")) %>% ungroup

    # Various binary classification tasks
    # Set up comparisons: each tuple specifies a column and the corresponding value used for filtering to produce the positive class or negative class
    comps <- list(
        "CV_PathVsBenign" = list(list("clinvarClass", "Pathogenic"), list("clinvarClass", "Benign")),
        "CV_NegclassVsBengin" = list(list("negClassNotBenign", TRUE), list("clinvarClass", "Benign")),
        "BE_PhenoVsNone" = list(list("BE_hasPheno", TRUE), list("BE_typeLabel", "No effect")),
        "NLS_InVsOut" = list(list("inNLS", TRUE), list("inNLS", FALSE))
    )
    for (i in 1:length(comps)){
        compName <- names(comps)[i]
        comp <- comps[[i]]
        dataset[[compName]] <- NA
        dataset[[compName]][which(dataset[[comp[[1]][[1]]]] == comp[[1]][[2]])] <- "X1"
        dataset[[compName]][which(dataset[[comp[[2]][[1]]]] == comp[[2]][[2]])] <- "X0"
        dataset[[compName]] <- factor(dataset[[compName]], levels = c("X1", "X0"))
    }
    # Calculate AUCs across resamples
    binaryAUCs <- dataset %>% split(.$Resample) %>% map_dfr(function(dat){
        AUCs <- map(names(comps), ~ROCAUC_fromPredictor(data = dat, obsCol = ., predCol = "predCol"))
        AUCs_protQuants <- map(names(comps), ~ROCAUC_fromPredictor(data = dat, obsCol = ., predCol = "predColQuant"))
        data.frame("AUC" = unlist(AUCs), "AUC_protQuants" = unlist(AUCs_protQuants), "name" = names(comps))
    }, .id = "Resample") %>% pivot_wider(names_from = "name", values_from = c("AUC", "AUC_protQuants"))

    # Combine metrics and summarise
    combinedMetrics <- pivot_wider(AUCresSumm, names_from = "Regime", values_from = "AUC") %>% 
        left_join(., pivot_wider(MADresSumm, names_from = "Regime", values_from = "MAD")) %>% 
        left_join(., prtmsiCorr) %>% 
        left_join(abundCorr) %>% 
        left_join(., binaryAUCs)
    combinedMetricsSumm <- combinedMetrics %>% pivot_longer(cols = setdiff(colnames(.), "Resample"), names_to = "metric", values_to = "value") %>% 
        group_by(metric) %>% summarise(mean = mean(value, na.rm = T), sd = sd(value, na.rm = T)) %>% ungroup %>% 
        mutate(metric = factor(metric, levels = intersect(colnames(combinedMetrics), unique(metric)))) %>% 
        .[order(.$metric), ]

    # Sites of interest
    sitesOfInterest_wRes <- left_join(sitesOfInterest, dataset[, c("uniprot_site", "predCol", "predColQuant", "predColRank", "numSites", "Resample")], by = "uniprot_site")
    sitesOfInterest_wResSumm <- sitesOfInterest_wRes %>%
    dplyr::select(-c("Resample")) %>% 
    group_by(uniprot_site) %>%
    summarise(
        across(where(is.numeric), ~mean(., na.rm = TRUE)),
        across(where(is.character), ~unique(.)),
        .groups = "drop"
    )

    # Final results
    resultsLst <- list(
        "combinedMetrics" = combinedMetrics, 
        "combinedMetricsSumm" = combinedMetricsSumm,
        "sitesOfInterest_wRes" = sitesOfInterest_wRes,
        "sitesOfInterest_wResSumm" = sitesOfInterest_wResSumm
    )
    return(resultsLst)
}
```


# Test individual features
```{r}
suffix <- "attempt11_revisions"

## Load coefficients
load(paste0("output/data/functional_scores/finalModel_logregs_coeffs__", suffix, ".RData"))

# Prepare individual features
load("data/functional_scores/datFeats__7.RData")
selFeats <- intersect(names(finalModel_logregs_coeffs[[1]]), colnames(datFeats))
testDf_comb <- left_join(testSetPreds, rownames_to_column(datFeats[, selFeats], "uniprot_site"), by = "uniprot_site") %>% 
    dplyr::rename("functional_score" = "X1")

# Run through evaluation
evaluationLstFeats <- list()
for (col in c("functional_score", selFeats)){
    res <- predictionEvaluationPipeline(dataset = testDf_comb, predCol = col)
    evaluationLstFeats[[col]] <- res
}
## Combine summaries
summNames <- names(evaluationLstFeats[[1]]) %>% .[grep("Summ$", .)]
evaluationCombLstFeats <- map(summNames, function(summName){
    dat <- map_dfr(evaluationLstFeats, function(x){x[[summName]]}, .id = "model")
    return(dat)
})
names(evaluationCombLstFeats) <- summNames
save(evaluationCombLstFeats, file = "output/data/functional_scores/featsIndivAblate/evaluationCombLstFeats.RData")
```

## Inspect results
```{r}
# AUCs
AUCdat <- evaluationCombLstFeats$combinedMetricsSumm %>% 
    subset(metric == "AUCoverall") 
AUCdat$AUCabs <- ifelse(AUCdat$mean < 0.5, 1 - AUCdat$mean, AUCdat$mean)
# Clinvar pathogenic
evaluationCombLstFeats$combinedMetricsSumm %>% 
    subset(metric == "AUC_CV_PathVsBenign") %>% 
    View
```


## Plot results
```{r}
for (selMetric in unique(evaluationCombLstFeats$combinedMetricsSumm$metric)){
    subDat <- subset(evaluationCombLstFeats$combinedMetricsSumm, metric == selMetric) 
    # Get metric label
    metricLabel <- getMetricName(selMetric)
    includeVline <- metricLabel == "AUC"
    # Plot
    plt <- bar_meanSd_features(
        data = subDat, meanCol = "mean", sdCol = "sd", featCol = "model", orderByMean = T,
        includeVline = includeVline, xLabel = metricLabel) +
            ggtitle(selMetric) + labs(y = "Feature")
    plot(plt)
    ggsave_pdfpng(file = paste0("output/images/functional_scores/featsIndivAblate/bar_", selMetric), width = 3, height = 2.5)
}
```

# Ablate individual feaures
## Retrain models without each feature
```{r}
# Set up
resDir <- "output/data/functional_scores/retrain1BFeatAblate/"
modNames <- dir(resDir) %>% .[grepl("^modsPredsTest_df_", .)] %>% gsub("modsPredsTest_df_", "", .) %>% gsub(".RData", "", ., fixed = T) %>% rev

# Load training results 
testDfLst <- list()
for (modName in modNames){
    load(paste0(resDir, "modsPredsTest_df_", modName, ".RData"))
    testDfLst[[modName]] <- modsPredsTest_df
}
testDfLst$functional_score <- testSetPreds # Add the score itself


# Run through evaluation
evaluationlst <- map(testDfLst, function(testDf){
    predictionEvaluationPipeline(dataset = testDf, predCol = "X1")
})
## Combine summaries
summNames <- names(evaluationlst[[1]]) %>% .[grep("Summ$", .)]
evaluationCombLst <- map(summNames, function(summName){
    dat <- map_dfr(evaluationlst, function(x){x[[summName]]}, .id = "model")
    return(dat)
})
names(evaluationCombLst) <- summNames
save(evaluationCombLst, file = paste0(resDir, "evaluationCombLst.RData"))
#load(paste0(resDir, "evaluationCombLst.RData"))
```

### Inspect results
```{r}
# Clinvar pathogenic
evaluationCombLst$combinedMetricsSumm %>% 
    subset(metric == "AUC_CV_PathVsBenign") %>% 
    View
```

### Plot results
```{r}
for (selMetric in unique(evaluationCombLst$combinedMetricsSumm$metric)){
    subDat <- subset(evaluationCombLst$combinedMetricsSumm, metric == selMetric) 
    # Get metric label
    metricLabel <- getMetricName(selMetric)
    includeVline <- metricLabel == "AUC"
    # Plot
    plt <- bar_meanSd_features(
        data = subDat, meanCol = "mean", sdCol = "sd", featCol = "model", orderByMean = T,
        includeVline = includeVline, xLabel = metricLabel) +
            ggtitle(selMetric) + labs(y = "Removed feature")
    plot(plt)
    ggsave_pdfpng(file = paste0("output/images/functional_scores/featsIndivAblate/bar_featRemove_", selMetric), width = 3, height = 2.5)
}
```