# Background
Here I analyse factors that could be confounding the score and its association with other things of interest.
In particular, I look at protein abundance and protein turnover


# Initialise
```{r}
library(tidyverse)
library(purrr)
library(ggplot2)
library(ggridges)
library(readxl)
library(openxlsx)

home_directory <- ""
source(paste0(home_directory, "scripts/essential_visualisations_JvG.R"))
source(paste0(home_directory, "scripts/enrichment_analysis_JvG.R"))
source("scripts/01_general.R")
source("scripts/00_objects.R")
source("scripts/04_ubnOmics.R")

#Load data
load("data/functional_scores/ubiFeatPreds__7.RData")
load("data/features/ubnOmics/ubnOmicsFeauresAll__4.RData")
load("data/features/ubnOmics/ubnOmicsComp_hsPRIDE_lst__8.RData")


#Set up bin breaks for the functional score, and other aesthetic paramters
funcScoreBreaks <- c(0, 0.2, 0.4, 0.6, 0.8, 1)
funcScoreBinCols <- brewer.pal(9, "Blues")[c(1, 3, 5, 7, 9)]
funcScoreFillAlpha <- 0.3

#Get gene identifiers for our ubi sites
library(AnnotationDbi)
library(org.Hs.eg.db)
collapseFunc <- function(x){
    x <- unique(x)
    if (is.na(x[1])){
        return(NA)
    } else {
        return(paste(x, collapse = ";"))
    }
}
geneMap <- AnnotationDbi::select(org.Hs.eg.db, keys = unique(ubiFeatPreds$uniprot), keytype = "UNIPROT",
                                    columns = c("SYMBOL")) %>% 
                                group_by(UNIPROT) %>% 
                                summarise(gene = collapseFunc(SYMBOL))
ubiFeatPreds <- left_join(ubiFeatPreds, dplyr::rename(geneMap, "uniprot" = "UNIPROT"), by = "uniprot")
##Expand out by gene when they are not unique
ubiFeatPreds_byGene <- separate_rows(ubiFeatPreds, "gene", sep = ";") %>% 
                        mutate(gene_site = paste(gene, site, sep = "_")) 
```

## Add proteasome inhibition data
```{r}
# Add prtsm inhib data
prtsmInhibCols <- c("Bortezomib_meanlog2FC", "MG-132_meanlog2FC", "Epoxomicin_meanlog2FC")
if (length(intersect(prtsmInhibCols, colnames(ubiFeatPreds))) == 0){
    ubiFeatPreds <- left_join(ubiFeatPreds, ubnOmicsFeaturesAll[, c("uniprot_pos", prtsmInhibCols)],
                                by = "uniprot_pos")
}
```

# Protein and site abundance 
## Prep data
```{r}
load("data/features/abundDat__1.RData")
abundDat$uniprot_pos <- gsub("_K", "_", rownames(abundDat))
abundFeatsLog <- colnames(abundDat) %>% .[grep("log$", .)]
ubiFeatPreds <- left_join(ubiFeatPreds, abundDat[, c(abundFeatsLog, "uniprot_pos")], by = "uniprot_pos")
```

## log-transform features so they match training
```{r}
logFeats <- c(
  "mutR_am_pathogenicity_all",
  
  "PSP_numAdj21",
  
  "numQuant_all",
  
  "numReg_notPrtsm"

  
)

for (col in logFeats){
  if (sum(ubiFeatPreds[, col] == 0, na.rm = TRUE) > 0){
    ubiFeatPreds[, col] <- log(1 + ubiFeatPreds[, col]) #Avoid NA for 0 values
  } else {
    ubiFeatPreds[, col] <- log(ubiFeatPreds[, col])
  }
}
```

## Correlate the score and model features with abundance measures
```{r}
# Set up aesthetics
library(viridis)
colScale <- c(seq(0, 0.25, length.out = 6 + 1)[-1], 
            seq(0.25, 0.7, length.out = 5 + 1)[-1], 1) 
hexScale <- scale_fill_gradientn(colours = viridis(10), 
                        values = rescale(colScale, c(0, 1)), 
                        limits = c(1, NA), name = "# sites")

# Define features
abunMets <- c("protAbundance_log", "PSMcounts_log")
feats <- c("functional_score", featureCols)

# Helper functions
removeExtremeOutliers <- function(dat, vars, sdCutoff = 10){
    keep <- rep(TRUE, nrow(dat))
    for (var in vars){
        x <- dat[[var]]
        if (!is.numeric(x)){
            next
        }
        xMean <- mean(x, na.rm = TRUE)
        xSd <- sd(x, na.rm = TRUE)
        if (is.na(xSd) || xSd == 0){
            next
        }
        keep <- keep & (is.na(x) | abs(x - xMean) <= sdCutoff * xSd)
    }
    dat[keep, ]
}

isBooleanVar <- function(x){
    x <- x[!is.na(x)]
    if (length(x) == 0){
        return(FALSE)
    }
    all(unique(x) %in% c(TRUE, FALSE, 0, 1))
}

# Plot
for (abunMet in abunMets){
    pltLst <- list()
    for (feat in feats){
        # Remove extreme outliers for the variables being plotted
        plotDat <- removeExtremeOutliers(ubiFeatPreds, c(abunMet, feat))

        # Plot continuous features as hex plots and boolean features as boxplots
        if (isBooleanVar(plotDat[[feat]])){
            plotDat[[feat]] <- factor(plotDat[[feat]])
            plt <- ggplot(plotDat, aes(y = !!sym(abunMet), x = !!sym(feat))) +
                geom_boxplot(outlier.shape = NA, size = 0.235) +
                comfy_theme()
        } else {
            corVal <- cor(plotDat[[abunMet]], plotDat[[feat]], use = "complete.obs")
            corLabel <- paste0("R = ", signif(corVal, 3))
            plt <- ggplot(plotDat, aes(y = !!sym(abunMet), x = !!sym(feat))) + 
                geom_hex(bins = 50) +
                geom_smooth(method = "lm", se = TRUE, color = "red", linewidth = 0.235, alpha = 0.3) +
                annotate("text", x = Inf, y = Inf, label = corLabel, hjust = 1.1, vjust = 1.1, size = 4) +
                hexScale +
                comfy_theme()
        }
        pltLst[[feat]] <- plt
    }
    ggarrange(plotlist = pltLst)
    ggsave_pdfpng(file = paste0("output/images/functional_scores/abundance/hex_modelFeats_", abunMet),
                    width = 2.25*5, height = 2*4)
}
```

## Bin protein abundance and compare to prtsmi 
```{r}
# Bin protein abundance
binN <- 10
ubiFeatPreds$protAbundance_log_bins <- cut_number(ubiFeatPreds$protAbundance_log, binN)
ubiFeatPreds$protAbundance_log_binsLab <- ubiFeatPreds$protAbundance_log_bins
levels(ubiFeatPreds$protAbundance_log_binsLab) <- paste0("Abun", 1:binN)

# Visualise score distribution in each bin
ggplot(ubiFeatPreds, aes(x = functional_score)) + 
    geom_histogram() + 
    comfy_theme() + 
    facet_wrap(~protAbundance_log_binsLab)
```



## Bin functional score and compare to prtsmi, coloured by abundance bin
```{r}
# Bin functional_score into 5 fixed-width bins
funcScoreBinBreaks <- c(0, 0.2, 0.4, 0.6, 0.8, 1.0)
ubiFeatPreds$functional_score_bins <- cut(ubiFeatPreds$functional_score, breaks = funcScoreBinBreaks, include.lowest = TRUE)
ubiFeatPreds$functional_score_binsLab <- ubiFeatPreds$functional_score_bins
levels(ubiFeatPreds$functional_score_binsLab) <- paste0("Score", 1:5)

# FC boxplot function (duplicated from 09_regulation_vs_conservation.Rmd)
box_FCs <- function(
  dat,
  logFCcol = "log2FC",
  xCol = "consBox_LysCont",
  colourCol = "protAbundance_log_binsLab",
  yLims = c(-2, 3),
  hlineInts = c(-1, 0, 1)
){
  # Prep data
  dat <- dat %>% .[rowSums(is.na(.[, c(logFCcol, xCol, colourCol)])) == 0,  ] %>%
    as.data.frame
  ## Combine columns to make one x-axis
  dat <- dat %>% .[order(.[, colourCol], .[, xCol]), ] %>%
    mutate(xFact = paste(!!sym(colourCol), !!sym(xCol), sep = "__")) %>%
    mutate(xFact = factor(xFact, levels = unique(xFact)))

  ## Calculate numbers to show above violins
  dataNums <- dat %>%
    group_by(xFact) %>%
    summarise(num = n(), .groups = "drop")
  ## Calculate y value for number annotations
  dataNums$y <- max(yLims) * 0.9

  # Plot
  plt <- ggplot(dat, aes(x = xFact, y = !!sym(logFCcol), color = !!sym(colourCol), fill = !!sym(colourCol))) +
    geom_boxplot(outlier.shape = NA, size = 0.235, alpha = 0.2, coef = 0, width = 0.7) +
    geom_hline(yintercept = hlineInts, size = 0.235, linetype = "dashed", alpha = 0.5) +
    geom_text(data = dataNums, aes(x = xFact, y = y, label = num),
                          size = 1.875 * 3 / 5, colour = "black",
                          position = position_dodge2(width = 0.9),
                          inherit.aes = F) +
    comfy_theme(rotate_x_text = T) +
    theme(axis.text.x = element_text(colour = "black",size = 5, hjust = 1, vjust = 1, angle = 45),
          axis.title.x = element_blank()) +
    lims(y = yLims)
  return(plt)
}

# Plot: one boxplot per treatment, x = functional_score bin, colour = abundance bin
pltLst <- map(prtsmInhibCols, function(treatCol){
  plt <- box_FCs(
                dat = ubiFeatPreds,
                logFCcol = treatCol,
                xCol = "functional_score_binsLab") +
      ggtitle(treatCol)
  return(plt)
})
ggarrange(plotlist = pltLst, nrow = 3)
ggsave_pdfpng(file = "output/images/functional_scores/abundance/box_scoreBinAbunBin_prtmsi", width = 0.6*10, height = 2*3)
```





# Protein half life (Mathieson et al)
```{r}
# Load table (defaults to the "protein half lives high qual" sheet, which is
# already pre-filtered by the authors on fit R_sq; dataQual ("good"/"weak") is
# a separate, per-replicate quality call that we filter on below)
TableS2 <- read_xlsx(path = paste0(home_directory, "data/supplementary_tables/Mathieson_NatComms_2018/Table_S2.xlsx"))
```

```{r}
# Reshape to one row per gene x cell line x replicate
halfLifeLong <- TableS2 %>%
  pivot_longer(cols = -gene_name,
               names_to = c("cellLine", "replicate", "measure"),
               names_pattern = "^(.*) replicate (\\d+) (half_life|dataQual|R_sq)$",
               values_to = "value",
               values_transform = list(value = as.character)) %>%
  mutate(cellLine = trimws(gsub(",$", "", cellLine))) %>%
  pivot_wider(names_from = measure, values_from = value) %>%
  mutate(half_life = as.numeric(half_life),
         R_sq = as.numeric(R_sq))

# Number of (per-replicate) values per cell line before filtering
nBefore <- halfLifeLong %>%
  group_by(cellLine) %>%
  summarise(n_before = sum(!is.na(half_life)), .groups = "drop")

# Keep only high-quality replicates, then average remaining replicates per gene x cell line
halfLifeFilt <- halfLifeLong %>%
  mutate(half_life = ifelse(dataQual == "good", half_life, NA)) %>%
  group_by(gene_name, cellLine) %>%
  summarise(half_life = mean(half_life, na.rm = TRUE), .groups = "drop") %>%
  mutate(half_life = ifelse(is.nan(half_life), NA, half_life))

# Number of values per cell line after filtering + collapsing replicates
nAfter <- halfLifeFilt %>%
  group_by(cellLine) %>%
  summarise(n_after = sum(!is.na(half_life)), .groups = "drop")

# Summary: values per cell line before vs. after filtering + collapsing
halfLifeSumm <- full_join(nBefore, nAfter, by = "cellLine")
```

```{r}
# log10-transform and reshape to one row per gene, one column per cell line
halfLifeFilt$log10_half_life <- log10(halfLifeFilt$half_life)
halfLifeWide <- halfLifeFilt %>%
  dplyr::select(gene_name, cellLine, log10_half_life) %>%
  pivot_wider(names_from = cellLine, values_from = log10_half_life)
cellLineCols <- setdiff(colnames(halfLifeWide), "gene_name")
halfLifeWide$Average <- rowMeans(halfLifeWide[, cellLineCols], na.rm = TRUE)
halfLifeWide$Average[is.nan(halfLifeWide$Average)] <- NA

# Cell line vs cell line correlation matrix (QC)
cor(halfLifeWide[, cellLineCols], use = "pairwise.complete.obs")
```

```{r}
# Map half lives into a copy of ubiFeatPreds, matching gene_name to gene
halfLifeCols <- paste0("halfLife_", c(cellLineCols, "Average"))
halfLifeWide_prefixed <- halfLifeWide %>% rename_with(~paste0("halfLife_", .), -gene_name)
ubiFeatPreds_wHalfLife <- left_join(ubiFeatPreds, halfLifeWide_prefixed, by = c("gene" = "gene_name"))
```

```{r}
# Correlation (with p-value and n) of each log half-life source to each model feature
corDat <- expand_grid(halfLifeSource = halfLifeCols, feat = feats) %>%
  pmap_dfr(function(halfLifeSource, feat){
    x <- ubiFeatPreds_wHalfLife[[halfLifeSource]]
    y <- as.numeric(ubiFeatPreds_wHalfLife[[feat]])
    n <- sum(!is.na(x) & !is.na(y))
    ct <- if (n >= 3) cor.test(x, y) else NULL
    tibble(halfLifeSource = halfLifeSource, feat = feat,
           cor = if (!is.null(ct)) unname(ct$estimate) else NA_real_,
           pval = if (!is.null(ct)) ct$p.value else NA_real_,
           n = n)
  })

subset(corDat, feat %in% c("functional_score", "numQuant_all", "numReg_notPrtsm", "FinalBoxProcNum")) %>% View
```

## Turnover (half-life) correlations by cell line
```{r}
featsSel <- c("functional_score", "numQuant_all", "numReg_notPrtsm", "FinalBoxProcNum")
corDatSel <- subset(corDat, feat %in% featsSel) %>%    
    mutate(feat = factor(feat, levels = featsSel))

corDatCellLines <- subset(corDatSel, halfLifeSource != "halfLife_Average") 
corDatMean <- corDatCellLines %>% group_by(feat) %>% summarise(cor = mean(cor, na.rm = TRUE), .groups = "drop")

ggplot(corDatMean, aes(x = feat, y = cor)) +
    geom_col(size = 0.235, colour = "black", fill = "grey", alpha = 0.7) +
    geom_point(data = corDatCellLines, shape = 16, size = 1) +
    comfy_theme(rotate_x_text = TRUE) +
    labs(title = "Half-life correlation (Mathieson et al. 2018)", y = "Pearson's R")
ggsave_pdfpng(file = "output/images/functional_scores/abundance/bar_halfLifeCor_byCellLine",
                width = 1.65, height = 2.3)
```

## Bin average protein half-life and compare to prtsmi
```{r}
# Bin average protein half-life
binN <- 10
ubiFeatPreds_wHalfLife$halfLife_Average_bins <- cut_number(ubiFeatPreds_wHalfLife$halfLife_Average, binN)
ubiFeatPreds_wHalfLife$halfLife_Average_binsLab <- ubiFeatPreds_wHalfLife$halfLife_Average_bins
levels(ubiFeatPreds_wHalfLife$halfLife_Average_binsLab) <- paste0("Turnover", 1:binN)

# Visualise score distribution in each bin
ggplot(ubiFeatPreds_wHalfLife, aes(x = functional_score)) +
    geom_histogram() +
    comfy_theme() +
    facet_wrap(~halfLife_Average_binsLab)

# Correlate within each bin
plotLst <- list()
k <- 1
datLst <- ubiFeatPreds_wHalfLife %>% subset(!is.na(halfLife_Average_binsLab)) %>%
    split(.$halfLife_Average_binsLab)
for (treatCol in prtsmInhibCols){
    for (i in 1:length(datLst)){
        plt <- ggplot(datLst[[i]], aes(x = functional_score, y = !!sym(treatCol))) +
            geom_hex(bins = 20) +
            geom_hline(yintercept = 0, linewidth = 0.235) +
            geom_smooth(method = "lm", se = TRUE, color = "red", linewidth = 0.235, alpha = 0.3) +
            hexScale +
            comfy_theme() + ggtitle(names(datLst)[i])
        plotLst[[k]] <- plt
        k <- k + 1
    }
}
ggarrange(plotlist = plotLst, nrow = 3, ncol = 10)
ggsave_pdfpng(file = "output/images/functional_scores/abundance/hex_halfLifeBin_score_prtsmi",
                width = 10*2.5, height = 3*2)
```