# Background
Here I analyse the accuracy of the model within individual proteins

# Initialise
NOTE: This entire section will easily take 5-10min to run
```{r}
library(tidyverse)
library(purrr)
library(ggplot2)
library(ggridges)
library(readxl)
library(openxlsx)
library(data.table)
library(caret)

home_directory <- ""
source(paste0(home_directory, "scripts/essential_visualisations_JvG.R"))
source(paste0(home_directory, "scripts/enrichment_analysis_JvG.R"))
source(paste0(home_directory, "scripts/caret_JvG.R"))
source("scripts/01_general.R")
source("scripts/00_objects.R")
source("scripts/03_visualisations.R")
source("scripts/04_ubnOmics.R")
source("scripts/02_model_training.R")
source("scripts/16_model_accuracy.R") # Auxiliary functions for assessing accuracy

# Load data
load("data/functional_scores/ubiFeatPreds__7.RData")
load("data/features/ubnOmics/ubnOmicsFeauresAll__4.RData")
load("data/features/ubnOmics/ubnOmicsComp_hsPRIDE_lst__8.RData")
load("data/hotspots/hotspotSites__3.RData")
load("data/functional_scores/datFeats__7.RData")

#Set up bin breaks for the functional score, and other aesthetic paramters
funcScoreBreaks <- c(0, 0.2, 0.4, 0.6, 0.8, 1)
funcScoreBinCols <- brewer.pal(9, "Blues")[c(1, 3, 5, 7, 9)]
funcScoreFillAlpha <- 0.3

#Get gene identifiers for our ubi sites
library(AnnotationDbi)
library(org.Hs.eg.db)
collapseFunc <- function(x){
    x <- unique(x)
    if (is.na(x[1])){
        return(NA)
    } else {
        return(paste(x, collapse = ";"))
    }
}
geneMap <- AnnotationDbi::select(org.Hs.eg.db, keys = unique(ubiFeatPreds$uniprot), keytype = "UNIPROT",
                                    columns = c("SYMBOL")) %>% 
                                group_by(UNIPROT) %>% 
                                summarise(gene = collapseFunc(SYMBOL))
ubiFeatPreds <- left_join(ubiFeatPreds, dplyr::rename(geneMap, "uniprot" = "UNIPROT"), by = "uniprot")
##Expand out by gene when they are not unique
ubiFeatPreds_byGene <- separate_rows(ubiFeatPreds, "gene", sep = ";") %>% 
                        mutate(gene_site = paste(gene, site, sep = "_")) 

# Load clinvar
load("output/data/genetics/clinvarDatSub__2.RData")
clinvarDatSub <- subset(clinvarDatSub, grepl("missense_variant", MC))
clinvarDatSub <- clinvarDatSub %>% mutate(clinvar_label = factor(clinvar_label, levels = c("Benign", "VUS", "Pathogenic")))
levels(clinvarDatSub$clinvar_label)[2] <- "Uncertain" 

# Load base editing
load("output/data/genetics/LysBEdat_funcScore_sites.RData")
LysBEdat_funcScore_sites$pos <- gsub("^.+_" ,"", LysBEdat_funcScore_sites$aa_id)

# Load NLS
load("output/data/functional_scores/ubiFeatPreds_NLS__7.RData")

## Load protein abundance data
load("data/features/abundDat__1.RData")
abundDat$uniprot_pos <- gsub("_K", "_", rownames(abundDat))
abundDat$uniprot <- gsub("_.+$", "", abundDat$uniprot_pos)
abundFeatsLog <- colnames(abundDat) %>% .[grep("log$", .)]
protAbundDat <- abundDat[, c("uniprot", "protAbundance_log")] %>% distinct

# Load PSP annotations
load("data/PSP/PSP_reg_ubCorr__3.RData")
```

## Quantile-normalise features within proteins
```{r}
cols <- featureCols
feats_quantNorm <- ubiFeatPreds[, c("uniprot", "uniprot_site", cols)] %>% 
    pivot_longer(cols = all_of(cols), names_to = "column", values_to = "predCol") %>%  
    group_by(uniprot, column) %>% 
    mutate(
        predColQuant = to_quantiles(predCol), 
        predColRank = rank(-predCol),
        numSites = n()) %>% ungroup  %>% 
    pivot_wider(names_from = "column", values_from = c("predColQuant", "predColRank", "predCol")) 
# Reorder columns
colsReord <- map(cols, function(x){colnames(feats_quantNorm)[grep(x, colnames(feats_quantNorm))]}) %>% unlist %>% unique
colsReord <- c(colsReord, setdiff(colnames(feats_quantNorm), colsReord))
feats_quantNorm <- feats_quantNorm[, colsReord]
```



## Prepare external data to add
This is done at the level of uniprot_site, so it can be done once and then mapped to multiple datasets more quickly
```{r}
# Set up data
siteData <- distinct(ubiFeatPreds[, c("uniprot", "uniprot_site", "uniprot_pos")])

# Add prtsm inhib data
prtsmInhibCols <- c("Bortezomib_meanlog2FC", "MG-132_meanlog2FC", "Epoxomicin_meanlog2FC")
if (length(intersect(prtsmInhibCols, colnames(siteData))) == 0){
    siteData <- left_join(siteData, ubnOmicsFeaturesAll[, c("uniprot_pos", prtsmInhibCols)],
                                by = "uniprot_pos")
}

# Add clinvar data
useStars <- FALSE
CVdatFilt <- clinvarDatSub # Optional filtering here
siteData$clinvarClass <- NA
levels <- c("Uncertain", "Benign", "Pathogenic") # This degines the order of evidence. Pathogenic trumps all
if (useStars){
    for (lev in levels){
        siteData$clinvarClass[which(siteData$uniprot_site %in% subset(CVdatFilt, clinvar_label == lev & CLNREVSTAT_STARS == 1)$uniprot_site)] <- paste0(lev, "_1star")
        siteData$clinvarClass[which(siteData$uniprot_site %in% subset(CVdatFilt, clinvar_label == lev & CLNREVSTAT_STARS > 1)$uniprot_site)] <- paste0(lev, "_2+star")
    }
    siteData$clinvarClass <- factor(siteData$clinvarClass, levels = sort(unique(siteData$clinvarClass[which(!is.na(siteData$clinvarClass))])))
} else {
    for (lev in levels){
        siteData$clinvarClass[which(siteData$uniprot_site %in% subset(CVdatFilt, clinvar_label == lev)$uniprot_site)] <- lev
    }
    siteData$clinvarClass <- factor(siteData$clinvarClass, levels = levels)
}

# Add base editing data
siteData <- left_join(
    siteData, 
    LysBEdat_funcScore_sites[, c("aa_id", "typeLabel")] %>% dplyr::rename("uniprot_pos" = "aa_id", "BE_typeLabel" = "typeLabel")
    )

# Add abundance
siteData <- left_join(
    siteData, 
    abundDat
    )

# Add NLS data
siteData <- left_join(siteData, ubiFeatPreds_NLS)

# Add PSP annotations
siteData <- left_join(siteData, PSP_reg_ubCorr[, c("uniprot_site", "ON_PROCESS", "ON_FUNCTION")])

# Add gene names
siteData <- left_join(siteData, distinct(ubiFeatPreds[, c("uniprot_site", "gene")]))

# Add # regulatory sites
numRegSites <- ubiFeatPreds %>% group_by(uniprot) %>% 
    summarise(numReg_notProtDeg = sum(PSP_reg_notProtDeg), numReg_protDeg = sum(PSP_reg_ProtDeg)) %>% ungroup
siteData <- left_join(siteData, numRegSites, by = "uniprot") %>% 
    dplyr::relocate(colnames(numRegSites))
```

## Load test-set predictions
### Load predictions on all sites and subset for those not in training
This ensures that sites excluded from the entire training process will still be there. 
For these, I include them only in test sets where the entire protein was included
```{r}
# Load
suffix <- "attempt11_revisions"
load(paste0("output/data/functional_scores/allPredsLstDf__", suffix, ".RData")) # Load predictions
load(paste0("output/data/functional_scores/trainTestforPredSplits__", suffix, ".RData")) # Load train-test splits
# Subset for data not in training
testSetPreds_wAll <- allPredsLstDf
testSetPreds_wAll$uniprot <- gsub("_[A-Z].+$", "", testSetPreds_wAll$uniprot_site)
testSetPreds_wAll$retain <- F
for (resamp in names(trainTestforPredSplits)){
    testDat <- trainTestforPredSplits[[resamp]]$testDat
    testProts <- unique(gsub("_.+$", "", rownames(testDat)))
    testSetPreds_wAll$retain[which(testSetPreds_wAll$fold == resamp & testSetPreds_wAll$uniprot %in% testProts)] <- T
}
testSetPreds_wAll <- subset(testSetPreds_wAll, retain == T) %>% dplyr::select(-c("retain"))
```


### Load only test-set predictions  
This excludes sites that were not in training at all e.g. degradation sites
```{r}
# Load
load(paste0("output/data/functional_scores/modsPreds_trainTestforPred_df__", suffix, ".RData"))
testSetPreds <- modsPreds_trainTestforPred_df 
```

### Process and add features
```{r}
datLst <- list("testSetPreds_wAll" = testSetPreds_wAll, "testSetPreds" = testSetPreds)

# Process
datLst <- map(datLst, function(data){
    # Add columns and data
    # Add identifier columns
    data$Resample <- data$fold
    data$uniprot <- gsub("_[A-Z].+$", "", data$uniprot_site)
    data$uniprot_pos <- gsub("_[A-Z]", "_", data$uniprot_site)
    # Assign prediction column
    predCol <- "X1"
    data$predCol <- data[[predCol]]
    # Quantile-normalise within proteins
    data <- data %>% group_by(uniprot, Resample) %>% 
        mutate(
            predColQuant = to_quantiles(predCol), 
            predColRank = rank(-predCol),
            numSites = n()) %>% ungroup 
    data$predColPercentile <- 100*data$predColQuant

    # Add features
    cols_to_add <- setdiff(colnames(siteData), colnames(data))
    if (length(cols_to_add) > 0){
        data <- left_join(data, siteData[, c(cols_to_add, "uniprot_site")], by = "uniprot_site")
    }
    return(data)
})
testSetPreds_wAll <- datLst$testSetPreds_wAll
testSetPreds <- datLst$testSetPreds

# Map training class annotations in
testSetPreds_wAll <- left_join(testSetPreds_wAll, distinct(testSetPreds[, c("uniprot_site", "obs")]), by = "uniprot_site")

# Subset for proteins with regulatory sites
regProts <- subset(testSetPreds, obs == "X1")$uniprot %>% unique
testSetPreds_regProts <- subset(testSetPreds, uniprot %in% regProts)
```

### Take median of values across Resamples
```{r}
testSetPreds_wAll_summ <- testSetPreds_wAll %>% 
    group_by(uniprot_site) %>%
    summarise(
        across(where(is.numeric), ~ median(.x, na.rm = TRUE)),
        across(where(~ !is.numeric(.x)), dplyr::first),
        .groups = "drop"
    )

# Add median test-set score to ubiFeatPreds
ubiFeatPreds <- left_join(
    ubiFeatPreds,
    testSetPreds_wAll_summ[, c("uniprot_site", "predCol")] %>% dplyr::rename("functional_score_testSet" = "predCol"),
    by = "uniprot_site"
)
```



# Inspect quantile-normalised scores for proteins with regulatory sites
## Summarise data at different levels
```{r}
outputDir <- "output/images/functional_scores/withinProtein/"

##### For training data sites #####
regSiteDat <- subset(testSetPreds, obs == "X1") 
# Median across resamples
regSiteDat_ResampleSumm <- regSiteDat %>% group_by(uniprot_site) %>% 
    summarise(predColPercentile = median(predColPercentile), predColRank = median(predColRank), uniprot = uniprot[1]) %>% ungroup 
# Max within proteins
regSiteDat_protMax <- regSiteDat %>% group_by(uniprot, Resample) %>% 
    summarise(predColPercentile = max(predColPercentile), predColRank = min(predColRank)) %>% ungroup
# Combining the two
regSiteDat_summToMax <- regSiteDat_ResampleSumm %>% group_by(uniprot) %>% 
    summarise(predColPercentile = max(predColPercentile), predColRank = min(predColRank)) %>% ungroup
regSiteDat_maxToSumm <- regSiteDat_protMax %>% group_by(uniprot) %>% 
    summarise(predColPercentile = median(predColPercentile), predColRank = median(predColRank)) %>% ungroup
```


## Plot quantile-normalised scores 
```{r}
# Function
quantile_violin <- function(data){
    data$dummy <- "a"
    plt <- ViolBoxplot(data, vars = list(x = "dummy", colour = 'dummy', fill = "dummy", y = "predColPercentile"),
                includeLegend = F) + labs(y = "Score percentile within protein") +
        theme(axis.text.x = element_blank(), axis.title.x = element_blank()) +
        scale_y_continuous(breaks = seq(0, 100, by = 20)) + 
        scale_colour_manual(values = ubiSiteColour) + 
        scale_fill_manual(values = alpha(ubiSiteColour, 0.1))
    return(plt)
}
##### For training data sites #####
# All data
quantile_violin(regSiteDat_ResampleSumm) + ggtitle("Regulatory sites")
ggsave_pdfpng(file = paste0(outputDir, "viol_scoreQuant_", "ResampSumm"), width = 1.5, height = 1.5)
# Protein max, averaged across resamples
quantile_violin(regSiteDat_maxToSumm) + ggtitle("Max per protein")
ggsave_pdfpng(file = paste0(outputDir, "viol_scoreQuant_", "maxToSumm"), width = 1.5, height = 1.5)
```

Compare to protein abundance
```{r}
# Add protein abundance
regSiteDat_ResampleSumm <- left_join(regSiteDat_ResampleSumm, distinct(abundDat[, c("uniprot", "protAbundance_log")]), by = "uniprot")
ggplot(regSiteDat_ResampleSumm, aes(x = protAbundance_log, y = predColPercentile)) + 
    geom_point(shape = 16, size = 1, alpha = 0.7) + 
    comfy_theme()
ggsave_pdfpng(file = paste0(outputDir, "pplot_scoreQuant_vs_protAbund_ResampSumm"), width = 2, height = 2)
```


## Inspect ranks for regulatory sites
```{r}
# Function to plot number of sites at each rank level
plot_rank_barplot <- function(data, rankLimit = 5, rankCol = predColRank, ylab = "% of sites"){
    rankCol <- enquo(rankCol)

    rankSumm <- data %>%
        mutate(
            rankBin = case_when(
                is.na(!!rankCol) ~ NA_character_,
                !!rankCol >= rankLimit ~ paste0(">=", rankLimit),
                TRUE ~ as.character(as.integer(!!rankCol))
            ),
            rankBin = factor(
                rankBin,
                levels = c(as.character(seq_len(rankLimit - 1)), paste0(">=", rankLimit))
            )
        ) %>%
        filter(!is.na(rankBin)) %>%
        dplyr::count(rankBin, name = "count") %>%
        mutate(percent = 100 * count / sum(count))
    levels(rankSumm$rankBin) <- sapply(levels(rankSumm$rankBin), function(x){
        if (sum(grepl(">=", x, fixed = T)) > 0){
            x <- gsub(">=", "", x, fixed = T)
            x <- paste0(x, "+")
            return(x)
        } else {return(x)}
    })

    ggplot(rankSumm, aes(x = rankBin, y = percent)) +
        geom_col(width = 0.8, size = 0.235/2, fill = alpha(ubiSiteColour, 0.75), colour = ubiSiteColour) +
        labs(x = "Score rank within protein", y = ylab) +
        scale_y_continuous(breaks = seq(0, 100, by = 20)) +
        comfy_theme()
}
##### For training data sites #####
# All data
plot_rank_barplot(regSiteDat_ResampleSumm, rankLimit = 6) + ggtitle("Regulatory sites")
ggsave_pdfpng(file = paste0(outputDir, "bar_scoreRank_", "ResampSumm"), width = 1.5, height = 1.7)
# Protein max, averaged across resamples
plot_rank_barplot(regSiteDat_maxToSumm, rankLimit = 6, ylab = "% of proteins") + ggtitle("Max per protein")
ggsave_pdfpng(file = paste0(outputDir, "bar_scoreRank_", "maxToSumm"), width = 1.5, height = 1.7)
```


# Inspect sites that are top-ranked within their protein
```{r}
rank1Sites <- subset(regSiteDat_ResampleSumm, predColRank == 1)$uniprot_site
rank1Sites_info <- subset(testSetPreds, uniprot_site %in% rank1Sites) %>% 
    group_by(uniprot_site) %>%
    summarise(
        across(where(is.numeric), ~ median(.x, na.rm = TRUE)),
        across(where(~ !is.numeric(.x)), dplyr::first),
        .groups = "drop"
    )
```

## Export all sites for inspection
```{r}
regSites_info <- subset(testSetPreds_wAll_summ, obs == "X1") 
# Add quantile-normalised features
regSites_info <- left_join(
    regSites_info, 
    feats_quantNorm %>% dplyr::select(-c("uniprot", "numSites")), 
    by = "uniprot_site")
# Order
regSites_info <- regSites_info %>% .[order(.$predColRank, -.$predCol, decreasing = F), ]
# Export
write.csv(regSites_info, file = "output/data/functional_scores/withinProteins/regSites_info.csv", na = "")
```



# Plot lollypop plots for rank-1 proteins
```{r}
# Select favourite sites
unipSiteVec <- c(
    "P46783_K138", #
    "P22314_K604", #
    "P60866_K8", #
    "Q13164_K184", #
    "Q15004_K24", #
    "P51149_K38", #
    "P54252_K117", #
    "O14757_K132", #
    "Q9BXW9_K561", #
    "O43318_K158", #
    "P49915_K182", #
    "Q9NVI1_K523", #
    "Q96BN8_K66", #
    "O94992_K289" #
)

unipVec <- unique(gsub("_.+$", "", unipSiteVec))

# Set up domains
domainListList <- map(unipVec, function(unip){
    ind <- which(ubiFeatPreds$uniprot == unip)
    domainsList <- list("a" = c(1, ubiFeatPreds$proteinLength[ind[1]]))
    names(domainsList)[1] <- ubiFeatPreds$gene[ind[1]]
    return(domainsList)
})
names(domainListList) <- unipVec
## Make domains
domainListList <- list( # From PhosphositePlus unless otherwise specified
    "Q9NVI1" = list(  # pfam
            "Cap" = c(1, 53),
            "S1" = c(64, 280),
            "HD1" = c(284, 370),
            "S2" = c(378, 541),
            "HD2" = c(555, 785),
            "S3" = c(804, 1032),
            "S4" = c(1045, 1294)
            ),

    "P46783" = list(  
            "S10_plectin" = c(3, 98)
            ),


    "P22314" = list(  # From pfam
            "FCCH" = c(227, 296),
            "4HB" = c(297, 366),
            "ThiF" = c(451, 637),
            "SCCH" = c(638, 884),
            "Thif" = c(885, 946),
            "UFD" = c(955, 1053)
            ),
    
    "P60866" = list(  
            "Ribosomal_S10" = c(21, 115)
            ),


    "Q13164" = list(  
            "Pkinase" = c(55, 347)
            ),

    "Q15004" = list(  
            "PAF" = c(1, 111)
            ),

    "P51149" = list(  
            "Ras" = c(10, 175)
            ),

    "P54252" = list(  
            "Josephin" = c(9, 166),
            "UIM" = c(224, 240),
            "UIM" = c(244, 260),
            "SUIM_assoc" = c(263, 329),
            "UIM" = c(335, 351)
            ),

    "O14757" = list(  
            "Pkinase" = c(9, 265)
            ),

    "Q9BXW9" = list(  
            "FancD2" = c(1, 1414)
            ),

    "O43318" = list(  
            "PK_Tyr_Ser-Thr" = c(36, 284)
            ),

    "P49915" = list(  
            "GATase" = c(20, 210),
            "NAD_synthase" = c(219, 349),
            "GMP_synt_C" = c(597, 692)
            ),

    "Q96BN8" = list(  
            "Peptidase_C101" = c(80, 344)
            ),

    "O94992" = list(  
            "HEXIM" = c(164, 301)
            )
)


# Make plots
yArgs <- list(
    list("functional_score", "Importance score", "finalScore"),
    list("functional_score_testSet", "Importance score (test set median)", "testSetScore")
)
for (unipSel in unipVec){
    ubiFeat_sub <- subset(ubiFeatPreds, uniprot == unipSel)
    gene <- ubiFeat_sub$gene[1]
    domainsList <- domainListList[[unipSel]]
    ## Plot, either final score or test-set score
    for (arg in yArgs){
        plot_lollipop(data = ubiFeat_sub, 
            domains = domainsList, 
            aes_colour = "PSP_reg_notProtDeg", aes_size = "PSP_reg_notProtDeg", # For now, show non-degradation sites
            colour_values = setNames(c("#203569", "#c7dfec"), c(T, F)), 
            size_values = setNames(c(2.5, 1.5), c(T, F)),
            ycol = arg[[1]], sitesToLabel = subset(ubiFeat_sub, PSP_reg_notProtDeg == T)$uniprot_site) + 
            labs(y = arg[[2]]) + theme(axis.title.x = element_blank())
        ggsave_pdfpng(file = paste0("output/images/functional_scores/withinProtein/examples/lolly_", gene, "_", unipSel, "_", arg[[3]]),
                        width = 2.4, height = 1.25)
    }
}
```