# Background
Here I process and prepare experimental data of ubiquitin site mutations in yeast.

# Initialise
```{r}
library(tidyverse)
library(purrr)
library(ggplot2)
library(readxl)
library(openxlsx)

home_directory <- ""
source(paste0(home_directory, "scripts/essential_visualisations_JvG.R"))
source("scripts/01_general.R")
source("scripts/00_objects.R")
source("scripts/03_visualisations.R")
source("scripts/04_ubnOmics.R")

load("data/functional_scores/ubiFeatPreds__7.RData")
load("data/features/ubnOmics/ubnOmicsComp_hsPRIDE_lst__8.RData")
```

Load in proteomes
```{r}
##Function: read in a fasta file of protein sequences and convert them into a data frame
library(seqinr)
read_fastaseq <- function(directory){
  seqs_raw <- read.fasta(directory)
  seqs <- data.frame("sequence" = map(seqs_raw, ~paste(., collapse = "")) %>% unlist)
  seqs$sequence <- toupper(seqs$sequence)
  seqs$uniprotkb <- strsplit(rownames(seqs), "\\|") %>% map(~.[2]) %>% unlist
  return(seqs)
}
#Load human proteomes
uniprotHsIso <- read_fastaseq(paste(home_directory, 
                                        "data/biol_databases/uniprot/uniprotkb_proteome_UP000005640_2025_02_19.fasta", sep = ""))
uniprotHsIsoRevi <- read_fastaseq(paste(home_directory, 
                                        "data/biol_databases/uniprot/uniprotkb_proteome_UP000005640_AND_revi_2025_02_19.fasta", sep = ""))
save(uniprotHsIso, file = "output/data/uniprotHsIso.RData")
save(uniprotHsIsoRevi, file = "output/data/uniprotHsIsoRevi.RData")
# Checkpoint
#load("output/data/uniprotHsIso.RData")
#load("output/data/uniprotHsIsoRevi.RData")
##Load yeast proteome
sgdProteome <- read_fastaseq(paste(home_directory, 
                                         "data/biol_databases/SGD/orf_trans_all.fasta", sep = "")) %>%
  mutate(ORF = rownames(.),
         sequence = gsub("*", "", sequence, fixed = T))


#Function for fetching sequence windows from uniprot using position and identifer
fetchSeqWindsUnip <- function(data, uniprotCol, positionCol){
  data <- as.data.frame(data)
  unipSeqWindVec <- c()
  for (i in 1:nrow(data)){
    unip <- data[i, uniprotCol]
    pos <- data[i, positionCol]
    unipSeq <- uniprotHsIso$sequence[which(uniprotHsIso$uniprotkb == unip)]
    ##Return NA if position is too long
    if (pos > str_length(unipSeq)){
      unipSeqWind <- NA
    } else {
      unipSeq <- paste(c(rep("_", 7), unipSeq, rep("_", 7)), collapse = "")
      unipSeqWind <- substr(unipSeq, start = pos, stop = pos + 14)
    }
    unipSeqWindVec[i] <- unipSeqWind
  }
  return(unipSeqWindVec)
}
```


# Load in and prep data
## Load in
```{r}
#Experimental data
mutSscores <- read.csv("data/experiments/Scores_KO_pointMut.csv")
sort(unique(mutSscores$anot))
##Extract mutation and type
mutSscores$mutation <- strsplit(mutSscores$anot, "_") %>% map(~.[2]) %>% unlist
mutSscores$mutType <- "Site"
mutSscores$mutType[which(mutSscores$mutation %in% c("DAMP", "KO"))] <- "Gene"
mutSscores$mutType[which(mutSscores$mutation %in% c("WT"))] <- "WT"
##Make nicer names
mutSscores$anotNice <- gsub("_", " ", mutSscores$anot) %>% gsub("DAMP", "KD", .)
mutSscores$anotNiceNoGene <- gsub("^.+ ", "", mutSscores$anotNice)

# Map between yeast and human sites
mutSiteMapProc <- read_csv("data/experiments/mutSiteMapProc__2.csv")
mutSscores <- subset(mutSscores, gen_name %in% c(mutSiteMapProc$geneMapped, "BY4741", "DHY214"))
```

# Prepare s-score data and export
## Prepare data
```{r}
#Order levels
mutSscores <- mutSscores %>%
  mutate(time = factor(time, levels = sort(unique(time), decreasing = T)),
         treatment = factor(treatment, levels = sort(unique(treatment), decreasing = T))) %>%
  .[order(.$treatment, .$time), ] %>%
  mutate(condition = factor(condition, levels = unique(condition)))

#Add replicate identifier
mutSscores <- mutSscores %>%
  .[order(.$strain), ] %>%
  group_by(anot) %>%
  mutate(replicateID = paste0(anot, "_", dense_rank(strain)))
```


## Export metadata for supps 
### Conditions 
```{r}
# Get conditions
condMeta <- data.frame("condition" = unique(datOneMut$condition)) %>%  
  mutate(time = gsub("-.+$", "", condition),
         treatment = gsub("^.+-", "", condition))
```


Using Bede's methods 
```{r}
# Load Bede's conditions
BedeMeta <- map(c("Sheet1", "Special"), function(sheet){
  read_xlsx(path = "data/experiments/yeastMethodsBede/conditions for screen.xlsx", sheet = sheet)
})
commonNames <- map(BedeMeta, colnames) %>% purrr::reduce(intersect)
BedeMeta <- map_dfr(BedeMeta, function(x){x[, commonNames]}, .id = "source") %>% 
  subset(!is.na(Abbreviation)) %>%  
  .[, c("abbrev_corrected", "condition", "carrier", "concentration")] %>% 
  distinct %>%  
  dplyr::rename("conditionDescription" = "condition", 
                "treatment" = "abbrev_corrected") %>%  
  mutate(treatment = toupper(gsub("48H-", "", treatment)))

# Map into metadata
condMeta <- left_join(condMeta, BedeMeta[, c("treatment", setdiff(colnames(BedeMeta), colnames(condMeta)))], by = "treatment") 
View(condMeta)

# Export into excel for supp info
## Prep
renameCols <- c(
  "condition" = 'Condition',
  "time" = "Timepoint (hours)",
  "treatment" = "Treatment",
  "conditionDescription" = "Condition description",
  "carrier" = "Carrier",
  "concentration" = "Concentration"
)
condMetaExp <- condMeta[, names(renameCols)] %>% 
  dplyr::rename(!!!setNames(names(renameCols), renameCols))
## Write
outDir = "output/data/experiments/SuppTable_yeastExps.xlsx"
sheet = "conditionDescriptions"
if (file.exists(outDir)){
  wb <- loadWorkbook(outDir)
} else {
  wb <- createWorkbook(outDir)
}
removeWorksheet(wb, sheet)
addWorksheet(wb, sheet)
writeData(wb, sheet = sheet, x = condMetaExp)
saveWorkbook(wb, outDir, overwrite = TRUE)
```


### Strain info  
```{r}
# Set up metadata
strainMeta <- mutSscores[, c("strain", "anot", "ID")] %>% distinct %>%  
  mutate(strainID = gsub("_.+$", "", strain),
         replicateID = strain %>% gsub("^.+_", "", .) %>% gsub("-.+$", "", .))

# Inspect strainIDs
strainIDMeta <- strainMeta[, c("anot", "strainID")] %>% distinct 

# Make metadata for individual mutants
mutMeta <- strainMeta %>% dplyr::select("anot") %>% distinct %>% 
  dplyr::rename("yeastMutant" = "anot")

# Map in human sites
matchDat <- mutSiteMapProc %>% subset(windows_match == "Y") %>% 
  dplyr::select("humanSite", "yeastMutant", "hasFuncScore")
mutMeta <- left_join(mutMeta, matchDat, by = "yeastMutant")

# Map in func score
mutMeta <- left_join(mutMeta, ubiFeatPreds[, c("uniprot_site", "functional_score")] %>%  
  dplyr::rename("humanSite" = "uniprot_site"), 
                    by = "humanSite")

# Export
## Prep
renameCols <- c(
  "yeastMutant" = 'Gene/strain + mutation',
  "humanSite" = "Human ubi-site",
  "functional_score" = "Functional score (human ubi-site)"
)
mutMetaExp <- mutMeta[, names(renameCols)] %>% 
  dplyr::rename(!!!setNames(names(renameCols), renameCols))
## Write
outDir = "output/data/experiments/SuppTable_yeastExps.xlsx"
sheet = "mutantDescriptions"
if (file.exists(outDir)){
  wb <- loadWorkbook(outDir)
} else {
  wb <- createWorkbook(outDir)
}
removeWorksheet(wb, sheet)
addWorksheet(wb, sheet)
writeData(wb, sheet = sheet, x = mutMetaExp)
saveWorkbook(wb, outDir, overwrite = TRUE)
```

### Primers and donor DNA
```{r}
primerDonorF <- "data/experiments/yeastMethodsBede/Final Guides to order_with_mod.xlsx"

# Primers
primerDat <- read_xlsx(path = primerDonorF, sheet = "Guides") %>%  
  as.data.frame %>%  
  subset(!is.na(Name)) %>% 
  dplyr::rename("sequence" = "5>3") %>%  
  mutate(Name = gsub("\\s+_", "_", Name)) %>%
  mutate(gene = strsplit(Name, "_") %>% map(~.[1]) %>% unlist,
          pos = strsplit(Name, "_") %>% map(~.[2]) %>% unlist) %>%  
  separate_rows("pos", sep = ",") %>%  
  mutate(gene_pos = paste(gene, pos, sep = "_"))
## Check if all mutations are present
muts_gene_pos <- mutMeta$yeastMutant %>% .[grep("_K\\d+R$", .)] %>% gsub("_K", "_", .) %>% gsub("R$", "", .)
setdiff(muts_gene_pos, primerDat$gene_pos)
## Subset for final mutations
primerDatSub <- subset(primerDat, gene_pos %in% muts_gene_pos) %>%  
  .[, c("Name", "sequence")] %>% distinct %>%  
  .[order(.$Name), ]
## Export
## Prep
renameCols <- c(
  "Name" = 'Target',
  "sequence" = "Sequence"
)
primerDatSubExp <- primerDatSub[, names(renameCols)] %>% 
  dplyr::rename(!!!setNames(names(renameCols), renameCols))
## Write
outDir = "output/data/experiments/SuppTable_yeastExps.xlsx"
sheet = "primers"
if (file.exists(outDir)){
  wb <- loadWorkbook(outDir)
} else {
  wb <- createWorkbook(outDir)
}
removeWorksheet(wb, sheet)
addWorksheet(wb, sheet)
writeData(wb, sheet = sheet, x = primerDatSubExp)
saveWorkbook(wb, outDir, overwrite = TRUE)

# Donor DNA
donorDat <- read_xlsx(path = primerDonorF, sheet = "Donors") %>%  
  as.data.frame %>%  
  subset(!is.na(Name)) %>%  
  subset(grepl("/", Name) == F) %>%  
  mutate(Name = gsub("\\s+_", "_", Name)) %>%
  mutate(gene = strsplit(Name, "_") %>% map(~.[1]) %>% unlist,
          pos = strsplit(Name, "_") %>% map(~.[2]) %>% unlist) %>%   
  mutate(gene_pos = paste(gene, pos, sep = "_"))
setdiff(muts_gene_pos, donorDat$gene_pos)
## Subset for final mutations
View(donorDat)
donorDatSub <- subset(donorDat, gene_pos %in% muts_gene_pos) %>%  
  .[, c("Name", "Sequence")] %>% distinct %>%  
  .[order(.$Name), ]
## Export
## Prep
renameCols <- c(
  "Name" = 'Target',
  "Sequence" = "Sequence"
)
donorDatSubExp <- donorDatSub[, names(renameCols)] %>% 
  dplyr::rename(!!!setNames(names(renameCols), renameCols))
## Write
outDir = "output/data/experiments/SuppTable_yeastExps.xlsx"
sheet = "donors"
if (file.exists(outDir)){
  wb <- loadWorkbook(outDir)
} else {
  wb <- createWorkbook(outDir)
}
removeWorksheet(wb, sheet)
addWorksheet(wb, sheet)
writeData(wb, sheet = sheet, x = donorDatSubExp)
saveWorkbook(wb, outDir, overwrite = TRUE)
```

### S-scores 
```{r}
# Map in  metadata
mutSscores_wMeta <- mutSscores %>% 
  ungroup %>% 
  dplyr::select(c("score", "qvalue", "condition", "strain")) %>%  
  as.data.frame
metaCols <- c("anot", "strainID", "replicateID")
mutSscores_wMeta$strain <- factor(mutSscores_wMeta$strain, 
  levels = strainMeta$strain)
strainMeta <- as.data.frame(strainMeta)
for (col in metaCols){
  mutSscores_wMeta[, col] <- mutSscores_wMeta$strain 
  levels(mutSscores_wMeta[, col]) <- strainMeta[, col]
  mutSscores_wMeta[, col] <- as.character(mutSscores_wMeta[, col])
}

# Export
renameCols <- c(
  "anot" = 'Gene/strain + mutation',
  "strainID" = "Strain ID",
  "replicateID" = "Replicate",
  "score" = "S-score",
  "qvalue" = "Q-value"
)
mutSscores_wMetaExp <- mutSscores_wMeta[, names(renameCols)] %>% 
  dplyr::rename(!!!setNames(names(renameCols), renameCols))
## Write
outDir = "output/data/experiments/SuppTable_yeastExps.xlsx"
sheet = "sscores"
if (file.exists(outDir)){
  wb <- loadWorkbook(outDir)
} else {
  wb <- createWorkbook(outDir)
}
removeWorksheet(wb, sheet)
addWorksheet(wb, sheet)
writeData(wb, sheet = sheet, x = mutSscores_wMetaExp)
saveWorkbook(wb, outDir, overwrite = TRUE)
```



# Collapse replica and timepoints, and plot results
```{r}
colsToKeep <- colnames(mutSscores)[-which(colnames(mutSscores) %in% c("treatment", "anot", "score", "qvalue"))]
mutSscores_coll <- mutSscores %>% 
  group_by(treatment, anot) %>%  
  summarise(score = mean(score),
            qvalue = geoMean(qvalue),
            across(all_of(colsToKeep), ~unique(.)[1]),
            .groups = "drop")
```


## Filter for significance 
```{r}
mutSscores_coll$sig <- mutSscores_coll$qvalue < 0.01

#Significance per mutant
mutSscores_collSumm <- mutSscores_coll %>% group_by(anot) %>% 
  summarise(numPhen = sum(sig), gen_name = unique(gen_name), mutType = unique(mutType)) %>% 
  .[order(.$numPhen, decreasing = T), ]

split(mutSscores_collSumm, mutSscores_collSumm$mutType)
table(mutSscores_collSumm$mutType)
```

## Plot heatmaps for site mutants with phenotypes
```{r}
SitesWithPhenos <- subset(mutSscores_collSumm, mutType == "Site" & numPhen > 0)$anot 

plotParams <- list(
  "YPT1_K122R" = list(width = 2, height = 2),
  "PGI1_K371R" = list(width = 1.75, height = 2),
  "TUB2_K58R" = list(width = 1.75, height = 2)
)

mutSscores_coll <- as.data.frame(mutSscores_coll)
for (i in 1:length(plotParams)){
  site <- names(plotParams)[i]
  params <- plotParams[[i]]
  gene <- subset(mutSscores_coll, anot == site)$gen_name %>% unique
  sscoreDatSub <- subset(mutSscores_coll, gen_name == gene)
  ## Get conditions significant in any site mutant for this gene
  sigConds <- subset(sscoreDatSub, sig == T)$treatment %>% unique #Use all significant for this gene
  sscoreDatSub <- subset(sscoreDatSub, treatment %in% sigConds)
  ## Sort data
  sortedTreats <- sscoreDatSub %>% 
    filter(anot == site) %>% 
    arrange(score) %>% 
    distinct(treatment) %>% 
    pull(treatment)
  sscoreDatSub$treatment <- factor(sscoreDatSub$treatment, levels = sortedTreats)
  ## Plot
  hmap_sscores(sscoreDat = sscoreDatSub, xaxis = "anotNiceNoGene", yaxis = "treatment") + 
              theme(axis.title.x = element_blank(), axis.title.y = element_blank()) +
              ggtitle(gene)
  ggsave_pdfpng(file = paste0("output/images/experiments/hmaps/hmap_sscore_", site), 
                width = params$width, height = params$height)
}
```


# Overview of scores and number of phenotypes 
```{r}
# Map in functional score
humanYeastSumm <- left_join(mutSiteMapProc, ubiFeatPreds[, c("uniprot_site", "functional_score")] %>% 
                                            dplyr::rename("humanSite" = "uniprot_site"), by = "humanSite")
# Map in number of phenotypes
humanYeastSumm <- left_join(humanYeastSumm, mutSscores_collSumm[, c("anot", "numPhen")] %>% 
                                            dplyr::rename("yeastMutant" = "anot"))
humanYeastSumm <- humanYeastSumm %>% mutate(humanAApos = gsub("^.+_", "", humanSite)) %>% 
                                     mutate(humanGeneSite = paste(Gene, humanAApos, sep = " "),   
                                            yeastGeneSite = gsub("_", " ", yeastMutant)) %>% 
                                     mutate(yeastGeneSite = gsub("R$", "", yeastGeneSite))
```

```{r}
# Plot
# Create the barplot. Bars for functional_score (red) and numPhen (blue) will be side‐by‐side.
plotDat <- humanYeastSumm %>% 
  .[order(.$functional_score, decreasing= T), ] %>%
  mutate(row = row_number()) %>% 
  pivot_longer(cols = c("functional_score", "numPhen"), names_to = "col", values_to = "val")   %>% 
  mutate(humanYeastSite = paste(yeastGeneSite, " (", humanGeneSite, ")", sep = "")) %>%   
  subset(col == "functional_score")
## Plot  
ggplot(plotDat, aes(x = row, y = val, fill = col)) +
  geom_col(width = 0.8, position = position_dodge2(), 
            fill = "#b3d1e0", colour = "#2373b6", linewidth = 0.235) +
  scale_x_continuous(
    name = "humanGeneSite",
    breaks = plotDat$row,
    labels = plotDat$humanGeneSite,
    sec.axis = sec_axis(~., 
                        name = "yeastGeneSite", 
                        breaks = plotDat$row, 
                        labels = plotDat$yeastGeneSite)
  ) +
  comfy_theme(rotate_x_text = T, x_text_angle = 90) +
  theme(panel.background = element_rect(colour = "black", fill = NA, linewidth = 0.235),
        axis.text.x.top = element_text(size = 7, colour = "black", angle = 90, vjust = 0.5, hjust = 0)) +
  labs(y = "Ubi-site functional score")
ggsave_pdfpng(file = "output/images/experiments/bplot_expSites_wScores",
              width = 2.25, height = 2.5)

# Just one axis  
plotDat %>%
  mutate(humanYeastSite = factor(humanYeastSite, levels = unique(humanYeastSite[order(val)]))) %>%
  ggplot(aes(y = humanYeastSite, x = val)) +
    geom_col(width = 0.75, position = position_dodge2(), 
             fill = "#b3d1e0", colour = "#2373b6", linewidth = 0.235) +
    scale_x_continuous(breaks = c(0, 0.5, 1.0)) +
    comfy_theme() +
    labs(x = "Ubi-site functional score", y = "Yeast ubi-site (human ubi-site)")
ggsave_pdfpng(file = "output/images/experiments/bplot_expSites_wScores_oneYaxis",
              width = 2.5, height = 2.4)

```