# Background
Here I calculate conservation levels from the compara alignment results


# Initialise
```{r}
setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # Set working directory to source file location
library(tidyverse)
library(purrr)
library(data.table)


home_directory <- ""
source(paste0(home_directory, "scripts/essential_visualisations_JvG.R"))

#Directory for results from ptmAge code
ubiAgeDir <- ""   # Directory with my results
speciesTag <- "PRIDE" # Tag used to identify results
```


# Load data 
My data
```{r}
# Table of all sites with their gene families and alignment positions
ancest <- read_delim(paste0(ubiAgeDir, "results\\all_origins_", speciesTag, "_w0.tab")) %>% as.matrix

# Map from genes to their directories
map <- read_delim(paste0(ubiAgeDir, "temp\\Gene_to_tree_file_human.txt"), col_names = F) %>% as.data.frame
## Trim directories (user-specific)
CORRECTIONTAG <- ""
map[, 3] <- gsub(CORRECTIONTAG, "", map[, 3], fixed = TRUE)
map <- as.matrix(map)
```

# Clean ancest data
```{r}
####Clean the spaces in the column variable and create an IDposition like identifier for site recognition
ancest[,"column"]=gsub(" ","",ancest[,"column"])
ancest[,"position"]=gsub(" ","",ancest[,"position"])
ancest=cbind(paste(ancest[,"family"],ancest[,"column"],sep="_"),ancest)
colnames(ancest)[1]="ID"
ancest=cbind(ancest,paste(ancest[,"gene"],ancest[,"position"],sep="_K"))
colnames(ancest)[ncol(ancest)]="IDpos"

######We select all alignments which includes human sequences
ancest.homo=ancest[grep("Homo",ancest[,"species"]),]

######Overview for sanity check
# Human sites
dim(ancest.homo)
length(unique(ancest.homo[, "IDpos"]))
sum(is.na(ancest.homo[, "family"]))
# Non-human sites
ancest %>% as.data.frame %>% subset(species != "Homo sapiens") %>% .[, "species"] %>% table
```



## Collapse sites
I use my faster implementation
```{r}
column <- ancest %>% as.data.frame %>% group_by(ID) %>%  
    summarise(
        Aligned.sites.count = n(),
        Organisms = paste(species, collapse = ";"),
        IDpos = paste(IDpos, collapse = ";")
        #common_ancestor_name = paste(common_ancestor_name, collapse = ";")
    ) %>% ungroup %>% 
    subset(Aligned.sites.count > 1)  %>% 
    as.matrix

###Filtering all sites for human
column.human=column[grep("Homo",column[,"Organisms"]),]
View(column.human)
```

# Prepare the map
```{r}
###Lines to finaly calculate the boxes 
###Includes the one to navigate the subfolder system that the script has created with the alignments

###This first part creates arrangements to navigate the system I was working at, tunned as required
new.map=map[!duplicated(map[,2]),2:3]
colnames(new.map)=c("family","path")
new.map[, "path"] <- paste0(new.map[, "path"], "/", speciesTag, "/region_w0/", new.map[, "family"],"_ptm_continue_region_w0.txt")
```


# Navigating through the map to collect alignment data
```{r}
##Geting ready fo rthe final big tab

# Subset for gene families with PTMs
## This is done by using the "ancest" file, which has PTMs as rows
new.map=new.map[new.map[,"family"]%in%ancest[,"family"],]
```

```{r}
# Set up arguments to read in and manipulate data
argList <- list()
for (i in 1:nrow(new.map)){
  argList[[i]] <- c(new.map[i, "path"], new.map[i, "family"])
}
# Loop and read in data
primerLst <- list()
for (i in 1:length(argList)){
  if (i %% 100 == 0){
    print(i)
  }
  args <- argList[[i]]
  path <- args[1]
  family <- args[2]
  dat <- read.delim(path)
  datLong <- pivot_longer(dat, cols = -Gene)
  datLong$family <- family
  primerLst[[i]] <- datLong
}
# Combine
primer <- rbindlist(primerLst) %>% 
  dplyr::rename("gene" = "Gene", "column" = "name")
primer$column <- gsub("^X", "", primer$column)
primer$ID <- paste(primer$family, primer$column, sep = "_")
primer <- as.matrix(primer)
```


# Assign species (looks at gene naming)
```{r}
##Annotate with organism names based on ID features

primer=cbind(primer,"")

colnames(primer)[ncol(primer)]="species"

primer[grep("ENSP",primer[,"gene"]),"species"]="Homo sapiens"
primer[grep("ENSMUSP",primer[,"gene"]),"species"]="Mus musculus"
primer[grep("ENSRNOP",primer[,"gene"]),"species"]="Rattus norvegicus"
primer[grep("ENSGALP",primer[,"gene"]),"species"]="Gallus gallus"
primer[grep("ENSECAP",primer[,"gene"]),"species"]="Equus caballus"
primer[grep("ENSDARP",primer[,"gene"]),"species"]="Danio rerio"
primer[grep("ENSXETP",primer[,"gene"]),"species"]="Xenopus tropicalis"
primer[grep("FBpp",primer[,"gene"]),"species"]="Drosophila melanogaster"

primer[primer[,"species"]=="","species"]="Rest"
```


# Assign boxes 
## Set up
```{r}
# Set up
primerDf <- data.frame(primer) %>% 
  mutate(value = as.numeric(value))
column.humanDf <- as.data.frame(column.human)
ancest.homoDf <- as.data.frame(ancest.homo)

# Get human sites that also have mouse/rat or other species sites
MRposID <- column.humanDf[
  grepl("Mus", column.humanDf$Organisms) | 
  grepl("Rattus", column.humanDf$Organisms), 
  "ID"]
MRposID <- unique(MRposID)
CDWYposID <- column.humanDf[
  grepl("Gallus", column.humanDf$Organisms) | 
  grepl("Drosophila", column.humanDf$Organisms) |
  grepl("Caenorhabditis", column.humanDf$Organisms) | 
  grepl("Saccharomyces", column.humanDf$Organisms), 
  "ID"]
CDWYposID <- unique(CDWYposID)
MRE_lys <- unique(subset(primerDf, # All IDs with mouse/rat/horse acceptors 
  species %in% c("Mus musculus", "Rattus norvegicus", "Equus caballus") & 
  value == 0.5)$ID)  
anyNoLys <- unique(subset(primerDf, value == 0)$ID)  # Sites with no lys in any gene in any species
otherSpecies_lys <- unique(subset(primerDf, # All IDs with a lysine in other species
  species %in% c("Homo sapiens", "Mus musculus", "Rattus norvegicus", "Equus caballus") == F &
  value > 0)$ID)

# Asign different values to ancest.homoDf
ancest.homoDf$hasUbi_MR <- ancest.homoDf$ID %in% MRposID
ancest.homoDf$hasUbi_CDWY <- ancest.homoDf$ID %in% CDWYposID
ancest.homoDf$hasLys_MRE <- ancest.homoDf$ID %in% MRE_lys
ancest.homoDf$hasLys_CDWYetc <- ancest.homoDf$ID %in% otherSpecies_lys
ancest.homoDf$hasUbi_hsOnly <- !(ancest.homoDf$hasUbi_CDWY) & !(ancest.homoDf$hasUbi_MR)
```

## Ubi-sites + Lysine content
### Calculate
```{r}
# Calculate % lysine content for non-human sites
primerDfnoHs <- subset(primerDf, species != "Homo sapiens")
nonHsLysContent <- primerDfnoHs %>% group_by(ID) %>% 
  summarise(totalNum = n(), lysNum = sum(value > 0)) %>% ungroup %>% 
  mutate(lysContent = lysNum / totalNum * 100)
subset(nonHsLysContent, is.na(lysContent)) %>% View
# Map in
ancest.homoDf <- left_join(ancest.homoDf, nonHsLysContent, by = "ID")
ancest.homoDf$lysContent[which(is.na(ancest.homoDf$lysContent))] <- 0 # Set to 0 for sites with only human genes in alignment
```


### Define categories based on lysine content
```{r}
# Function to filter for lysine content
lysContFilt <- function(data, hasUbi_MR, hasUbi_CDWY, lysContent){
  bool <- (data$hasUbi_MR == hasUbi_MR) &  #Ubi-site filtering
    (data$hasUbi_CDWY == hasUbi_CDWY) & 
    (data$lysContent >= lysContent[1]) &  # Lysine content lower bound
    (data$lysContent < lysContent[2])   # Lysine content upper bound
}
lysContGroups <- list(
  "Level1" = lysContFilt(ancest.homoDf, F, F, c(0, 10)),
  "Level2" = lysContFilt(ancest.homoDf, F, F, c(10, 50)),
  "Level3" = lysContFilt(ancest.homoDf, F, F, c(50, 101)),
  "Level4" = lysContFilt(ancest.homoDf, T, F, c(0, 101)),
  "Level5" = lysContFilt(ancest.homoDf, T, T, c(0, 101))
)
ancest.homoDf$LevelsLyscont <- NA 
for (i in 1:length(lysContGroups)){
  ancest.homoDf$LevelsLyscont[lysContGroups[[i]]] <- names(lysContGroups)[i]
}
```



# Map back to uniprot and export
```{r}
# Load identifier map
load("../../data/PRIDE/PRIDE_uniprotToEnsp__3.RData")

# Map
ancest.homoDfMapped <- ancest.homoDf %>% 
  mutate(ensp_pos = paste(gene, position, sep = "_")) %>% 
  left_join(., PRIDE_uniprotToEnsp, by = "ensp_pos") %>% 
  mutate(uniprot_site = gsub("_", "_K", uniprot_pos))

# Rename
conservationBoxesDf <- ancest.homoDfMapped %>% 
  dplyr::rename("consBox_v1" = "FinalBox",
                "consBox_LysCont" = "LevelsLyscont")

# Collapse redundancy at the level of uniprot_site
## Order so that we always pick the highest conservation box when given a choice
conservationBoxesDfUniq <- conservationBoxesDf[order(conservationBoxesDf$consBox_LysCont, decreasing = T), ]
## Collapse at the level of uniprots
conservationBoxesDfUniq <- conservationBoxesDfUniq %>% 
  group_by(uniprot_site) %>% 
  summarise(across(everything(), ~ first(.))) %>% 
  ungroup

# Export
save(conservationBoxesDf, file = "results/conservationBoxesDf__1.RData")
save(conservationBoxesDfUniq, file = "results/conservationBoxesDfUniq__1.RData")
```